### 2012-04-19 docs/proposal/paper.tex
Added proposal and related work writeups

\documentclass[preprint]{acm_proc_article-sp}
%\documentclass[preprint]{sig-alternate}
\usepackage{url}
\usepackage{graphicx,subfigure}
\usepackage{xspace}

\newcommand{\ie}{{\em i.e.,}~}
\newcommand{\eg}{{\em e.g.,}~}

\newenvironment{denseitemize}{
\begin{itemize}[topsep=2pt, partopsep=0pt, leftmargin=1.5em]
  \setlength{\itemsep}{4pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{itemize}}

\newcommand{\eat}[1]{}

\begin{document}

\title{CS 294-1 Project Proposal: \\
Detecting Spam on Social Networking Sites}

\numberofauthors{3}
\author{
Antonio Lupher,
Cliff Engle,
Reynold Xin\\\\
\texttt{\{alupher, cengle, rxin\}@cs.berkeley.edu}
}

\maketitle

\section{Problem}

Most social networks of any significant size see constant spam, scams
and phishing attacks. The nature of these attacks can be quite diverse
and difficult detect. Marketers can spam members with unwanted
advertisements, fraudsters lure users with advance fee frauds and
other confidence tricks, while others attempt to steal user
information by directing users to external phishing pages. To make
matters more difficult, a site with global reach sees communication
among its members in a number foreign languages with varying levels
of ability. This means that much benign content shares characteristics
like misspellings, awkward phrases, etc. that might have made certain
types of common frauds and spam more easy to distinguish on US-based
(or English-language) sites.

\section{Approach}

This project will examine in detail the types of malicious and benign
content that are encountered on social networks by analyzing
experimental data available from InterPals, an international social network
for cultural exchange and language practice. For example, the site
attracts a wide variety of financial scams, ranging from Nigerian
"419" scams to romance scams. Another prevalent problem is spam with
links to third-party websites, directing users to various porn/webcam
sites, phishing sites or various untrustworthy online marketplaces.

We will then examine various methods of detecting and preventing abuse
on the site, including those measures that have already been taken
(e.g. various heuristics including IP/location anomaly detection,
frequency capping, duplicate account detection, etc.). However, the
main focus will be on mining experimental data from the site and using
features derived from this data to build and evaluate classifiers to
detect unwanted behavior programmatically. The large volume of data
available to us will provide a unique perspective both on the types of
malicious content that exist on such sites as well as on the
effectiveness of classifier/learning-based approaches to identifying
these activities.

\section{Data sets}

We plan to make heavy use of our unrestricted access to the data of
InterPals, which has over 1.2 million active members. This data
includes a corpus of 90 million private messages and another 1.5
million messages that have been labeled as spam by users. Other data
includes 40 million or so "wall" comments, 5 million photos, and 8
million photo comments.

\section{Techniques}

Using this data, we plan to first identify the most prevalent types of 
malicious activities on the site. We will investigate various machine learning 
techniques to automatically detect these activities, including sampling, clustering and 
classifiers. Sampling and clustering, in addition to the user-labeled and moderator-verified 
spam corpus, will help us identify training data.

Classifiers that we plan to explore 
include:

\begin{itemize}
\item Gradient boosting
\item Naive Bayes
\item SVMs 
\item Decision trees
\end{itemize}

We will then test the classifiers on new site data to evaluate their performance.

One important part of this project will be to identify features of undesirable activity that are useful in classification. We plan to examine:
\begin{itemize}
\item Profile and message data (keywords, n-grams, age, sex, registration date, amount of data in profile, etc.)
\item User info (geographical location, IPs, browser settings)
\item Complaints and user labeling of spam
\item Profile photographs
\item User reputation, social graph
\end{itemize}

\section{Frameworks}
We hope to implement the detection algorithms using Spark, an
in-memory distributed computing framework that is particularly
well-suited for machine learning and iterative computations.

\balancecolumns
\end{document}

### 2012-04-19 docs/related/paper.tex
Added proposal and related work writeups

\documentclass[preprint]{acm_proc_article-sp}
%\documentclass[preprint]{sig-alternate}
\usepackage{url}
\usepackage{graphicx,subfigure}
\usepackage{xspace}

\newcommand{\ie}{{\em i.e.,}~}
\newcommand{\eg}{{\em e.g.,}~}

\newenvironment{denseitemize}{
\begin{itemize}[topsep=2pt, partopsep=0pt, leftmargin=1.5em]
  \setlength{\itemsep}{4pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{itemize}}

\newcommand{\eat}[1]{}

\begin{document}

\title{Detecting Spam on Social Networking Sites:\\
Related Work}

\numberofauthors{3}
\author{
Antonio Lupher,
Cliff Engle,
Reynold Xin\\\\
\texttt{\{alupher, cengle, rxin\}@cs.berkeley.edu}
}

\maketitle

\section{Related Work}
The rise of social media has made Social Networking Services (SNSs) more attractive targets for spam and fraud, leading to increasingly sophisticated attacks. This trend is reflected in recent research, as papers have focused on identifying and classifying the various types of social media spam. Many of these studies employ techniques previously used to combat conventional email and web spam. SNSs also provide opportunities to take advantage of user reputation and other social graph-dependent features to improve classification. Nevertheless, most research has been carried out on publicly-available data from SNSs, making it difficult up until now to measure the effect of private user data on algorithms for detecting site misuse.

\subsection{Social Spam Features}

Heymann et al. \cite{heymann} survey the field of spam on SNSs, identifying several common approaches. Identification-based approaches identify spam to train classifiers based on labels submitted by users or trusted moderators. Rank-based approaches demote visibility of questionable content, while interface-based approaches apply policies to prevent unwanted behavior. This work groups classification-based approaches with detection, although classifiers can be used in conjuction with user information to prevent spam before it happens.

A number of researchers have focused on collecting, identifying features and classifying various genres of spam on social networks. Zinman and Donath \cite{zinman} extract bundles of profile-based and comment-based features from MySpace profiles, but the relatively poor performance of their classifier highlights the difficulties in manual classification social network spam. Several studies take the approach of baiting spammers with social ``honeypots", profiles created with the sole intent of attracting spam.\cite{stringhini, lee} They then use the data collected to train classifiers with features including friend request rate and ratios of URLs to text. Webb et al. \cite{webb} use the honeypot approach as well and provide examples of various types of spammers, the typical demographics of their profiles as well as the web pages that they tend to advertise.

Gao et al. \cite{gao} look at Facebook wall posts, analyzing temporal properties, URL characteristics, post ratios and other features of malicious accounts. They also pinpoint various spam ``campaigns" based on products advertised in a given time frame. They note that spam on Facebook often exhibits burstiness and is mainly sent from compromised accounts. Benevenuto et al. \cite{benevenuto} identify attributes of spam on video SNSs and use a Support Vector Machine (SVM) for classification.

Not all undesirable content on SNSs is necessarily spam or a scam. SNSs and online communities witness inappropriate user behavior, where users post offensive and harassing content. Yin et al. \cite{yin} combine sentiment analysis and profanity word lists with contextual features to identify harassment on datasets from Slashdot and MySpace. Other work looks at SNSs as platforms to collect data about users in order to aid direct attacks on the user's computers or to compromise a large number of accounts. \cite{patsakis, huber}

\subsection{Social Spam Detection Systems}

SocialSpamGuard \cite{jin} is a social media spam detection system that analyzes text and image features of social media posts. The demo system uses GAD clustering \cite{jingad} for sampling spam and ham posts, then trains a classifier with text and image features. However, the system is built on top of Facebook features that are publicly accessible and thus cannot make use of sensitive user data (\eg IP addresses) to increase its effectiveness.

De Wang et al. \cite{wang} propose a cross-site spam detection framework to share spam data across all social networking sites, building classifiers to identify spam in profiles, messages and web pages. This multi-pronged approach lends itself to associative classification, in which, for example, a message would be classified as spam if it contained a link to a web page that had a high probability of being spam. Unfortunately, the differing characterestics of various social networks \eg the length of messages in Facebook vs. Twitter, can reduce the benefits of sharing spam corpora across diverse sites.

Facebook \cite{stein} provides an overview of their ``immune system" defences against phishing, fraud and spam. The system is composed of classifier services, an ML-derived Feature Extraction Language (FXL), feature loops to aggregate and prepare features for classification and a policy engine to take action on suspected misuse. While the discussion remains high-level and includes few implemention particulars, it does include significant detail on the various types and characterists of undesirable activity on the site, including fake profiles, harassment, compromised accounts, malware and spam.

In contrast to research that focuses on dynamically detecting spam based on user activity, Irani et al. \cite{irani} show that static features associated with user signups on MySpace are enough to train an effective social spam classifier. They note that C4.5 decision tree algorithms provide better performance than naive Bayes in this case. As in other works, this only examines publicly available profile information collected by social honeypots. Private data collected on users including browser features, IP addresses and geographic location would conceivably improve classifier performance substantially.

Bosma et al. \cite{bosma} explore user-generated spam reports as a tool for building an unsupervised spam detection framework for SNSs. Their approach counts the number of spam reports against a suspected spammer and adds weight to reports based on user reputation. Determining reputation and trustworthiness of users in social networks has been well studied \cite{bian, guha, zhang} and appears to be a promising addition to social spam classification. The framework uses a Bayesian classifier and links messages with similar content, but does not take into account other features. Nevertheless, this is one of the few studies to test its framework on non-public data, inlcuding private messages, spam reports and user profiles from a large Dutch social networking site.

\subsection{Spam Email \& Web}

Much work has been done on protecting traditional email systems from spam. Blanzieri \cite{blanzieri} offers a comprehensive overview of machine learning techniques that can be applied to email filtering. Hao et al. \cite{hao} describe a reputation engine based on lightweight features such as geographic distance between sender and receiver, geolocation anomalies and diurnal patterns. While the target was conventional spam, these and similar features are applicable to spam on SNSs as well.

Whittaker et al. \cite{whittaker} describe a scalable phishing machine learning classifier and blacklisting system with high accuracy. Since a considerable amount of social media spam includes links to phishing sites, being able to detect them is critical. Along similar lines, Monarch \cite{thomas} is a system that provides scalable real-time detection of URLs that point to spam web pages as determined by URL features, page content and hosting properties of the target domain.

Blog comment spam have also attracted considerable attention from researchers who have applied machine learning \cite{kolari, nag} and NLP \cite{mishne} techniques to the problem. Likewise, Markines et al. \cite{markines} apply similar techniques to spam on social bookmarking sites.

\subsection{Machine Learning and Data Mining}

Many of the data mining algorithms used to detect spam and patterns of misuse on SNSs are designed with the assumption that the data and the classifier are independent. However, in the case of spam, fraud and other malicious content, users will often modify their behavior to evade detection, leading to degraded classifier performance and the need to re-train classifiers frequently. Several researchers tackle this adversarial problem. Dalvi et al. \cite{dalvi} offer a modified Naive Bayes classifier to detect and reclassify data taking into account the optimal modification strategy that an adversary could choose. Lowd and Meek \cite{lowd} provide a framework for reverse engineering a classifier to determine whether an adversary can efficiently learn enough about a classifier to effectively defeat it.

\bibliographystyle{abbrv}
\bibliography{paper}

\balancecolumns
\end{document}

### 2012-05-08 docs/paper/paper.tex
Latest paper draft

\documentclass[preprint]{acm_proc_article-sp}
%\documentclass[preprint]{sig-alternate}
\usepackage{url}
\usepackage{graphicx,subfigure}
\usepackage{xspace}

\newcommand{\ie}{{\em i.e.,}~}
\newcommand{\eg}{{\em e.g.,}~}

\newenvironment{denseitemize}{
\begin{itemize}[topsep=2pt, partopsep=0pt, leftmargin=1.5em]
  \setlength{\itemsep}{4pt}
  \setlength{\parskip}{0pt}
  \setlength{\parsep}{0pt}
}{\end{itemize}}

\newcommand{\eat}[1]{}

\begin{document}

\title{Project Report: \\
Detecting Spam on Social Networking Sites}

\numberofauthors{3}
\author{
Antonio Lupher,
Cliff Engle,
Reynold Xin\\\\
\texttt{\{alupher, cengle, rxin\}@cs.berkeley.edu}
}

\maketitle

\section{Introduction}

\subsection{Problem}

Social networks of any significant size see constant spam, scams
and phishing attacks. The nature of these attacks can be quite diverse
and difficult detect. Marketers can spam members with unwanted
advertisements, fraudsters lure users with advance fee frauds and
other confidence tricks, while others attempt to steal user
information by directing users to external phishing pages. To make
matters more difficult, a site with global reach sees communication
among its members in a number foreign languages with varying levels
of ability. This means that much benign content shares characteristics
like misspellings, awkward phrases, etc. that might have made certain
types of common frauds and spam more easy to distinguish on US-based
(or English-language) sites.
% focus on extra impact of non-public features vs. public (bag-of-words)

\subsection{Approach}

This project examines in detail the types of malicious and benign
content that are encountered on social networks by analyzing
experimental data available from InterPals, an international social network
for cultural exchange and language practice. For example, the site
attracts a wide variety of financial scams, ranging from Nigerian
"419" scams to romance scams. Another prevalent problem is spam with
links to third-party websites, directing users to various porn/webcam
sites, phishing sites or various untrustworthy online marketplaces.

We will then examine various methods of detecting and preventing abuse
on the site, including those measures that have already been taken
(e.g. various heuristics including IP/location anomaly detection,
frequency capping, duplicate account detection, etc.). However, the
main focus will be on mining experimental data from the site and using
features derived from this data to build and evaluate classifiers to
detect unwanted behavior programmatically. The large volume of data
available to us will provide a unique perspective both on the types of
malicious content that exist on such sites as well as on the
effectiveness of classifier/learning-based approaches to identifying
these activities.

\subsection{Data sets}

We plan to make heavy use of our unrestricted access to the data of
InterPals, which has over 1.3 million active members. This data
includes a corpus of over a 100 million private messages and another 2 
million messages that have been labeled as spam by users. Other data
includes 40 million or so "wall" comments, 5 million photos, and 8
million photo comments.

\subsection{Techniques}

Using this data, we plan to first identify the most prevalent types of 
malicious activities on the site. We will investigate various machine learning 
techniques to automatically detect these activities, including sampling, clustering and 
classifiers. Sampling and clustering, in addition to the user-labeled and moderator-verified 
spam corpus, will help us identify training data.

Classifiers that we plan to explore 
include:

\begin{itemize}
\item Naive Bayes
\item Regression
\item SVMs 
\end{itemize}

We then test the classifiers on new site data to evaluate their performance.

One important part of this project will be to identify features of undesirable activity that are useful in classification. We plan to examine:

\subsection{Frameworks}
We implement the detection algorithms using Matlab, ScalaNLP, LibSVM, and Spark, an in-memory distributed computing framework designed for machine learning and iterative computation.

\section{Background}

The rapid growth of social media has made Social Networking Services (SNSs) increasingly attractive targets for spam and fraud, leading to a proliferation of sophisticated attacks. This trend is reflected in recent research, as papers have focused on identifying and classifying the various types of social media spam. Many of these studies employ techniques previously used to combat conventional email and web spam. SNSs also provide opportunities to take advantage of user reputation and other social graph-dependent features to improve classification. Nevertheless, most research has been carried out on publicly-available data from SNSs, making it difficult up until now to measure the effect of private user data on algorithms for detecting site misuse.

\subsection{Social Spam Features}

Heymann et al. \cite{heymann} survey the field of spam on SNSs, identifying several common approaches. Identification-based approaches identify spam to train classifiers based on labels submitted by users or trusted moderators. Rank-based approaches demote visibility of questionable content, while interface-based approaches apply policies to prevent unwanted behavior. This work groups classification-based approaches with detection, although classifiers can be used in conjuction with user information to prevent spam before it happens.

A number of researchers have focused on collecting, identifying features and classifying various genres of spam on social networks. Zinman and Donath \cite{zinman} extract bundles of profile-based and comment-based features from MySpace profiles, but the relatively poor performance of their classifier highlights the difficulties in manual classification social network spam. Several studies take the approach of baiting spammers with social ``honeypots", profiles created with the sole intent of attracting spam.\cite{stringhini, lee} They then use the data collected to train classifiers with features including friend request rate and ratios of URLs to text. Webb et al. \cite{webb} use the honeypot approach as well and provide examples of various types of spammers, the typical demographics of their profiles as well as the web pages that they tend to advertise.

Gao et al. \cite{gao} look at Facebook wall posts, analyzing temporal properties, URL characteristics, post ratios and other features of malicious accounts. They also pinpoint various spam ``campaigns" based on products advertised in a given time frame. They note that spam on Facebook often exhibits burstiness and is mainly sent from compromised accounts. Benevenuto et al. \cite{benevenuto} identify attributes of spam on video SNSs and use a Support Vector Machine (SVM) for classification.

Not all undesirable content on SNSs is necessarily spam or a scam. SNSs and online communities witness inappropriate user behavior, where users post offensive and harassing content. Yin et al. \cite{yin} combine sentiment analysis and profanity word lists with contextual features to identify harassment on datasets from Slashdot and MySpace. Other work looks at SNSs as platforms to collect data about users in order to aid direct attacks on the user's computers or to compromise a large number of accounts. \cite{patsakis, huber}

\subsection{Social Spam Detection Systems}

SocialSpamGuard \cite{jin} is a social media spam detection system that analyzes text and image features of social media posts. The demo system uses GAD clustering \cite{jingad} for sampling spam and ham posts, then trains a classifier with text and image features. However, the system is built on top of Facebook features that are publicly accessible and thus cannot make use of sensitive user data (\eg IP addresses) to increase its effectiveness.

De Wang et al. \cite{wang} propose a cross-site spam detection framework to share spam data across all social networking sites, building classifiers to identify spam in profiles, messages and web pages. This multi-pronged approach lends itself to associative classification, in which, for example, a message would be classified as spam if it contained a link to a web page that had a high probability of being spam. Unfortunately, the differing characterestics of various social networks \eg the length of messages in Facebook vs. Twitter, can reduce the benefits of sharing spam corpora across diverse sites.

Facebook \cite{stein} provides an overview of their ``immune system" defences against phishing, fraud and spam. The system is composed of classifier services, an ML-derived Feature Extraction Language (FXL), feature loops to aggregate and prepare features for classification and a policy engine to take action on suspected misuse. While the discussion remains high-level and includes few implemention particulars, it does include significant detail on the various types and characterists of undesirable activity on the site, including fake profiles, harassment, compromised accounts, malware and spam.

In contrast to research that focuses on dynamically detecting spam based on user activity, Irani et al. \cite{irani} show that static features associated with user signups on MySpace are enough to train an effective social spam classifier. They note that C4.5 decision tree algorithms provide better performance than naive Bayes in this case. As in other works, this only examines publicly available profile information collected by social honeypots. Private data collected on users including browser features, IP addresses and geographic location would conceivably improve classifier performance substantially.

Bosma et al. \cite{bosma} explore user-generated spam reports as a tool for building an unsupervised spam detection framework for SNSs. Their approach counts the number of spam reports against a suspected spammer and adds weight to reports based on user reputation. Determining reputation and trustworthiness of users in social networks has been well studied \cite{bian, guha, zhang} and appears to be a promising addition to social spam classification. The framework uses a Bayesian classifier and links messages with similar content, but does not take into account other features. Nevertheless, this is one of the few studies to test its framework on non-public data, inlcuding private messages, spam reports and user profiles from a large Dutch social networking site.

\subsection{Spam Email \& Web}

Much work has been done on protecting traditional email systems from spam. Blanzieri \cite{blanzieri} offers a comprehensive overview of machine learning techniques that can be applied to email filtering. Hao et al. \cite{hao} describe a reputation engine based on lightweight features such as geographic distance between sender and receiver, geolocation anomalies and diurnal patterns. While the target was conventional spam, these and similar features are applicable to spam on SNSs as well.

Whittaker et al. \cite{whittaker} describe a scalable phishing machine learning classifier and blacklisting system with high accuracy. Since a considerable amount of social media spam includes links to phishing sites, being able to detect them is critical. Along similar lines, Monarch \cite{thomas} is a system that provides scalable real-time detection of URLs that point to spam web pages as determined by URL features, page content and hosting properties of the target domain.

Blog comment spam have also attracted considerable attention from researchers who have applied machine learning \cite{kolari, nag} and NLP \cite{mishne} techniques to the problem. Likewise, Markines et al. \cite{markines} apply similar techniques to spam on social bookmarking sites.

\subsection{Machine Learning and Data Mining}

Many of the data mining algorithms used to detect spam and patterns of misuse on SNSs are designed with the assumption that the data and the classifier are independent. However, in the case of spam, fraud and other malicious content, users will often modify their behavior to evade detection, leading to degraded classifier performance and the need to re-train classifiers frequently. Several researchers tackle this adversarial problem. Dalvi et al. \cite{dalvi} offer a modified Naive Bayes classifier to detect and reclassify data taking into account the optimal modification strategy that an adversary could choose. Lowd and Meek \cite{lowd} provide a framework for reverse engineering a classifier to determine whether an adversary can efficiently learn enough about a classifier to effectively defeat it.

\section{Data set}

For this project, we had unlimited access to data from InterPals, a SNS for users who wish to communicate with others from other countries, whether for language practice, cultural exchange or friendship. Users sign up by completing a registration form with information about themselves, including age, sex and location. After registering, users can expand their profile page to include descriptions, interests, languages they speak, etc. and upload photographs. After clicking on a link sent to their email addres, a user can begin to interact with others on the site via private message, public ``wall" posts, comments on photos and on a bulletin board system.

Currently, the site combats spam and other Terms of Service violations through volunteer moderators. Users can report content, including private messages, profiles and photographs to moderators, using a form that includes a drop-down menu of pre-selected reasons with the option to add a more detailed message in a text field. Likewise, the message interface allows a user to report a private message spam with a single click. Moderators have access to a queue of these reports. In addition to the material being reported, moderators are able to make decisions based on the data from the reported user's account, including outgoing private messages, IP addresses, as well as a list of other users who have logged in from the same computer (determined via IP address as well as by browser fingerprints). Moderators can then decide to delete the user, send a warning, or clear reports on a user. All moderator actions require them to annotate their decision with a brief log message.

The data in our

\subsection{Categories of Spam}

Our ongoing observation of spam on the InterPals website provided direct intuition into the major classes of unwanted behavior that we classified as spam for this project. We noted the following broad categories:

\textbf{Advance Fee Fraud} We observed this scam to come in various flavors.  (inheritance, lottery,  etc)
\textbf{Romance Scam} 
\textbf{Sexual Solicitation}
\textbf{Porn Ads}
\textbf{Advertising External Sites}
\textbf{Webcam Ads}
\textbf{Money Muling}
\textbf{Begging \& Gift Requests}
\textbf{Business Proposals}

This categorizations are based only on observation. Analyzing the relevant terms in log comments entered by moderators when deleting spammer accounts provides the following breakdown:

\subsection{Spam Set}

\subsection{Ham Set}

To simplify the labeling of ham messages, we made the assumption that messages remaining in the inboxes of active users after a period of several months would most likely not be spam. Consequently, we extracted a working set of 2 million ham messages

\section{Methodology}

We decided to focus on spam in private messages for this paper, given that messages account for a majority of user spam reports. In addition to using the bag-of-words representation of the message content, we aimed to identify a subset of additional, relevant ``expert" features based on public and non-public message and account information that would augment the classifier's accuracy. We first extracted a number of fields that we expected might improve classification. To choose a sample of relevant features, we computed statistics on the extracted features, both on the ham and spam corpora.

\subsection{Message Features}

\textbf{Message Body \& Subject:} We extracted both the message

\textbf{Sender IP:} The IP address of the sender at the time that the user sent the message.

\textbf{Recipient Age:} The age listed on the message recipient's profile.

\textbf{Recipient Sex:} The sex listed on the message recipient's profile.

\textbf{Recipient Country:} This was the ISO 3166-1 alpha-2 code of the recipient of the message.

\textbf{Recipient Replied:} A boolean, 1 or 0, depending on whether the recipient replied to the message.

Due to data collection issues, recipient data (age, sex, country) was only available for the most recent 11,000 spam messages.

\subsection{Sender Account Features}

\textbf{Sender IP:} Spammers in general had a smaller number of unique IP addresses associated with them than ham users. This is likely due to their comparatively low length of time on the site and the fact that many legitimate users are behind NATs that yield a high number of dynamic IP addresses. One possibility for future research is to investigate the ratio of unique IPs for users as a function of the time they have been registered. Likewise, examining unique /16 network blocks instead of unique IP addresses could mitigate the influence of NATs.

\textbf{Sender Email:} We found that spammers were

\textbf{Sender Age:}

\textbf{Sender Country:}

\textbf{Sender Birthday:}

\textbf{Sender Age:}

\textbf{Sender Name:}

\textbf{Sender Username:}

\textbf{Sender Sex:}

\textbf{Sender Sex:}

\subsection{Mixed Features}
\textbf{Account Lifetime.}

\textbf{IP Mismatch.} This

\section{Classification}

\subsection{Naive Bayes}

% Modified from Reynold's HW1  write-up
We use a na\"ive Bayes model, in which the documents are generated according to the following process.  
First, a class \(c\) is chosen from a probability distribution \(\Pr(c)\).
Then, a document from class \(c\) is generated.
We tried a multinomial model and a Bernoulli model to model this process.
\begin{itemize}
\item \emph{Multinomial model.}  A document length \(m_d\) is chosen and \(m_d\) words \(w_{d,1}, \dotsc, w_{d,m_d}\) are chosen independently from a distribution \(\Pr(w|c)\).
\item \emph{Bernoulli model.}  For each word \(w\) in the vocabulary, \(w\) is added to the document with probability \(\Pr(w|c)\).
\end{itemize}
The distributions \(\Pr(c)\) and \(\Pr(w|c)\) are parameters of the model.

\subsection{Linear Regression}

% Modified from Cliff & Antonio's HW2 write-up
Linear regression is a method to model the relationship between an output variable, $y$, and explanatory variables, $X$. The output variable is a linear sum of the explanatory variables multiplied by their corresponding coeffecients, represented by $\beta$. In our case, $y$ represents a review's ranking from one to five, while $X$ is the bag-of-words representation of the review text. This implies that $\beta$ represents the weight of a word's correlation to a higher review. This gives us the simple regression formula: $$y=X\beta$$

In this project, we performed ordinary least squares regression. By minimizing the sum of squared residuals, we can solve for the unkown parameter vector $\beta$ using a closed form solution. $$\beta=(X'X)^{-1}X'y$$ In order to ensure that the matrix $X'X$ is invertible, we perform ridge regression. In ridge regression, we add the identity matrix, $I$, scaled by a factor, $\lambda$, to produce the following equation. $$\beta=(X'X+\lambda I)^{-1}X'y$$ We calculate this $\beta$ for a training set of review text along with their corresponding ratings which range from 1 to 5. We then evaluate our model $\beta$ on the withheld reviews in order to perform a 10-fold cross-validation.

All of our linear regression code was implemented in MATLAB because of its ease-of-use and optimizations on matrix multiplications. We parsed the reviews from the tokenized binary format into a bag-of-words representation stored as a sparse matrix. Matlab uses Compressed Sparse Column format, so we actually stored the matrix $X'$ because the dictionary is very sparse. We also filtered out duplicate reviews by hashing the first ten words in a review to determine distinctness. This filtered the one million reviews to 492,130.

The actual matrix multiplications were done using the sparse $X$ matrix along with our $y$ vector and the ridge parameter. The multiplication $X'X$ takes up too much memory if we use the entire dictionary, so we chose to limit the dictionary size. We found that a dictionary of the 10,000 most frequent words gave sufficient features while also being computable using MATLAB's built-in sparse matrix multiplications. We also tried using a dictionary of 15,000 words, but only had a marginal improvement along with much slower execution. Using our 10,000 word dictionary, we calculated the performance of our matrix multiply to be 0.123 Gflops/s by dividing the number of non-zero elements in our matrices by the time to compute the multiplication.

Since we used ridge regression, we had the ability to tune the $\lambda$ parameter in order to get the best performance. As shown in figure \ref{fig:lambda}, we observed an AUC improvement of about .5\% between the default value of 1 and a tuned parameter of 1500. This is a fairly major difference since the AUC is at 93\% using the default value. Additionally, we also noticed a major difference in the top words when using better $\lambda$ values. Using the default value of 1, our top words included obscure words including ``\_c'', ``\_j'', ``donovan'', and ``matlock.'' When we used the tuned $\lambda$ value of 1500, the words were much more representative, as shown in table \ref{tab:words}.

\subsection{Logistic Regression}

We attempted to train the multiple logistic regression classifier implemented by the ScalaNLP package, but encountered a bug that prevented the model from being generated correctly. With the expectation that SVMs would outperform logistic regression, we decided to focus our efforts on tuning the SVM instead of pursuing logistic regression.

\subsection{Support Vector Machines (SVMs)}

%intro to svms
SVMs

% scalanlp svm

% libsvm

LIBSVM is library for SVM training and classification in C++ and Java, with wrappers for Python and other languages. The package allows one to specify the type of kernel to use, kernel parameters ($\gamma$, $\rho$, $d$), as well as a penalty parameter ($C$). We used Perl to convert the sparse feature matrices generated for use with ScalaNLP to a format recognized by LIBSVM. We then used one of the package's tools to perform simple scaling on the data.

Given the complexity of choosing kernels and parameters, the package includes a Python script that performs cross-validation to suggest appropriate $\gamma$ and $C$ values for the recommended radial basis function (RBF) kernel. We ran this script on a subset of 2500 ham and 2500 spam messages to obtain $C = 8$ and $\gamma = 0.0078125$.

The LIBSVM scaling tool increases file size considerably for data sets with large feature matrices. This large file size, coupled with the computational resources required for SVM training on large data sets, prevented us from completing our analysis of the

\begin{tabular}{c|c|c}
    \multicolumn{3}{c}{Unscaled: Data set size, Accuracy (bag-of-words), Accuracy (expert features)}
    Data size & Accuracy, bag-of-words & Accuracy, expert features \\
    \hline
    2k & 0.0 & 0.0 \\
    10k & 0.0 & 0.0 \\
    22k & 0.0 & 0.0 \\ % note we chose 22k because 11k had the recipient info
    50k  & 0.0 & 0.0 \\
    100k  & 0.0 & 0.0 \\
\end{tabular}

\begin{tabular}{c|c|c}
    \multicolumn{3}{c}{Scaled: Data set size, Accuracy (bag-of-words), Accuracy (expert features)}
    Data size & Accuracy, bag-of-words & Accuracy, expert features \\
    \hline
    2k & 0.0 & 0.0 \\
    10k & 0.0 & 0.0 \\
    22k & 0.0 & 0.0 \\ % note we chose 22k because 11k had the recipient info
\end{tabular}

\section{Results}

\subsection{Linear Regression}

\begin{tabular}{c|c}
    \multicolumn{2}{c}{Tokens most likely to be positive} \\
    token & \(log(p_+ / p_-)\) \\
    \hline
    pos & 0.0 \\
    pos2 & 0.1 \\
\end{tabular}

\begin{tabular}{c|c}
    \multicolumn{2}{c}{Tokens most likely to be negative} \\
    token & \(log(p_+ / p_-)\) \\
    \hline
    neg & 0.0 \\
    neg2 & 0.1 \\
\end{tabular}

\subsection{SVMs}

\begin{tabular}{c|c}
    \multicolumn{2}{c}{Tokens most likely to be positive} \\
    token & \(log(p_+ / p_-)\) \\
    \hline
    pos & 0.0 \\
    pos2 & 0.1 \\
\end{tabular}

\begin{tabular}{c|c}
    \multicolumn{2}{c}{Tokens most likely to be negative} \\
    token & \(log(p_+ / p_-)\) \\
    \hline
    neg & 0.0 \\
    neg2 & 0.1 \\
\end{tabular}

\section{Analysis}
% Vern wants site owner perspective on how to apply classifiers
From a site's perspective, ...

\section{Future Work}

In this project, we chose only a small subset of ``expert" features to use. However, we believe that it would be beneficial to train and evaluate the classifiers on a wider range of features. The features would include both the ones we extracted in this project but omitted from the classifier features matrices, as well as new features. Adding N-grams and message similarity tests seem particularly promising. N-grams would allow the classifier to take into account certain phrases \eg{``Western Union" or ``money transfer"}.

We have also recently begun to collect browser fingerprints from users based on user agent and JavaScript-acquired browser plugin details, time zone, screen size, color depth, system fonts, keyboard layout and locale information. Subsequent classifiers could leverage this information either as a hash that effectively tags individual known spammers or in part, adding features based on time zone or keyboard layout.

Characteristics of spam messages vary widely depending on the type of spam
we currently lump all spam into a single category; we plan to build on the classes we mentioned in section()
We are currently in the process of extending the moderation interface to include more specific tagging capabilities
- certain classifiers might work better on certain types of spam
- clustering to see how well these classes fit the spam types on the site

Server access logs, search logs and user viewing histories from InterPals, all of which are available to us, offer another avenue for future research. It is conceivable that certain spammers may have similar site usage patterns, in terms of HTTP request intervals,

\section{Conclusion}

figure \ref{fig:cprof}
figure \ref{fig:cip}
figure \ref{fig:day}
figure \ref{fig:month}
figure \ref{fig:drop}
figure \ref{fig:email}
figure \ref{fig:friends}
figure \ref{fig:usaspam}
figure \ref{fig:usahamfig:lambda}
figure \ref{fig:roclin}
figure \ref{fig:photos}
figure \ref{fig:recipage}
figure \ref{fig:sendage}
figure \ref{fig:sendsex}
figure \ref{fig:uniqip}
\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/country-prof.pdf}
    \caption{Country as stated on user profile.}
    \label{fig:cprof}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/country-ip.pdf}
    \caption{Country detected by IP address using MaxMind GeoIP database.}
    \label{fig:cip}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/dob-day.pdf}
    \caption{Date of month from birthday submitted by user at signup.}
    \label{fig:day}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/dob-month.pdf}
    \caption{Month of birthday as submitted by user at signup.}
    \label{fig:month}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/dropdown.png}
    \caption{Drop-down select menu on signup form.}
    \label{fig:drop}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/email.pdf}
    \caption{Domain name of email provided at signup. Users cannot send messages before 
        activating their account via a link sent to their email address.}
    \label{fig:email}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/friends.pdf}
    \caption{Number of friends associated with sender account.}
    \label{fig:friends}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/ips-usa-spam.pdf}
    \caption{Distribution of IPs by country for spammers with profiles stating a USA location.}
    \label{fig:usaspam}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/ips-usa-ham.pdf}
    \caption{Distribution of IPs by country for ham users with profiles stating a USA location.}
    \label{fig:usaham}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/linear-ridge.pdf}
    \caption{AUC based on varying lambda values for Ridge regression.}
    \label{fig:lambda}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/linear-roc.pdf}
    \caption{Linear regression ROC curves for classifier using bag-of-words and bag-of-words with ``expert" features.}
    \label{fig:roclin}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/photos.pdf}
    \caption{Number of photos associated with sender account.}
    \label{fig:photos}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/recip-age.pdf}
    \caption{Age of message recipient as stated on user profile. Note that 2 million message spam sample 
        had only 11,000 messages with this information.}
    \label{fig:recipage}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/sender-age.pdf}
    \caption{Age of message sender as stated on user profile. }
    \label{fig:sendage}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/sex.pdf}
    \caption{Sender sex as stated on profile.}
    \label{fig:sendsex}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[width=\linewidth]{figures/unique-ips.pdf}
    \caption{Number of unique IP addresses associated with user account. Spam: mean 11.82, std. dev. 52.52. Ham: mean 175.51, std. dev. 322.19}
    \label{fig:uniqip}
\end{figure}

\bibliographystyle{abbrv}
\bibliography{paper}

\balancecolumns
\end{document}

### 2012-05-08 docs/paper/paper.tex
Latest paper draft.

@@
 \section{Introduction}
+% Modified from project proposal
 
 \subsection{Problem}
-
 Social networks of any significant size see constant spam, scams
@@
 
-This project examines in detail the types of malicious and benign
+This paper examines in detail the types of malicious and benign
 content that are encountered on social networks by analyzing
@@
 
-We will then examine various methods of detecting and preventing abuse
+We examine various methods of detecting and preventing abuse
 on the site, including those measures that have already been taken
 (e.g. various heuristics including IP/location anomaly detection,
-frequency capping, duplicate account detection, etc.). However, the
-main focus will be on mining experimental data from the site and using
-features derived from this data to build and evaluate classifiers to
-detect unwanted behavior programmatically. The large volume of data
-available to us will provide a unique perspective both on the types of
+frequency capping, duplicate account detection, etc.). We then analyze 
+message and user account data points to try to differentiate legitimate users
+from malicious ones. By mining this data, we extract features 
+to build and evaluate classifiers that can detect unwanted behavior 
+programmatically. The large volume of data available to us, although we do 
+not use all of it, provides a unique perspective both on the types of
 malicious content that exist on such sites as well as on the
-effectiveness of classifier/learning-based approaches to identifying
+effectiveness of classifier-based approaches to identifying
 these activities.
+
+We investigated several machine learning techniques to automatically 
+detect spam and scams in the private message data, and chose to use 
+train and evaluate Naive Bayes, Linear Regression, and 
+Support Vector Machine (SVM) classifiers. Our implementations used 
+a variety of tools, including Matlab, ScalaNLP, LIBSVM, Lucene and Spark, 
+an in-memory distributed computing framework designed for machine 
+learning and iterative computation.
 
@@
 
-We plan to make heavy use of our unrestricted access to the data of
-InterPals, which has over 1.3 million active members. This data
-includes a corpus of over a 100 million private messages and another 2 
-million messages that have been labeled as spam by users. Other data
-includes 40 million or so "wall" comments, 5 million photos, and 8
-million photo comments.
-
-\subsection{Techniques}
-
-Using this data, we plan to first identify the most prevalent types of 
-malicious activities on the site. We will investigate various machine learning 
-techniques to automatically detect these activities, including sampling, clustering and 
-classifiers. Sampling and clustering, in addition to the user-labeled and moderator-verified 
-spam corpus, will help us identify training data.
-
-Classifiers that we plan to explore 
-include:
-
-\begin{itemize}
-\item Naive Bayes
-\item Regression
-\item SVMs 
-\end{itemize}
-
-We then test the classifiers on new site data to evaluate their performance.
-
-One important part of this project will be to identify features of undesirable activity that are useful in classification. We plan to examine:
-
-\subsection{Frameworks}
-We implement the detection algorithms using Matlab, ScalaNLP, LibSVM, and Spark, an in-memory distributed computing framework designed for machine learning and iterative computation.
+We enjoyed unrestricted access to the data of InterPals.net, a SNS with 
+over 1.3 million active members. This data includes a corpus of over a 
+100 million private messages and another 2 million messages that have been 
+labeled as spam by users. Other data includes 40 million or so "wall" 
+comments, 5 million photos, and 8 million photo comments.
 
@@
 
-The data in our
+\subsection{Spam Set}
+
+\subsection{Ham Set}
+
+To simplify the labeling of ham messages, we made the assumption that messages remaining in the inboxes of active users after a period of several months would most likely not be spam.
+Consequently, we extracted a working set of 2 million ham messages
 
@@
 This categorizations are based only on observation. Analyzing the relevant terms in log comments entered by moderators when deleting spammer accounts provides the following breakdown:
-
-\subsection{Spam Set}
-
-\subsection{Ham Set}
-
-To simplify the labeling of ham messages, we made the assumption that messages remaining in the inboxes of active users after a period of several months would most likely not be spam.
-Consequently, we extracted a working set of 2 million ham messages
 
@@
 \subsection{Mixed Features}
+
 \textbf{Account Lifetime.}
@@
 % Modified from Reynold's HW1  write-up
-We use a na\"ive Bayes model, in which the documents are generated according to the following process.  
-First, a class \(c\) is chosen from a probability distribution \(\Pr(c)\).
-Then, a document from class \(c\) is generated.
-We tried a multinomial model and a Bernoulli model to model this process.
-\begin{itemize}
-\item \emph{Multinomial model.}  A document length \(m_d\) is chosen and \(m_d\) words \(w_{d,1}, \dotsc, w_{d,m_d}\) are chosen independently from a distribution \(\Pr(w|c)\).
-\item \emph{Bernoulli model.}  For each word \(w\) in the vocabulary, \(w\) is added to the document with probability \(\Pr(w|c)\).
-\end{itemize}
-The distributions \(\Pr(c)\) and \(\Pr(w|c)\) are parameters of the model.
-
+%We use a na\"ive Bayes model, in which the documents are generated according to the following process.  
+%First, a class \(c\) is chosen from a probability distribution \(\Pr(c)\).
+%Then, a document from class \(c\) is generated.
+%We tried a multinomial model and a Bernoulli model to model this process.
+%\begin{itemize}
+%\item \emph{Multinomial model.}  A document length \(m_d\) is chosen and \(m_d\) words \(w_{d,1}, \dotsc, w_{d,m_d}\) are chosen independently from a distribution \(\Pr(w|c)\).
+%\item \emph{Bernoulli model.}  For each word \(w\) in the vocabulary, \(w\) is added to the document with probability \(\Pr(w|c)\).
+%\end{itemize}
+%The distributions \(\Pr(c)\) and \(\Pr(w|c)\) are parameters of the model.
+%
+%
 \subsection{Linear Regression}
@@
 % Modified from Cliff & Antonio's HW2 write-up
-Linear regression is a method to model the relationship between an output variable, $y$, and explanatory variables, $X$.
+%Linear regression is a method to model the relationship between an output variable, $y$, and explanatory variables, $X$.
 The output variable is a linear sum of the explanatory variables multiplied by their corresponding coeffecients, represented by $\beta$.
@@
 This gives us the simple regression formula: $$y=X\beta$$
-
-In this project, we performed ordinary least squares regression.
+%
+%In this project, we performed ordinary least squares regression.
 By minimizing the sum of squared residuals, we can solve for the unkown parameter vector $\beta$ using a closed form solution.
@@
 We then evaluate our model $\beta$ on the withheld reviews in order to perform a 10-fold cross-validation.
-
-All of our linear regression code was implemented in MATLAB because of its ease-of-use and optimizations on matrix multiplications.
+%
+%All of our linear regression code was implemented in MATLAB because of its ease-of-use and optimizations on matrix multiplications.
 We parsed the reviews from the tokenized binary format into a bag-of-words representation stored as a sparse matrix.
@@
 This filtered the one million reviews to 492,130.
-
-The actual matrix multiplications were done using the sparse $X$ matrix along with our $y$ vector and the ridge parameter.
+%
+%The actual matrix multiplications were done using the sparse $X$ matrix along with our $y$ vector and the ridge parameter.
 The multiplication $X'X$ takes up too much memory if we use the entire dictionary, so we chose to limit the dictionary size.
@@
 Using our 10,000 word dictionary, we calculated the performance of our matrix multiply to be 0.123 Gflops/s by dividing the number of non-zero elements in our matrices by the time to compute the multiplication.
-
-Since we used ridge regression, we had the ability to tune the $\lambda$ parameter in order to get the best performance.
+%
+%
+%Since we used ridge regression, we had the ability to tune the $\lambda$ parameter in order to get the best performance.
 As shown in figure \ref{fig:lambda}, we observed an AUC improvement of about .5\% between the default value of 1 and a tuned parameter of 1500.
@@
 Using the default value of 1, our top words included obscure words including ``\_c'', ``\_j'', ``donovan'', and ``matlock.'' When we used the tuned $\lambda$ value of 1500, the words were much more representative, as shown in table \ref{tab:words}.
+%
 
@@
 %intro to svms
-SVMs
 
@@
 
-\begin{tabular}{c|c|c}
-    \multicolumn{3}{c}{Unscaled: Data set size, Accuracy (bag-of-words), Accuracy (expert features)}
-    Data size & Accuracy, bag-of-words & Accuracy, expert features \\
-    \hline
-    2k & 0.0 & 0.0 \\
-    10k & 0.0 & 0.0 \\
-    22k & 0.0 & 0.0 \\ % note we chose 22k because 11k had the recipient info
-    50k  & 0.0 & 0.0 \\
-    100k  & 0.0 & 0.0 \\
-\end{tabular}
-
-\begin{tabular}{c|c|c}
-    \multicolumn{3}{c}{Scaled: Data set size, Accuracy (bag-of-words), Accuracy (expert features)}
-    Data size & Accuracy, bag-of-words & Accuracy, expert features \\
-    \hline
-    2k & 0.0 & 0.0 \\
-    10k & 0.0 & 0.0 \\
-    22k & 0.0 & 0.0 \\ % note we chose 22k because 11k had the recipient info
-\end{tabular}
+%\begin{tabular}{c|c|c}
+%    \multicolumn{3}{c}{Unscaled: Data set size, Accuracy (bag-of-words), Accuracy (expert features)}
+%    Data size & Accuracy, bag-of-words & Accuracy, expert features \\
+%    \hline
+%    2k & 0.0 & 0.0 \\
+%    10k & 0.0 & 0.0 \\
+%    22k & 0.0 & 0.0 \\ % note we chose 22k because 11k had the recipient info
+%    50k  & 0.0 & 0.0 \\
+%    100k  & 0.0 & 0.0 \\
+%\end{tabular}
+%
+%
+%\begin{tabular}{c|c|c}
+%    \multicolumn{3}{c}{Scaled: Data set size, Accuracy (bag-of-words), Accuracy (expert features)}
+%    Data size & Accuracy, bag-of-words & Accuracy, expert features \\
+%    \hline
+%    2k & 0.0 & 0.0 \\
+%    10k & 0.0 & 0.0 \\
+%    22k & 0.0 & 0.0 \\ % note we chose 22k because 11k had the recipient info
+%\end{tabular}
+%
 
@@
 
-\begin{tabular}{c|c}
-    \multicolumn{2}{c}{Tokens most likely to be positive} \\
-    token & \(log(p_+ / p_-)\) \\
-    \hline
-    pos & 0.0 \\
-    pos2 & 0.1 \\
-\end{tabular}
-
-\begin{tabular}{c|c}
-    \multicolumn{2}{c}{Tokens most likely to be negative} \\
-    token & \(log(p_+ / p_-)\) \\
-    \hline
-    neg & 0.0 \\
-    neg2 & 0.1 \\
-\end{tabular}
+%\begin{tabular}{c|c}
+%    \multicolumn{2}{c}{Tokens most likely to be positive} \\
+%    token & \(log(p_+ / p_-)\) \\
+%    \hline
+%    pos & 0.0 \\
+%    pos2 & 0.1 \\
+%\end{tabular}
+%
+%\begin{tabular}{c|c}
+%    \multicolumn{2}{c}{Tokens most likely to be negative} \\
+%    token & \(log(p_+ / p_-)\) \\
+%    \hline
+%    neg & 0.0 \\
+%    neg2 & 0.1 \\
+%\end{tabular}
+%
 
@@
 We are currently in the process of extending the moderation interface to include more specific tagging capabilities
-- certain classifiers might work better on certain types of spam
-- clustering to see how well these classes fit the spam types on the site
+%- certain classifiers might work better on certain types of spam
+%- clustering to see how well these classes fit the spam types on the site
 

### 2012-05-08 docs/paper/paper.tex
Paper: data set section update

@@
 All moderator actions require them to annotate their decision with a brief log message.
+When deleting a user, moderators have the option of flagging the reported message (if there is one) as spam, or of flagging all of the user's outgoing private messages as spam.
 
@@
 
+At the beginning of this research project, there were just over two million messages that had been flagged by moderators as spam.
+We extracted the contents of exactly two million of these messages from the site's database and merged them with records of account information for the deleted users who had sent the messages.
+In addition to using SQL to extract much of this information, we wrote Perl scripts to clean the data, tokenize messages and prepare data for further processing.
+The methodology section (X) offers more details on precisely which items of data were collected and why.
+
 \subsection{Ham Set}
 
+Unlike spam messages, we did not have access to a comparable corpus of human-labeled ham messages.
 To simplify the labeling of ham messages, we made the assumption that messages remaining in the inboxes of active users after a period of several months would most likely not be spam.
-Consequently, we extracted a working set of 2 million ham messages
+Consequently, we extracted a working set of 2 million ham messages that were sent in the beginning of January, 2012 and still existed in the recipient's inbox.
+To reduce the possibility of collecting messages sent by uncaught spammers to inactive or dormant users, we selected messages only from users and to recipients who had logged in within the last two weeks.
+As in the case of spam, we collected two million messages with associated account data and merged them into text files for feature extraction.
 

### 2012-05-08 docs/paper/paper.tex
paper update

@@
 
+\subsection{Current Anti-Spam Measures}
+
 Currently, the site combats spam and other Terms of Service violations through volunteer moderators.
@@
 When deleting a user, moderators have the option of flagging the reported message (if there is one) as spam, or of flagging all of the user's outgoing private messages as spam.
+
+Other anti-spam measures include widespread use of CAPTCHAs across the site and frequency caps.
+Short-window frequency caps are in place for all users, limiting the number of messages that can be sent per short time interval (1, 5 and 10 minutes).
+New users are also subject to a per-day cap on the number of unique users with whom they are able to initiate contact.
+
+Users are able to filter messages by age, sex, continent and country. At this time,
 
@@
 
+In the course of collecting the data, we observed a number of different types of undesirable messages.
 Our ongoing observation of spam on the InterPals website provided direct intuition into the major classes of unwanted behavior that we classified as spam for this project.
@@
 \textbf{Advance Fee Fraud} We observed this scam to come in various flavors.  (inheritance, lottery,  etc)
-\textbf{Romance Scam} 
+\textbf{Romance Scams:} These 
 \textbf{Sexual Solicitation}
@@
 We first extracted a number of fields that we expected might improve classification.
+
 To choose a sample of relevant features, we computed statistics on the extracted features, both on the ham and spam corpora.
+These statistics were generated using SQL on a table created from the merged and cleaned data generated in the process described above.
+SQL makes it trivial to calculate means and standard errors, but medians are more resource-intensive, so we omitted them from our
 
@@
 
-\textbf{Recipient Age:} The age listed on the message recipient's profile.
-
-\textbf{Recipient Sex:} The sex listed on the message recipient's profile.
-
-\textbf{Recipient Country:} This was the ISO 3166-1 alpha-2 code of the recipient of the message.
+\textbf{Recipient Age/Sex/Country:} These fields were the age, sex and ISO 3166-1 alpha-2 code of the message 
+recipient, as listed on their profile. Unfortunately, due to data collection issues, this data was only available 
+for the most recent 11,000 spam messages. While the male-to-female ratio of the recipients was very close to that 
+of non-spam recipients (both were 47\% male to 53\% female), we found that the age of recipients was typically higher.
 
@@
 
-Due to data collection issues, recipient data (age, sex, country) was only available for the most recent 11,000 spam messages.
-
 \subsection{Sender Account Features}
 
-\textbf{Sender IP:} Spammers in general had a smaller number of unique IP addresses associated with them than ham users.
+\textbf{Sender IPs:} Spammers in general had a smaller number of unique IP addresses associated with them than ham users, as per figure \ref{fig:uniqip}.
+The mean for spammers was 11.82 with a standard deviation of 52.52, whereas ham users saw a mean of 175.51 and standard deviation of 322.19.
 This is likely due to their comparatively low length of time on the site and the fact that many legitimate users are behind NATs that yield a high number of dynamic IP addresses.
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/unique-ips.pdf}
+    \caption{Number of unique IP addresses associated with user account. Spam: mean 11.82, std. dev. 52.52. Ham: mean 175.51, std. dev. 322.19}
+    \label{fig:uniqip}
+\end{figure}
+
 \textbf{Sender Email:} We found that spammers were
 
-\textbf{Sender Age:}
+\textbf{Sender Age:} Spam user age
 
-\textbf{Sender Country:}
+\textbf{Sender Country:} The distribution of countries as stated by users on their profile and as revealed by their IP address differed remarkably between the spam and ham user groups.
+48\% of spam users claimed to be from the United States, with the next most popular countries being UK, Ghana, Senegal, Germany and Canada taking combined 25\%.
+In contrast, only 15\% ham users claimed to be from the United States, with Russia, Korea, UK, France and Germany combining for 24\%.
+
+figure \ref{fig:usaspam}
+figure \ref{fig:usahamfig:lambda}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/ips-usa-spam.pdf}
+    \caption{Distribution of IPs by country for spammers with profiles stating a USA location.}
+    \label{fig:usaspam}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/ips-usa-ham.pdf}
+    \caption{Distribution of IPs by country for ham users with profiles stating a USA location.}
+    \label{fig:usaham}
+\end{figure}
 
 \textbf{Sender Birthday:}
+
+We also noticed that Afghanistan and Albania featured prominently at the 11th and 15th most popular countries f......
 
@@
 
-\textbf{Sender Username:}
-
-\textbf{Sender Sex:}
+\textbf{Username, Name:} While we collected these fields, we did not end up processing or analyzing them further.
+That said, we saw considerable repetition in the usernames of scammers and it seems that analyzing the substrings in these fields could yield useful features.
 
@@
 
+Support Vector Machines are a classification method  \cite{boser}, \cite{cortesv95}
+
 % scalanlp svm
@@
 Given the complexity of choosing kernels and parameters, the package includes a Python script that performs cross-validation to suggest appropriate $\gamma$ and $C$ values for the recommended radial basis function (RBF) kernel.
+The RBF kernel is a real-valued function such that
+
+\begin{itemize}
+\item $K(\mathbf{x}_i, \mathbf{x}_j) = (\gamma\mathbf{x}_i^\intercal\mathbf{x}_j+r)^d, \gamma>0$
+\end{itemize}
 We ran this script on a subset of 2500 ham and 2500 spam messages to obtain $C = 8$ and $\gamma = 0.0078125$.
@@
 
+\subsection{Spam Grouping}
+
+As a part of this project, we performed some qualitative analysis on spam and differentiated between the major categories of undesirable messages found on this SNS. However, we
+
+Characteristics of spam messages vary widely depending on the type of spam
+we currently lump all spam into a single category; we plan to build on the classes we mentioned in section()
+We are currently in the process of extending the moderation interface to include more specific tagging capabilities
+%- certain classifiers might work better on certain types of spam
+%- clustering to see how well these classes fit the spam types on the site
+
+\subsection{Features}
+
 In this project, we chose only a small subset of ``expert" features to use.
@@
 
-Characteristics of spam messages vary widely depending on the type of spam
-we currently lump all spam into a single category; we plan to build on the classes we mentioned in section()
-We are currently in the process of extending the moderation interface to include more specific tagging capabilities
-%- certain classifiers might work better on certain types of spam
-%- clustering to see how well these classes fit the spam types on the site
+It is also anecdotally clear from moderator log messages that spammers re-use their profile photos and these photos are often of celebrities, porn stars or simply stock photos. Hashing these images
 
@@
 It is conceivable that certain spammers may have similar site usage patterns, in terms of HTTP request intervals,
+
+\subsection{Testing \& Evaluation}
+
+- test SVM and other classifiers on newer data
+-
 
@@
 figure \ref{fig:friends}
-figure \ref{fig:usaspam}
-figure \ref{fig:usahamfig:lambda}
 figure \ref{fig:roclin}
@@
 figure \ref{fig:sendsex}
-figure \ref{fig:uniqip}
+figure \ref{fig:roc-25}
 \begin{figure}[h]
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/ips-usa-spam.pdf}
-    \caption{Distribution of IPs by country for spammers with profiles stating a USA location.}
-    \label{fig:usaspam}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figures/ips-usa-ham.pdf}
-    \caption{Distribution of IPs by country for ham users with profiles stating a USA location.}
-    \label{fig:usaham}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
     \includegraphics[width=\linewidth]{figures/linear-ridge.pdf}
@@
     \label{fig:roclin}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/roc-25.pdf}
+    \caption{LIBSVM ROC curves for classifier using bag-of-words and bag-of-words with ``expert" features on 50k messages.}
+    \label{fig:roc-25}
 \end{figure}
@@
 
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figures/unique-ips.pdf}
-    \caption{Number of unique IP addresses associated with user account. Spam: mean 11.82, std. dev. 52.52. Ham: mean 175.51, std. dev. 322.19}
-    \label{fig:uniqip}
-\end{figure}
-
 \bibliographystyle{abbrv}

### 2012-05-08 docs/paper/paper.tex
added median

@@
 These statistics were generated using SQL on a table created from the merged and cleaned data generated in the process described above.
-SQL makes it trivial to calculate means and standard errors, but medians are more resource-intensive, so we omitted them from our
 
@@
 \textbf{Sender IPs:} Spammers in general had a smaller number of unique IP addresses associated with them than ham users, as per figure \ref{fig:uniqip}.
-The mean for spammers was 11.82 with a standard deviation of 52.52, whereas ham users saw a mean of 175.51 and standard deviation of 322.19.
+The mean for spammers was 11.82 with a standard deviation of 52.52 and a median of 2, whereas ham users saw a mean of 175.51, a standard deviation of 322.19 and median of 64.
 This is likely due to their comparatively low length of time on the site and the fact that many legitimate users are behind NATs that yield a high number of dynamic IP addresses.
@@
     \includegraphics[width=\linewidth]{figures/unique-ips.pdf}
-    \caption{Number of unique IP addresses associated with user account. Spam: mean 11.82, std. dev. 52.52. Ham: mean 175.51, std. dev. 322.19}
+    \caption{Number of unique IP addresses associated with user account. Spam: mean 11.82, std. dev. 52.52, median 2. Ham: mean 175.51, std. dev. 322.19, median 64}
     \label{fig:uniqip}
@@
 
+- plan to implement SVM, despite speed tradeoff
+
 \section{Future Work}

### 2012-05-08 docs/paper/paper.tex
Added table of feature statistics
update

@@
 \usepackage{xspace}
+\usepackage{multirow}
 
@@
 
-\title{Project Report: \\
-Detecting Spam on Social Networking Sites}
+\title{Detecting Spam on Social Networking Sites}
 
@@
 \maketitle
+\begin{abstract}
+Social networking sites (SNSs) see a variety of spam and scams targeted at 
+their users. In contrast to the more limited amounts of secondary information 
+available when analyzing email spam, spam on SNSs is accompanied by a wealth 
+of data on the sender, which can be used to build more accurate detection mechanisms. 
+We analyze public and private data from a popular social network in order to 
+gain insight into the various features spam messages and the accompanying user 
+accounts and profiles available to site operators. By examining the ....
+\end{abstract}
+
+\maketitle
 
@@
 
-\subsection{Problem}
 Social networks of any significant size see constant spam, scams
@@
 advertisements, fraudsters lure users with advance fee frauds and
-other confidence tricks, while others attempt to steal user
-information by directing users to external phishing pages. To make
-matters more difficult, a site with global reach sees communication
-among its members in a number foreign languages with varying levels
+other confidence tricks, while others may attempt to steal user
+information by directing users to external phishing pages. 
+Sites with global reach see communication among its members in a 
+number foreign languages with varying levels
 of ability. This means that much benign content shares characteristics
@@
 
+Email spam and techniques with which to combat it has been studied 
+extensively. Like
 This paper examines in detail the types of malicious and benign
@@
 detect spam and scams in the private message data, and chose to use 
-train and evaluate Naive Bayes, Linear Regression, and 
+train and evaluate Na\"ive Bayes, Linear Regression, and 
 Support Vector Machine (SVM) classifiers. Our implementations used 
@@
 
-\section{Background}
+\section{Related Work}
 
@@
 \cite{irani} show that static features associated with user signups on MySpace are enough to train an effective social spam classifier.
-They note that C4.5 decision tree algorithms provide better performance than naive Bayes in this case.
+They note that C4.5 decision tree algorithms provide better performance than Na\"ive Bayes in this case.
 As in other works, this only examines publicly available profile information collected by social honeypots.
@@
 Dalvi et al.
-\cite{dalvi} offer a modified Naive Bayes classifier to detect and reclassify data taking into account the optimal modification strategy that an adversary could choose.
+\cite{dalvi} offer a modified Na\"ive Bayes classifier to detect and reclassify data taking into account the optimal modification strategy that an adversary could choose.
 Lowd and Meek \cite{lowd} provide a framework for reverse engineering a classifier to determine whether an adversary can efficiently learn enough about a classifier to effectively defeat it.
@@
 
-At the beginning of this research project, there were just over two million messages that had been flagged by moderators as spam.
-We extracted the contents of exactly two million of these messages from the site's database and merged them with records of account information for the deleted users who had sent the messages.
-In addition to using SQL to extract much of this information, we wrote Perl scripts to clean the data, tokenize messages and prepare data for further processing.
+At the beginning of this research project, slightly two million messages had been flagged by moderators as spam.
+We extracted the contents of exactly two million spam-labeled messages from accounts deleted by moderators between October 2011 and March 2012.
+As moderators can flag a user's entire list of outgoing messages as spam on deletion, the earliest sent dates of some messages in our data set begin in May 2010.
+
+InterPals stores private messages and user account data in a number of separate MySQL tables.
+Account information for deleted users is stored for six months, so account details for all spam and ham accounts was available.
+Extracting the data of interest to this project required dumping the contents from a query with multiple joins and aggregations across 8 tables with a combined size of 154 GB.
+To minimize the impact on the load of the site's production database instances, we took an LVM snapshot of slave instance and cloned the database to a spare server.
+In addition to using SQL to extract the data, we wrote Perl scripts to clean the data, tokenize messages and prepare data for further processing.
 The methodology section (X) offers more details on precisely which items of data were collected and why.
@@
 
+\textbf{Sender Photos:} We found that the number of photos that user has uploaded is also indicative of their spam or ham reputation. Spammers
+
 \textbf{Sender Birthday:}
@@
 \textbf{Sender Name:}
+
+\begin{table}
+\begin{tabular}{|l|l|r|r|r|}
+    \hline
+    Feature & Class & Mean & Median & Std. Dev.\\
+    \hline
+    \multirow{2}{*}{Unique IPs} & Spam & 11.82 & 2 & 52.52\\
+    & Ham & 175.51 & 64 & 322.19\\
+    \hline
+    \multirow{2}{*}{Sender Age} & Spam & 33.52 & 31 & 11.09\\
+    & Ham & 24.86 & 22 & 10.30\\
+    \hline
+    \multirow{2}{*}{Photos} & Spam & 2.09 & 1 & 8.04\\
+    & Ham & 15.69 & 6 & 43.40\\
+    \hline
+    \multirow{2}{*}{Photo Albums} & Spam & 0.91 & 1 & 0.56\\
+    & Ham & 1.89 & 1 & 2.43\\
+    \hline
+    \multirow{2}{*}{Friends} & Spam & 11.82 & 0 & 52.52\\
+    & Ham & 23.26 & 10 & 58.91\\
+    \hline
+    \multirow{2}{*}{Recipient Age} & Spam & 39.98 & 41 & 12.69\\
+    & Ham & 25.02 & 22 & 10.70\\
+    \hline
+    \multirow{2}{*}{Birth Day} & Spam & 12.31 & 10 & 8.52\\
+    & Ham & 15.01 & 15 & 8.92\\
+    \hline
+    \multirow{2}{*}{Birth Month} & Spam & 5.85 & 5 & 3.33\\
+    & Ham & 6.38 & 6 & 3.48\\
+    \hline
+    \multirow{2}{*}{Description Length} & Spam & 267.30 & 87 & 483.19\\
+    & Ham & 439.33 & 267 & 613.26\\
+    \hline
+    \multirow{2}{*}{\% Fields Completed} & Spam & 22.15 & 14.00 & 30.75\\
+    & Ham & 79.29 & 86.00 & 32.85\\
+    \hline
+\end{tabular}
+\caption{Feature statistics}
+\label{featsat}
+\end{table}
 
@@
 
+\subsection{Features for Classification}
+In choosinnng
+
 \section{Classification}
 
-\subsection{Naive Bayes}
+Our goal in classification was not only to build an effective classifier for spam detection, but also to see how the presence of ``expert features", or features based on account and message metadata, could affect classifier accuracy.
+To this end, with of each of the classifiers we chose to implement, we evaluated performance on a bag-of-words representation of message data as well as a combination of bag-of-words with expert features.
+
+%feature extraction tools: perl, spark, lucene analyzers
+%- feature extraction pipeline is parallelized / distributed using Spark
+%- it produces a feature matrix used later for training classifiers
+%- Stage 1 generates a dictionary sorted by document frequencies in descending order
+%- Stage 2 create bag-of-words representation of the examples, as well as “expert-features” along with the documents.
+%
+%- two feature sets: one just with bag-of-words. the other with expert features chosen by a domain expert (site owner)
+%
+%classification:
+%- implemented 3 models: naive bayes, linear regression, and SVM.
+%- used scalanlp for SVM: Pegasos: Primal Estimated sub-GrAdient SOlver for SVM Extended with Wang, Crammer, Vucetic's work for Multiclass.
+The optimizer runs a stochastic subgradient descent on the primal objective using batches provided.
+%- used scalanlp for naive bayes
+%- used matlab for linear regression with ridge regularization.
+%- also attempted logistics regression with scalanlp, but training result was buggy (always output 0 probably. perhaps a bug in scalanlp)
+%
+%- naive bayes is super fast. accuracy around 80%. when expert features are added, accuracy dropped.
+%
+%- SVM is slow, and doesn’t produce very good result. Suspect we don’t have enough training data (22k training examples, 5k features)
+%
+%- linear regression performed the best, with an AUC of 0.xx. When expert features are added, AUC went up to 0.xx.
+%
+
+\subsection{Na\"ive Bayes}
 
@@
 
+\subsection{Na\"ive Bayes}
+
 \subsection{Linear Regression}
@@
 \section{Conclusion}
+
+In this paper, we presented a brief overview of the breakdown of spam and scams on the InterPals web site. 
+We then examined a number of statistics on both public and private data that showed that spam users, even when 
+grouped together without regard to their specific angle of attack, share a number of characteristics that differentiate 
+them from legitimate users. To simulate an approach 
+These features allowed us to improve the accuracy of classifiers
 

### 2012-05-08 docs/paper/paper.tex
update

@@
 \newcommand{\eg}{{\em e.g.,}~}
+
+\setcounter{page}{1}
+\pagenumbering{arabic}
 
@@
 
-\textbf{Sender IPs:} Spammers in general had a smaller number of unique IP addresses associated with them than ham users, as per figure \ref{fig:uniqip}.
-The mean for spammers was 11.82 with a standard deviation of 52.52 and a median of 2, whereas ham users saw a mean of 175.51, a standard deviation of 322.19 and median of 64.
-This is likely due to their comparatively low length of time on the site and the fact that many legitimate users are behind NATs that yield a high number of dynamic IP addresses.
-One possibility for future research is to investigate the ratio of unique IPs for users as a function of the time they have been registered.
-Likewise, examining unique /16 network blocks instead of unique IP addresses could mitigate the influence of NATs.
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figures/unique-ips.pdf}
-    \caption{Number of unique IP addresses associated with user account. Spam: mean 11.82, std. dev. 52.52, median 2. Ham: mean 175.51, std. dev. 322.19, median 64}
-    \label{fig:uniqip}
-\end{figure}
-
-\textbf{Sender Email:} We found that spammers were
-
-\textbf{Sender Age:} Spam user age
-
-\textbf{Sender Country:} The distribution of countries as stated by users on their profile and as revealed by their IP address differed remarkably between the spam and ham user groups.
-48\% of spam users claimed to be from the United States, with the next most popular countries being UK, Ghana, Senegal, Germany and Canada taking combined 25\%.
-In contrast, only 15\% ham users claimed to be from the United States, with Russia, Korea, UK, France and Germany combining for 24\%.
-
-figure \ref{fig:usaspam}
-figure \ref{fig:usahamfig:lambda}
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figures/ips-usa-spam.pdf}
-    \caption{Distribution of IPs by country for spammers with profiles stating a USA location.}
-    \label{fig:usaspam}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figures/ips-usa-ham.pdf}
-    \caption{Distribution of IPs by country for ham users with profiles stating a USA location.}
-    \label{fig:usaham}
-\end{figure}
-
-\textbf{Sender Photos:} We found that the number of photos that user has uploaded is also indicative of their spam or ham reputation. Spammers
-
-\textbf{Sender Birthday:}
-
-We also noticed that Afghanistan and Albania featured prominently at the 11th and 15th most popular countries f......
-
-\textbf{Sender Age:}
-
-\textbf{Sender Name:}
-
+We chose message and user account features to analyze primarily by the amount of data that we had for each them, while making sure to include the most vital account information.
+Table \ref{tab:feats} provides averages on a number of the account statistics that we examined.
 \begin{table}
@@
     \hline
-    Feature & Class & Mean & Median & Std. Dev.\\
+    \textbf{Feature} & \textbf{Label} & \textbf{Mean} & \textbf{Median} & \textbf{Std. Dev.}\\
     \hline
@@
     \hline
-    \multirow{2}{*}{Description Length} & Spam & 267.30 & 87 & 483.19\\
+    \multirow{2}{*}{Profile Desc. Len.} & Spam & 267.30 & 87 & 483.19\\
     & Ham & 439.33 & 267 & 613.26\\
@@
 \caption{Feature statistics}
-\label{featsat}
+\label{tab:feats}
 \end{table}
+
+\textbf{Sender IPs:} Spammers in general had a smaller number of unique IP addresses associated with them than ham users, as per figure \ref{fig:uniqip}.
+The mean for spammers was 11.82 with a standard deviation of 52.52 and a median of 2, whereas ham users saw a mean of 175.51, a standard deviation of 322.19 and median of 64.
+This is likely due to their comparatively low length of time on the site and the fact that many legitimate users are behind NATs that yield a high number of dynamic IP addresses.
+One possibility for future research is to investigate the ratio of unique IPs for users as a function of the time they have been registered.
+Likewise, examining unique /16 network blocks instead of unique IP addresses could mitigate the influence of NATs.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/unique-ips.pdf}
+    \caption{Number of unique IP addresses associated with user account. Spam: mean 11.82, std. dev. 52.52, median 2. Ham: mean 175.51, std. dev. 322.19, median 64}
+    \label{fig:uniqip}
+\end{figure}
+
+\textbf{Sender Email:} Users must provide email addresses upon signup. The site does not allow users to contact other 
+members before verifying that the email address provided is valid, via an account activation link sent to the signup 
+address. Subsequent changes to a user's email address on file require a similar confirmation process. Thus, we can be certain 
+that the email accounts that we extracted were in use by the senders, at least at the time of registration.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/email.pdf}
+    \caption{Domain name of email provided at signup.}
+    \label{fig:email}
+\end{figure}
+
+In the data processing step, we extracted the domain name from each email address. We found that ham user email domains 
+were distributed across top email providers in similar proportions (Hotmail, Yahoo and Gmail accounted for 22\%, 17\% and 17\% 
+respectively). For spam user accounts, Hotmail, Yahoo and Gmail accounted for 7\%, 72\% and 6\% respectively. This striking 
+predominance of Yahoo email accounts among spammers can be seen in figure \ref{fig:email}. While we have no explanation for 
+this, figure \cite{fig:ysmap} shows that Ghana and Nigeria alone account for 38\% of the IP addresses associated with these 
+Yahoo accounts. This popularity of Yahoo among West African users has been noted before\cite{burrell}, and appears to be 
+due to the early penetration of Yahoo services in the region.
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/yahoo-spam.pdf}
+    \caption{IP distribution for scam accounts with Yahoo email addresses.}
+    \label{fig:yspam}
+\end{figure}
+
+\textbf{Sender Age:} Spam user age tended to be higher
+
+\textbf{Sender Birthday:} We analyzed
+
+We also noticed that Afghanistan and Albania featured prominently at the 11th and 15th most popular countries f......
+
+figure \ref{fig:day}
+figure \ref{fig:month}
+figure \ref{fig:drop}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/dob-day.pdf}
+    \caption{Date of month from birthday submitted by user at signup.}
+    \label{fig:day}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/dob-month.pdf}
+    \caption{Month of birthday as submitted by user at signup.}
+    \label{fig:month}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/dropdown.png}
+    \caption{Drop-down select menu on signup form.}
+    \label{fig:drop}
+\end{figure}
+
+\textbf{Sender Country:} The distribution of countries as stated by users on their profile and as revealed by their IP address differed remarkably between the spam and ham user groups.
+48\% of spam users claimed to be from the United States, with the next most popular countries being UK, Ghana, Senegal, Germany and Canada taking combined 25\%.
+In contrast, only 15\% ham users claimed to be from the United States, with Russia, Korea, UK, France and Germany combining for 24\%.
+
+figure \ref{fig:usaspam}
+figure \ref{fig:usahamfig:lambda}
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/ips-usa-spam.pdf}
+    \caption{Distribution of IPs by country for spammers with profiles stating a USA location.}
+    \label{fig:usaspam}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/ips-usa-ham.pdf}
+    \caption{Distribution of IPs by country for ham users with profiles stating a USA location.}
+    \label{fig:usaham}
+\end{figure}
+
+\textbf{Sender Photos:} We found that the number of photos that user has uploaded is also indicative of their spam or ham reputation. Spammers
+
+\textbf{Sender Age:}
+
+\textbf{Sender Name:}
 
@@
 
-\subsection{Features for Classification}
-In choosinnng
-
 \section{Classification}
@@
 
-%feature extraction tools: perl, spark, lucene analyzers
-%- feature extraction pipeline is parallelized / distributed using Spark
-%- it produces a feature matrix used later for training classifiers
-%- Stage 1 generates a dictionary sorted by document frequencies in descending order
-%- Stage 2 create bag-of-words representation of the examples, as well as “expert-features” along with the documents.
-%
-%- two feature sets: one just with bag-of-words. the other with expert features chosen by a domain expert (site owner)
-%
+In order to train our classifiers, we had to create a feature matrix from the message and account data. We did this using 
+Perl, Spark, and Lucene analyzers. The first step was to generate a dictionary sorted by document frequencies in descending 
+order, mapping each feature (word or expert feature value) to an integer key. The second step was to create a sparse 
+matrix representation of each document (in this case, spam or ham message) and the associated features. Spark provides a 
+parallelized in-memory computation framework that dramatically reduced the time necessary to extract these features, 
+build the dictionary and generate the feature matrix.
+
+We did not use all of the expert features available to us, due to time constraints. Likewise, for many expert features could 
+have a wide range of values, which would expand the dictionary size considerably. In addition to the bag-of-words from the 
+message text, we used the following features:
+\begin{itemize}
+\item Sex of sender
+\item Age of sender
+\item Age of recipient
+\item Account lifetime
+\item Month of birth 
+\item Whether sender has friends (boolean)
+\item Whether sender has photo(s) (boolean)
+\item Profile country and IP-detected country match (boolean)
+\item IP-detected country is Ghana, Nigeria, Senegal, Malaysia, Turkey, Gambia, Ivory Coast, Togo (``high-risk" countries) (boolean)
+\end{itemize}
+
 %classification:
@@
 figure \ref{fig:cip}
-figure \ref{fig:day}
-figure \ref{fig:month}
 figure \ref{fig:drop}
-figure \ref{fig:email}
 figure \ref{fig:friends}
@@
     \label{fig:cip}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figures/dob-day.pdf}
-    \caption{Date of month from birthday submitted by user at signup.}
-    \label{fig:day}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figures/dob-month.pdf}
-    \caption{Month of birthday as submitted by user at signup.}
-    \label{fig:month}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figures/dropdown.png}
-    \caption{Drop-down select menu on signup form.}
-    \label{fig:drop}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figures/email.pdf}
-    \caption{Domain name of email provided at signup. Users cannot send messages before 
-        activating their account via a link sent to their email address.}
-    \label{fig:email}
 \end{figure}

### 2012-05-08 docs/paper/paper.tex
update

@@
 
-We decided to focus on spam in private messages for this paper, given that messages account for a majority of user spam reports.
+We chose to focus on spam in private messages for this paper, given that messages account for a majority of user spam reports.
 In addition to using the bag-of-words representation of the message content, we aimed to identify a subset of additional, relevant ``expert" features based on public and non-public message and account information that would augment the classifier's accuracy.
@@
 
-\textbf{Message Body \& Subject:} We extracted both the message
-
-\textbf{Sender IP:} The IP address of the sender at the time that the user sent the message.
+\textbf{Message Body \& Subject:} We extracted both the message body and subject from 
+each spam and ham message. The cleaning script generated two additional features based 
+on this. First, we counted the frequency of non-standard punctuation (we noticed that many 
+spam messages would, for example, put spaces before commas, periods and quotation marks, 
+and omit spaces after these characters). Second, we calculated the ratio of uppercase to 
+lowercase letters in the text, after observing relatively high amounts of uppercase text in spam messages.
+%% add numbers
 
@@
 
-\textbf{Recipient Replied:} A boolean, 1 or 0, depending on whether the recipient replied to the message.
+\textbf{Recipient Replied:} A boolean, 1 or 0, depending on whether the recipient replied to the message. We saw 
+that the mean reply rate for ham messages was 81.77\% (SD: 38.61), while it was only 0.81\% (SD: 8.97) for spam messages.
 
@@
 \end{table}
+
+\textbf{Sender Country:} The distribution of countries as stated by users on their profile and as 
+revealed by their IP address differed remarkably between the spam and ham user groups. 
+48\% of spam users claimed to be from the United States, with the next most popular 
+countries being the UK, Ghana, Senegal, Germany and Canada taking combined 25\%. 
+In contrast, only 15\% ham users claimed to be from the United States, with Russia, 
+Korea, UK, France and Germany combining for 24\%. This distribution of profile countries can 
+be seen in Figure \ref{fig:cprof}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/country-prof.pdf}
+    \caption{Country as stated on user profile.}
+    \label{fig:cprof}
+\end{figure}
+
+Users are free to choose any country on their profile. We record each IP address from which a 
+user logs in to the site. By examining the distribution of countries (using the MaxMind GeoIP 
+database) associated with the unique IP addresses (Figure \ref{fig:cip}), we see that the top 
+countries for spammers are significantly different, while the top ham countries are 
+virtually unchanged. Furthermore, only 30\% of the IP addresses of spam users who 
+claimed to be from the United States actually mapped back to a US-based ISP, with a combined 
+46\% indicating a Ghana, Nigeria and Senegal ISP (Figure \ref{fig:usaspam}). This contrasts 
+with the 73\% that US-based IP addresses comprised for ham users whose profiles stated that 
+they were in the United States (Figure \ref{fig:usaham}).
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/country-ip.pdf}
+    \caption{Country detected by IP address using MaxMind GeoIP database.}
+    \label{fig:cip}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/ips-usa-spam.pdf}
+    \caption{Distribution of IPs by country for spammers with profiles stating a USA location.}
+    \label{fig:usaspam}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/ips-usa-ham.pdf}
+    \caption{Distribution of IPs by country for ham users with profiles stating a USA location.}
+    \label{fig:usaham}
+\end{figure}
 
@@
     \label{fig:drop}
-\end{figure}
-
-\textbf{Sender Country:} The distribution of countries as stated by users on their profile and as revealed by their IP address differed remarkably between the spam and ham user groups.
-48\% of spam users claimed to be from the United States, with the next most popular countries being UK, Ghana, Senegal, Germany and Canada taking combined 25\%.
-In contrast, only 15\% ham users claimed to be from the United States, with Russia, Korea, UK, France and Germany combining for 24\%.
-
-figure \ref{fig:usaspam}
-figure \ref{fig:usahamfig:lambda}
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figures/ips-usa-spam.pdf}
-    \caption{Distribution of IPs by country for spammers with profiles stating a USA location.}
-    \label{fig:usaspam}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figures/ips-usa-ham.pdf}
-    \caption{Distribution of IPs by country for ham users with profiles stating a USA location.}
-    \label{fig:usaham}
 \end{figure}
@@
 figure \ref{fig:roc-25}
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figures/country-prof.pdf}
-    \caption{Country as stated on user profile.}
-    \label{fig:cprof}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figures/country-ip.pdf}
-    \caption{Country detected by IP address using MaxMind GeoIP database.}
-    \label{fig:cip}
-\end{figure}
 

### 2012-05-08 docs/paper/paper.tex
update

@@
 Rank-based approaches demote visibility of questionable content, while interface-based approaches apply policies to prevent unwanted behavior.
-This work groups classification-based approaches with detection, although classifiers can be used in conjuction with user information to prevent spam before it happens.
+This work groups classification-based approaches with detection, although classifiers can be used in conjunction with user information to prevent spam before it happens.
 
@@
 This multi-pronged approach lends itself to associative classification, in which, for example, a message would be classified as spam if it contained a link to a web page that had a high probability of being spam.
-Unfortunately, the differing characterestics of various social networks \eg the length of messages in Facebook vs.
+Unfortunately, the differing characteristics of various social networks \eg the length of messages in Facebook vs.
 Twitter, can reduce the benefits of sharing spam corpora across diverse sites.
@@
 The system is composed of classifier services, an ML-derived Feature Extraction Language (FXL), feature loops to aggregate and prepare features for classification and a policy engine to take action on suspected misuse.
-While the discussion remains high-level and includes few implemention particulars, it does include significant detail on the various types and characterists of undesirable activity on the site, including fake profiles, harassment, compromised accounts, malware and spam.
+While the discussion remains high-level and includes few implemention particulars, it does include significant detail on the various types and characteristics of undesirable activity on the site, including fake profiles, harassment, compromised accounts, malware and spam.
 
@@
 The framework uses a Bayesian classifier and links messages with similar content, but does not take into account other features.
-Nevertheless, this is one of the few studies to test its framework on non-public data, inlcuding private messages, spam reports and user profiles from a large Dutch social networking site.
+Nevertheless, this is one of the few studies to test its framework on non-public data, including private messages, spam reports and user profiles from a large Dutch social networking site.
 
@@
 and upload photographs.
-After clicking on a link sent to their email addres, a user can begin to interact with others on the site via private message, public ``wall" posts, comments on photos and on a bulletin board system.
+After clicking on a link sent to their email address, a user can begin to interact with others on the site via private message, public ``wall" posts, comments on photos and on a bulletin board system.
 
@@
 
-\textbf{Message Body \& Subject:} We extracted both the message body and subject from 
+\textbf{Message body \& subject:} We extracted both the message body and subject from 
 each spam and ham message. The cleaning script generated two additional features based 
@@
 
-\textbf{Recipient Age/Sex/Country:} These fields were the age, sex and ISO 3166-1 alpha-2 code of the message 
+\textbf{Recipient age, sex and country:} These fields were the age, sex and ISO 3166-1 alpha-2 code of the message 
 recipient, as listed on their profile. Unfortunately, due to data collection issues, this data was only available 
@@
 
-\textbf{Recipient Replied:} A boolean, 1 or 0, depending on whether the recipient replied to the message. We saw 
+\textbf{Recipient replied:} This boolean value indicates whether the recipient replied to the message. We saw 
 that the mean reply rate for ham messages was 81.77\% (SD: 38.61), while it was only 0.81\% (SD: 8.97) for spam messages.
@@
     \includegraphics[width=\linewidth]{figures/unique-ips.pdf}
-    \caption{Number of unique IP addresses associated with user account. Spam: mean 11.82, std. dev. 52.52, median 2. Ham: mean 175.51, std. dev. 322.19, median 64}
+    \caption{Number of unique IP addresses associated with user account.}
     \label{fig:uniqip}
@@
 
-% Modified from Cliff & Antonio's HW2 write-up
-%Linear regression is a method to model the relationship between an output variable, $y$, and explanatory variables, $X$.
-The output variable is a linear sum of the explanatory variables multiplied by their corresponding coeffecients, represented by $\beta$.
-In our case, $y$ represents a review's ranking from one to five, while $X$ is the bag-of-words representation of the review text.
-This implies that $\beta$ represents the weight of a word's correlation to a higher review.
-This gives us the simple regression formula: $$y=X\beta$$
-%
-%In this project, we performed ordinary least squares regression.
-By minimizing the sum of squared residuals, we can solve for the unkown parameter vector $\beta$ using a closed form solution.
+Linear regression is a method to model the relationship between an output variable, 
+$y$, and explanatory variables, $X$. The output variable is a linear sum of the explanatory 
+variables multiplied by their corresponding coeffecients, represented by $\beta$. 
+In our case, $y$ represents a message's label as spam or ham (1 or 0), 
+while $X$ is the feature vector, a bag-of-words representation of the message text and expert features. 
+This implies that $\beta$ represents the weight of a feature's correlation to whether 
+the message is spam or not. This gives us the simple regression formula: $$y=X\beta$$
+
+In this project, we performed ordinary least squares regression. By minimizing the sum of squared 
+residuals, we can solve for the unkown parameter vector $\beta$ using a closed form solution. 
 $$\beta=(X'X)^{-1}X'y$$ In order to ensure that the matrix $X'X$ is invertible, we perform ridge regression. 
-In ridge regression, we add the identity matrix, $I$, scaled by a factor, $\lambda$, to produce the following equation.
-$$\beta=(X'X+\lambda I)^{-1}X'y$$ We calculate this $\beta$ for a training set of review text along with their corresponding ratings which range from 1 to 5.
-We then evaluate our model $\beta$ on the withheld reviews in order to perform a 10-fold cross-validation.
-%
-%All of our linear regression code was implemented in MATLAB because of its ease-of-use and optimizations on matrix multiplications.
-We parsed the reviews from the tokenized binary format into a bag-of-words representation stored as a sparse matrix.
+In ridge regression, we add the identity matrix, $I$, scaled by a factor, $\lambda$, to 
+produce the following equation. $$\beta=(X'X+\lambda I)^{-1}X'y$$
+
+We calculate this $\beta$ for a training set of bag-of-words features along with their corresponding labels. 
+We then evaluate our model $\beta$ on the remaining messages in order to perform 10-fold cross-validation. 
+We repeat this process with the set that includes expert features.
+
+All of our linear regression code was implemented in MATLAB because of its ease-of-use 
+and optimizations on matrix multiplications. We read in the text-based feature dictionaries and feature matrix
+representations created above,
+
+%We parse the messages from the text format produced by our cleaning script, using the dictionary generated earlier map words to integers, saving the feature data into two sparse matrices.
+The first matrix only includes the bag-of-words representation of the message body, while the second includes the expert features as well.
 Matlab uses Compressed Sparse Column format, so we actually stored the matrix $X'$ because the dictionary is very sparse.
-We also filtered out duplicate reviews by hashing the first ten words in a review to determine distinctness.
-This filtered the one million reviews to 492,130.
-%
-%The actual matrix multiplications were done using the sparse $X$ matrix along with our $y$ vector and the ridge parameter.
+
+%For each feature set, the actual matrix multiplications were done using the sparse $X$ matrix along with our $y$ vector and the ridge parameter.
 The multiplication $X'X$ takes up too much memory if we use the entire dictionary, so we chose to limit the dictionary size.
@@
 Given the complexity of choosing kernels and parameters, the package includes a Python script that performs cross-validation to suggest appropriate $\gamma$ and $C$ values for the recommended radial basis function (RBF) kernel.
-The RBF kernel is a real-valued function such that
-
-\begin{itemize}
-\item $K(\mathbf{x}_i, \mathbf{x}_j) = (\gamma\mathbf{x}_i^\intercal\mathbf{x}_j+r)^d, \gamma>0$
-\end{itemize}
+The RBF kernel is a real-valued function such that: $$K(\mathbf{x}_i, \mathbf{x}_j) = (\gamma\mathbf{x}_i^\intercal\mathbf{x}_j+r)^d, \gamma>0$$
 We ran this script on a subset of 2500 ham and 2500 spam messages to obtain $C = 8$ and $\gamma = 0.0078125$.
@@
 \subsection{Linear Regression}
+
+We tried a number of values for $\lambda$: 1, 10, 50, 100, 500, 1000, 1500, 3000. Figure \ref{fig:lambda}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/linear-ridge.pdf}
+    \caption{AUC based on varying lambda values for Ridge regression.}
+    \label{fig:lambda}
+\end{figure}
+
+\begin{tabular}{c|c
 
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/linear-ridge.pdf}
-    \caption{AUC based on varying lambda values for Ridge regression.}
-    \label{fig:lambda}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
     \includegraphics[width=\linewidth]{figures/linear-roc.pdf}

### 2012-05-08 docs/paper/paper.tex
update

@@
 
-We chose to focus on spam in private messages for this paper, given that messages account for a majority of user spam reports.
-In addition to using the bag-of-words representation of the message content, we aimed to identify a subset of additional, relevant ``expert" features based on public and non-public message and account information that would augment the classifier's accuracy.
-We first extracted a number of fields that we expected might improve classification.
+We chose to focus on spam in private messages for this paper, given that messages account for 
+a majority of user spam reports. In addition to using the bag-of-words representation of 
+the message content, we aimed to identify a subset of additional, relevant ``expert" features 
+based on public and non-public message and account information that would augment the classifier's 
+accuracy. We first extracted a number of fields that we expected might improve classification.
 
-To choose a sample of relevant features, we computed statistics on the extracted features, both on the ham and spam corpora.
-These statistics were generated using SQL on a table created from the merged and cleaned data generated in the process described above.
+To choose a sample of relevant features, we computed statistics and generated histograms on 
+the extracted features, comparing the ham and spam corpora. These statistics were generated 
+using SQL on a table created from the merged and cleaned data generated in the process described above.
 
@@
 
-\textbf{Sender Age:} Spam user age tended to be higher
+\textbf{Sender \& Recipient Age:} The age indicated on their profile tended be higher for spam user than 
+for ham users, with median ages of 31 and 22 respectively (Figure \ref{fig:sendage}). Likewise, the 
+age of message recipients was higher for spam recipients that for recipients of ham messages, with 
+median ages of 41 and 22 (Figure \ref{fig:recipage}). The sample size for spam recipient age 
+was only 11,000 messages due to late data collection. We posit that the higher ages indicated on spam profiles, 
+mirrored in recipient ages, reflects targeting of older users who are more likely to be financially 
+stable.
 
-\textbf{Sender Birthday:} We analyzed
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/recip-age.pdf}
+    \caption{Age of message recipient as stated on user profile. Note that 2 million message spam sample 
+        had only 11,000 messages with this information.}
+    \label{fig:recipage}
+\end{figure}
 
-We also noticed that Afghanistan and Albania featured prominently at the 11th and 15th most popular countries f......
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/sender-age.pdf}
+    \caption{Age of message sender as stated on user profile. }
+    \label{fig:sendage}
+\end{figure}
+
+\textbf{Sender Birthday:} We analyzed the birthdays and birth months of ham and spam users and noticed that spam users 
+were more likely to have birthdays early in the month, and birth months early in the year. This is illustrated in 
+Figures \ref{fig:day} and \ref{fig:month}. Given the drop-down select menus on the web site's registration form, reproduced 
+Figure \ref{cite:drop}, it seems likely that this is due to an unwillingness on the part of spam users to 
+scroll down to lower options (as well as, perhaps, a hesitance to disclose real birthdates). A similar trend is 
+observed in the countries with the most discrepancies between IP-detected countries and countries stated on spam 
+user profiles, with Afghanistan and Albania comprising 19\%, despite their relatively low representation on the 
+site.
 
@@
 
-\textbf{Sender Photos:} We found that the number of photos that user has uploaded is also indicative of their spam or ham reputation. Spammers
+\textbf{Extra Profile Features:} We found that the number of photos that user has uploaded is also indicative of their spam or ham reputation. Spammers
 
@@
 
-\begin{tabular}{c|c
-
 %\begin{tabular}{c|c}
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/recip-age.pdf}
-    \caption{Age of message recipient as stated on user profile. Note that 2 million message spam sample 
-        had only 11,000 messages with this information.}
-    \label{fig:recipage}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figures/sender-age.pdf}
-    \caption{Age of message sender as stated on user profile. }
-    \label{fig:sendage}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
     \includegraphics[width=\linewidth]{figures/sex.pdf}

### 2012-05-08 docs/paper/paper.tex
update

@@
 
-%\begin{tabular}{c|c|c}
-%    \multicolumn{3}{c}{Unscaled: Data set size, Accuracy (bag-of-words), Accuracy (expert features)}
-%    Data size & Accuracy, bag-of-words & Accuracy, expert features \\
-%    \hline
-%    2k & 0.0 & 0.0 \\
-%    10k & 0.0 & 0.0 \\
-%    22k & 0.0 & 0.0 \\ % note we chose 22k because 11k had the recipient info
-%    50k  & 0.0 & 0.0 \\
-%    100k  & 0.0 & 0.0 \\
-%\end{tabular}
-%
-%
-%\begin{tabular}{c|c|c}
-%    \multicolumn{3}{c}{Scaled: Data set size, Accuracy (bag-of-words), Accuracy (expert features)}
-%    Data size & Accuracy, bag-of-words & Accuracy, expert features \\
-%    \hline
-%    2k & 0.0 & 0.0 \\
-%    10k & 0.0 & 0.0 \\
-%    22k & 0.0 & 0.0 \\ % note we chose 22k because 11k had the recipient info
-%\end{tabular}
-%
-
 \section{Results}
@@
 
-%\begin{tabular}{c|c}
-%    \multicolumn{2}{c}{Tokens most likely to be positive} \\
-%    token & \(log(p_+ / p_-)\) \\
-%    \hline
-%    pos & 0.0 \\
-%    pos2 & 0.1 \\
-%\end{tabular}
-%
-%\begin{tabular}{c|c}
-%    \multicolumn{2}{c}{Tokens most likely to be negative} \\
-%    token & \(log(p_+ / p_-)\) \\
-%    \hline
-%    neg & 0.0 \\
-%    neg2 & 0.1 \\
-%\end{tabular}
-%
+\begin{table}
+\begin{tabular}{l|l|l|l}
+\multicolumn{2}{c}{\textbf{Ham}} & 
+\multicolumn{2}{c}{\textbf{Spam}} \\
+\hline
+\textbf{Feature} & \textbf{Weight} & \textbf{Feature} & \textbf{Weight} \\
+\hline
+happy & -0.1314 & sex & 0.2295\\
+hahaha & -0.13 & cafe4tune & 0.1906\\
+turkey & -0.1271 & tony & 0.1715\\
+learn & -0.124 & oky & 0.1549\\
+fb & -0.1186 & photo & 0.1467\\
+goin & -0.1177 & hehe & 0.1379\\
+kk & -0.1164 & rochelle & 0.1346\\
+uganda & -0.1159 & register & 0.1327\\
+new & -0.1156 & irene & 0.1298\\
+msn & -0.1114 & okay & 0.1294\\
+ya & -0.11 & sexy & 0.1278\\
+thx & -0.1095 & ali & 0.1199\\
+question & -0.1095 & displaying & 0.1112\\
+xxx & -0.1081 & ghana & 0.111\\
+joyce & -0.1078 & correctly & 0.1102\\
+year & -0.1067 & ok & 0.1093\\
+hw & -0.1067 & m? & 0.1087\\
+snail & -0.1059 & sand & 0.1072\\
+direct & -0.1052 & check & 0.1058\\
+bonne & -0.1014 & sabina & 0.105\\
+\end{tabular}
+\caption{Linear Regression top features (bag-of-words) with weights}
+\label{tab:lrfeats}
+\end{table}
+
+\begin{table}
+\begin{tabular}{l|l|l|l}
+\multicolumn{2}{c}{\textbf{Ham}} & 
+\multicolumn{2}{c}{\textbf{Spam}} \\
+\hline
+\textbf{Feature} & \textbf{Weight} & \textbf{Feature} & \textbf{Weight} \\
+\hline
+
+requests & -0.1349 & F-country\_match & 0.3365\\
+giggle & -0.1257 & sex & 0.1954\\
+turkey & -0.1213 & cafe4tune & 0.1723\\
+hw & -0.1148 & register & 0.1324\\
+F-photo\_exists & -0.1087 & F-spam\_country & 0.1231\\
+xxx & -0.1047 & rochelle & 0.1054\\
+goin & -0.1008 & m? & 0.1019\\
+teacher & -0.1006 & tony & 0.1017\\
+happy & -0.0912 & hehe & 0.0991\\
+yie & -0.0904 & oky & 0.0968\\
+tea & -0.089 & sexy & 0.0967\\
+snail & -0.0852 & ewelina & 0.0874\\
+malaysia & -0.084 & card & 0.0866\\
+russian & -0.0832 & return & 0.0856\\
+fall & -0.0805 & video & 0.0842\\
+question & -0.0789 & pay & 0.0841\\
+ann? & -0.0785 & nas? & 0.0839\\
+learn & -0.0784 & model & 0.0826\\
+new & -0.0783 & displaying & 0.0808\\
+daichi & -0.0775 & k?  & 0.0806\\
+
+\end{tabular}
+\caption{Linear Regression top features (bag-of-words + expert features) with weights. Expert features
+are denoted by the 'F-' prefix.}
+\label{tab:lrfeats}
+\end{table}
 
 \subsection{SVMs}
-
-%\begin{tabular}{c|c}
-%    \multicolumn{2}{c}{Tokens most likely to be positive} \\
-%    token & \(log(p_+ / p_-)\) \\
-%    \hline
-%    pos & 0.0 \\
-%    pos2 & 0.1 \\
-%\end{tabular}
-%
-%\begin{tabular}{c|c}
-%    \multicolumn{2}{c}{Tokens most likely to be negative} \\
-%    token & \(log(p_+ / p_-)\) \\
-%    \hline
-%    neg & 0.0 \\
-%    neg2 & 0.1 \\
-%\end{tabular}
-%
 

### 2012-05-08 docs/paper/paper.tex
update

@@
 
+% LR B-O-W
 \begin{table}
-\begin{tabular}{l|l|l|l}
+\begin{tabular}{l|l||l|l}
 \multicolumn{2}{c}{\textbf{Ham}} & 
@@
 
+% LR + expert
 \begin{table}
-\begin{tabular}{l|l|l|l}
+\begin{tabular}{l|l||l|l}
 \multicolumn{2}{c}{\textbf{Ham}} & 
@@
 \subsection{SVMs}
+
+\begin{table}
+\begin{tabular}{l|l||l|l}
+\multicolumn{2}{c}{\textbf{Bag-of-Words}} & 
+\multicolumn{2}{c}{\textbf{Expert Features}} \\
+\hline
+\textbf{Feature} & \textbf{Ham Weight} & \textbf{Feature} & \textbf{Ham Weight} \\
+\hline
+
+year & 0.2253 & F-account\_lifetime & -0.98\\
+new & 0.225 & F-friends & -0.3671\\
+am & -0.1722 & F-recip\_age & -0.2129\\
+happy & 0.16 & F-age & -0.183\\
+d & 0.1513 & i & -0.05231\\
+ok & -0.1497 & you & -0.0483\\
+yahoo & -0.136 & to & -0.04026\\
+it & 0.12 & and & -0.03636\\
+me & -0.118 & F-month\_of\_birth & -0.03452\\
+too & 0.1173 & F-country\_match & -0.0269\\
+i'm & 0.1116 & me & -0.0224\\
+haha & 0.0998 & am & -0.022\\
+can & -0.0971 & my & -0.0209\\
+okay & -0.0947 & ? & 0.0199\\
+dear & -0.0882 & ? & 0.0191\\
+p & 0.0862 & a & -0.01573\\
+here & -0.0844 & your & -0.01554\\
+2012 & 0.084 & F-photo\_exists & -0.0146\\
+wish & 0.0836 & the & -0.01412\\
+doing & -0.0779 & f-sex & -0.01258\\
+
+\end{tabular}
+\caption{SVM term weights (ScalaNLP Pegasos implementation). The spam term weights were the same, but opposite signs.
+Question marks indicate strings in a different character encoding.}
+\label{tab:svmfeats}
+\end{table}
 

### 2012-05-08 docs/paper/paper.tex
update

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Related Work}
 
-The rapid growth of social media has made Social Networking Services (SNSs) increasingly attractive targets for spam and fraud, leading to a proliferation of sophisticated attacks.
-This trend is reflected in recent research, as papers have focused on identifying and classifying the various types of social media spam.
-Many of these studies employ techniques previously used to combat conventional email and web spam.
-SNSs also provide opportunities to take advantage of user reputation and other social graph-dependent features to improve classification.
-Nevertheless, most research has been carried out on publicly-available data from SNSs, making it difficult up until now to measure the effect of private user data on algorithms for detecting site misuse.
+The rapid growth of social media has made Social Networking Services (SNSs) 
+increasingly attractive targets for spam and fraud, leading to a proliferation 
+of sophisticated attacks. This trend is reflected in recent research, as papers 
+have focused on identifying and classifying the various types of social media spam. 
+Many of these studies employ techniques previously used to combat conventional 
+email and web spam. SNSs also provide opportunities to take advantage of user 
+reputation and other social graph-dependent features to improve classification. 
+Nevertheless, most research has been carried out on publicly-available data 
+from SNSs, making it difficult up until now to measure the effect of private user 
+data on algorithms for detecting site misuse.
 
@@
 
-Heymann et al.
-\cite{heymann} survey the field of spam on SNSs, identifying several common approaches.
-Identification-based approaches identify spam to train classifiers based on labels submitted by users or trusted moderators.
-Rank-based approaches demote visibility of questionable content, while interface-based approaches apply policies to prevent unwanted behavior.
-This work groups classification-based approaches with detection, although classifiers can be used in conjunction with user information to prevent spam before it happens.
+Heymann et al. \cite{heymann} survey the field of spam on SNSs, identifying 
+several common approaches. Identification-based approaches identify spam to 
+train classifiers based on labels submitted by users or trusted moderators. 
+Rank-based approaches demote visibility of questionable content, while interface-based 
+approaches apply policies to prevent unwanted behavior. This work groups 
+classification-based approaches with detection, although classifiers can be 
+used in conjunction with user information to prevent spam before it happens.
 
-A number of researchers have focused on collecting, identifying features and classifying various genres of spam on social networks.
-Zinman and Donath \cite{zinman} extract bundles of profile-based and comment-based features from MySpace profiles, but the relatively poor performance of their classifier highlights the difficulties in manual classification social network spam.
-Several studies take the approach of baiting spammers with social ``honeypots", profiles created with the sole intent of attracting spam.\cite{stringhini, lee} They then use the data collected to train classifiers with features including friend request rate and ratios of URLs to text.
-Webb et al.
-\cite{webb} use the honeypot approach as well and provide examples of various types of spammers, the typical demographics of their profiles as well as the web pages that they tend to advertise.
+A number of researchers have focused on collecting, identifying features 
+and classifying various genres of spam on social networks. Zinman and 
+Donath \cite{zinman} extract bundles of profile-based and comment-based 
+features from MySpace profiles, but the relatively poor performance of their classifier highlights 
+the difficulties in manual classification social network spam. Several 
+studies take the approach of baiting spammers with social ``honeypots", profiles 
+created with the sole intent of attracting spam.\cite{stringhini, lee} They 
+then use the data collected to train classifiers with features including 
+friend request rate and ratios of URLs to text. Webb et al. \cite{webb} use 
+the honeypot approach as well and provide examples of various types of spammers, 
+the typical demographics of their profiles as well as the web pages that they tend to advertise.
 
-Gao et al.
-\cite{gao} look at Facebook wall posts, analyzing temporal properties, URL characteristics, post ratios and other features of malicious accounts.
-They also pinpoint various spam ``campaigns" based on products advertised in a given time frame.
-They note that spam on Facebook often exhibits burstiness and is mainly sent from compromised accounts.
-Benevenuto et al.
-\cite{benevenuto} identify attributes of spam on video SNSs and use a Support Vector Machine (SVM) for classification.
+Gao et al. \cite{gao} look at Facebook wall posts, analyzing temporal properties, 
+URL characteristics, post ratios and other features of malicious accounts. They 
+also pinpoint various spam ``campaigns" based on products advertised in a given 
+time frame. They note that spam on Facebook often exhibits burstiness and is mainly 
+sent from compromised accounts. Benevenuto et al. \cite{benevenuto} identify 
+attributes of spam on video SNSs and use a Support Vector Machine (SVM) for classification.
 
-Not all undesirable content on SNSs is necessarily spam or a scam.
-SNSs and online communities witness inappropriate user behavior, where users post offensive and harassing content.
-Yin et al.
-\cite{yin} combine sentiment analysis and profanity word lists with contextual features to identify harassment on datasets from Slashdot and MySpace.
-Other work looks at SNSs as platforms to collect data about users in order to aid direct attacks on the user's computers or to compromise a large number of accounts.
-\cite{patsakis, huber}
+Not all undesirable content on SNSs is necessarily spam or a scam. SNSs and 
+online communities witness inappropriate user behavior, where users post offensive 
+and harassing content. Yin et al. \cite{yin} combine sentiment analysis and profanity 
+word lists with contextual features to identify harassment on datasets from Slashdot 
+and MySpace. Other work looks at SNSs as platforms to collect data about users in 
+order to aid direct attacks on the user's computers or to compromise a large number of accounts. \cite{patsakis, huber}
 
@@
 
-SocialSpamGuard \cite{jin} is a social media spam detection system that analyzes text and image features of social media posts.
-The demo system uses GAD clustering \cite{jingad} for sampling spam and ham posts, then trains a classifier with text and image features.
-However, the system is built on top of Facebook features that are publicly accessible and thus cannot make use of sensitive user data (\eg IP addresses) to increase its effectiveness.
+SocialSpamGuard \cite{jin} is a social media spam detection system that analyzes 
+text and image features of social media posts. The demo system uses GAD 
+clustering \cite{jingad} for sampling spam and ham posts, then trains a 
+classifier with text and image features. However, the system is built on top 
+of Facebook features that are publicly accessible and thus cannot make use of 
+sensitive user data (\eg IP addresses) to increase its effectiveness.
 
-De Wang et al.
-\cite{wang} propose a cross-site spam detection framework to share spam data across all social networking sites, building classifiers to identify spam in profiles, messages and web pages.
-This multi-pronged approach lends itself to associative classification, in which, for example, a message would be classified as spam if it contained a link to a web page that had a high probability of being spam.
-Unfortunately, the differing characteristics of various social networks \eg the length of messages in Facebook vs.
-Twitter, can reduce the benefits of sharing spam corpora across diverse sites.
+De Wang et al. \cite{wang} propose a cross-site spam detection framework to share 
+spam data across all social networking sites, building classifiers to identify spam 
+in profiles, messages and web pages. This multi-pronged approach lends itself to 
+associative classification, in which, for example, a message would be classified as 
+spam if it contained a link to a web page that had a high probability of being spam. 
+Unfortunately, the differing characteristics of various social networks \eg the length 
+of messages in Facebook vs. Twitter, can reduce the benefits of sharing spam corpora across diverse sites.
 
-Facebook \cite{stein} provides an overview of their ``immune system" defences against phishing, fraud and spam.
-The system is composed of classifier services, an ML-derived Feature Extraction Language (FXL), feature loops to aggregate and prepare features for classification and a policy engine to take action on suspected misuse.
-While the discussion remains high-level and includes few implemention particulars, it does include significant detail on the various types and characteristics of undesirable activity on the site, including fake profiles, harassment, compromised accounts, malware and spam.
+Facebook \cite{stein} provides an overview of their ``immune system" defences 
+against phishing, fraud and spam. The system is composed of classifier services, an 
+ML-derived Feature Extraction Language (FXL), feature loops to aggregate and prepare 
+features for classification and a policy engine to take action on suspected misuse. While 
+the discussion remains high-level and includes few implemention particulars, it does 
+include significant detail on the various types and characteristics of undesirable 
+activity on the site, including fake profiles, harassment, compromised accounts, malware and spam.
 
-In contrast to research that focuses on dynamically detecting spam based on user activity, Irani et al.
-\cite{irani} show that static features associated with user signups on MySpace are enough to train an effective social spam classifier.
-They note that C4.5 decision tree algorithms provide better performance than Na\"ive Bayes in this case.
-As in other works, this only examines publicly available profile information collected by social honeypots.
-Private data collected on users including browser features, IP addresses and geographic location would conceivably improve classifier performance substantially.
+In contrast to research that focuses on dynamically detecting spam based on user 
+activity, Irani et al. \cite{irani} show that static features associated with user signups 
+on MySpace are enough to train an effective social spam classifier. They note that
+C4.5 decision tree algorithms provide better performance than Na\"ive Bayes in this case. 
+As in other works, this only examines publicly available profile information collected 
+by social honeypots. Private data collected on users including browser features, IP addresses 
+and geographic location would conceivably improve classifier performance substantially.
 
-Bosma et al.
-\cite{bosma} explore user-generated spam reports as a tool for building an unsupervised spam detection framework for SNSs.
-Their approach counts the number of spam reports against a suspected spammer and adds weight to reports based on user reputation.
-Determining reputation and trustworthiness of users in social networks has been well studied \cite{bian, guha, zhang} and appears to be a promising addition to social spam classification.
-The framework uses a Bayesian classifier and links messages with similar content, but does not take into account other features.
-Nevertheless, this is one of the few studies to test its framework on non-public data, including private messages, spam reports and user profiles from a large Dutch social networking site.
+Bosma et al. \cite{bosma} explore user-generated spam reports as a tool for building an unsupervised 
+spam detection framework for SNSs. Their approach counts the number of spam reports against a 
+suspected spammer and adds weight to reports based on user reputation. Determining reputation 
+and trustworthiness of users in social networks has been well studied \cite{bian, guha, zhang} 
+and appears to be a promising addition to social spam classification. The framework uses a 
+Bayesian classifier and links messages with similar content, but does not take into account 
+other features. Nevertheless, this is one of the few studies to test its framework on 
+non-public data, including private messages, spam reports and user profiles from a 
+large Dutch social networking site.
 
@@
 Much work has been done on protecting traditional email systems from spam. 
-Blanzieri \cite{blanzieri} offers a comprehensive overview of machine learning techniques that can be applied to email filtering.
-Hao et al.
-\cite{hao} describe a reputation engine based on lightweight features such as geographic distance between sender and receiver, geolocation anomalies and diurnal patterns.
-While the target was conventional spam, these and similar features are applicable to spam on SNSs as well.
+Blanzieri \cite{blanzieri} offers a comprehensive overview of machine learning techniques 
+that can be applied to email filtering. Hao et al. \cite{hao} describe a reputation 
+engine based on lightweight features such as geographic distance between sender and 
+receiver, geolocation anomalies and diurnal patterns. While the target was conventional 
+spam, these and similar features are applicable to spam on SNSs as well.
 
-Whittaker et al.
-\cite{whittaker} describe a scalable phishing machine learning classifier and blacklisting system with high accuracy.
-Since a considerable amount of social media spam includes links to phishing sites, being able to detect them is critical.
-Along similar lines, Monarch \cite{thomas} is a system that provides scalable real-time detection of URLs that point to spam web pages as determined by URL features, page content and hosting properties of the target domain.
+Whittaker et al. \cite{whittaker} describe a scalable phishing machine learning 
+classifier and blacklisting system with high accuracy. Since a considerable 
+amount of social media spam includes links to phishing sites, being able to 
+detect them is critical. Along similar lines, Monarch \cite{thomas} is a 
+system that provides scalable real-time detection of URLs that point to spam web pages as 
+determined by URL features, page content and hosting properties of the target domain.
 
-Blog comment spam have also attracted considerable attention from researchers who have applied machine learning \cite{kolari, nag} and NLP \cite{mishne} techniques to the problem.
-Likewise, Markines et al.
-\cite{markines} apply similar techniques to spam on social bookmarking sites.
+Blog comment spam have also attracted considerable attention from researchers who 
+have applied machine learning \cite{kolari, nag} and NLP \cite{mishne} techniques to the 
+problem. Likewise, Markines et al. \cite{markines} apply similar techniques to spam on social bookmarking sites.
 
@@
 
-Many of the data mining algorithms used to detect spam and patterns of misuse on SNSs are designed with the assumption that the data and the classifier are independent.
-However, in the case of spam, fraud and other malicious content, users will often modify their behavior to evade detection, leading to degraded classifier performance and the need to re-train classifiers frequently.
-Several researchers tackle this adversarial problem.
-Dalvi et al.
-\cite{dalvi} offer a modified Na\"ive Bayes classifier to detect and reclassify data taking into account the optimal modification strategy that an adversary could choose.
-Lowd and Meek \cite{lowd} provide a framework for reverse engineering a classifier to determine whether an adversary can efficiently learn enough about a classifier to effectively defeat it.
+Many of the data mining algorithms used to detect spam and patterns of misuse 
+on SNSs are designed with the assumption that the data and the classifier are 
+independent. However, in the case of spam, fraud and other malicious content, 
+users will often modify their behavior to evade detection, leading to degraded 
+classifier performance and the need to re-train classifiers frequently. Several 
+researchers tackle this adversarial problem. Dalvi et al. \cite{dalvi} offer a 
+modified Na\"ive Bayes classifier to detect and reclassify data taking into account 
+the optimal modification strategy that an adversary could choose. Lowd and Meek \cite{lowd} 
+provide a framework for reverse engineering a classifier to determine whether 
+an adversary can efficiently learn enough about a classifier to effectively defeat it.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Data set}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Methodology}
@@
     \hline
-    \multirow{2}{*}{\% Fields Completed} & Spam & 22.15 & 14.00 & 30.75\\
-    & Ham & 79.29 & 86.00 & 32.85\\
+    \multirow{2}{*}{Fields Complete} & Spam & 22.15\% & 14.00\% & 30.75\\
+    & Ham & 79.29\% & 86.00\% & 32.85\\
+    \hline
+    \multirow{2}{*}{Account Lifetime} & Spam & 900.36 h & 1.12 h & 4163.25\\
+    & Ham & 8963.19 h & 5297.09 h & 11750.39 \\
     \hline
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/lifetime.pdf}
+    \caption{Sender account lifetime at the time that the message was sent.}
+    \label{fig:lifetime}
+\end{figure}
+
 \textbf{Sender Email:} Users must provide email addresses upon signup. The site does not allow users to contact other 
@@
 respectively). For spam user accounts, Hotmail, Yahoo and Gmail accounted for 7\%, 72\% and 6\% respectively. This striking 
-predominance of Yahoo email accounts among spammers can be seen in figure \ref{fig:email}. While we have no explanation for 
-this, figure \cite{fig:ysmap} shows that Ghana and Nigeria alone account for 38\% of the IP addresses associated with these 
-Yahoo accounts. This popularity of Yahoo among West African users has been noted before\cite{burrell}, and appears to be 
-due to the early penetration of Yahoo services in the region.
+predominance of Yahoo email accounts among spammers can be seen in figure \ref{fig:email}. While we have no certain 
+explanation for this, Figure \ref{fig:yspam} shows that Ghana and Nigeria alone account for 38\% of the IP 
+addresses associated with these Yahoo accounts. This popularity of Yahoo among West African users has 
+been noted before\cite{burrell}, and appears to be due to the early penetration of Yahoo services in the region.
 
@@
 
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/photos.pdf}
+    \caption{Number of photos associated with sender account.}
+    \label{fig:photos}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/friends.pdf}
+    \caption{Number of friends associated with sender account.}
+    \label{fig:friends}
+\end{figure}
+
 \textbf{Sender Age:}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Classification}
@@
 
-In order to train our classifiers, we had to create a feature matrix from the message and account data. We did this using 
+In order to train our classifiers, we created a feature matrix from the message and account data. We did this using 
 Perl, Spark, and Lucene analyzers. The first step was to generate a dictionary sorted by document frequencies in descending 
@@
 
+We initially trained and evaluated three classifiers on a small subset of the data: Na\"ive Bayes, Linear Regression and SVM. 
+We chose an 11k spam and 11k ham sample for training, using this particular sample size due to the size of the spam corpus 
+section containing recipient age feature data.
+
 %classification:
 %- implemented 3 models: naive bayes, linear regression, and SVM.
-%- used scalanlp for SVM: Pegasos: Primal Estimated sub-GrAdient SOlver for SVM Extended with Wang, Crammer, Vucetic's work for Multiclass.
-The optimizer runs a stochastic subgradient descent on the primal objective using batches provided.
 %- used scalanlp for naive bayes
@@
 
-% scalanlp svm
+We initially used ScalaNLP's built-in SVM solver, which implements the Pegasos algorithm \cite{pegasos} as extended 
+by Wang, Crammer and Vucetic\cite{wang2}. The optimizer runs stochastic subgradient descent on the primal objective 
+using batches provided. However, the ScalaNLP SVM interface does not allow changing kernel functions or parameters, 
+which seems to have led to the classifier's sub-par performance.
 
-% libsvm
+We then turned to LIBSVM, a library for SVM training and classification in C++ and Java, with wrappers for 
+Python and other languages. The package allows one to specify the type of kernel to use, kernel parameters 
+($\gamma$, $\rho$, $d$), as well as a penalty parameter ($C$). We used Perl to convert the sparse feature 
+matrices generated for use with ScalaNLP to a format recognized by LIBSVM. We then used one of the package'
+s tools to perform simple scaling on the data.
 
-LIBSVM is library for SVM training and classification in C++ and Java, with wrappers for Python and other languages.
-The package allows one to specify the type of kernel to use, kernel parameters ($\gamma$, $\rho$, $d$), as well as a penalty parameter ($C$).
-We used Perl to convert the sparse feature matrices generated for use with ScalaNLP to a format recognized by LIBSVM.
-We then used one of the package's tools to perform simple scaling on the data.
-
-Given the complexity of choosing kernels and parameters, the package includes a Python script that performs cross-validation to suggest appropriate $\gamma$ and $C$ values for the recommended radial basis function (RBF) kernel.
-The RBF kernel is a real-valued function such that: $$K(\mathbf{x}_i, \mathbf{x}_j) = (\gamma\mathbf{x}_i^\intercal\mathbf{x}_j+r)^d, \gamma>0$$
+Given the complexity and time-intensive nature of choosing and testing parameters, the 
+package includes a Python script that performs cross-validation to suggest appropriate 
+$\gamma$ and $C$ values for the recommended radial basis function (RBF) kernel. The RBF kernel is 
+a real-valued function such that: 
+$$K(\mathbf{x}_i, \mathbf{x}_j) = (\gamma\mathbf{x}_i^\intercal\mathbf{x}_j+r)^d, \gamma>0$$
 We ran this script on a subset of 2500 ham and 2500 spam messages to obtain $C = 8$ and $\gamma = 0.0078125$.
 
-The LIBSVM scaling tool increases file size considerably for data sets with large feature matrices.
-This large file size, coupled with the computational resources required for SVM training on large data sets, prevented us from completing our analysis of the
-
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Analysis}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Future Work}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/friends.pdf}
-    \caption{Number of friends associated with sender account.}
-    \label{fig:friends}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
     \includegraphics[width=\linewidth]{figures/linear-roc.pdf}
@@
     \centering
-    \includegraphics[width=\linewidth]{figures/photos.pdf}
-    \caption{Number of photos associated with sender account.}
-    \label{fig:photos}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
     \includegraphics[width=\linewidth]{figures/sex.pdf}

### 2012-05-08 docs/paper/paper.tex
update

@@
 %
+
 \subsection{Linear Regression}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results}
+\section{Results \& Analysis}
 
 \subsection{Na\"ive Bayes}
+
+One of the advantages of the Na\"ive Bayes classifier was that it took very little time to train the classifier, 
+and the steps involved were all trivially parallelizable, which makes it convenient for large data sets.
+
+The decline in accuracy after adding expert features is most likely explained by the "naive" assumption in this 
+algorithm that the features are independent. Several of the features that we used, \eg{account lifetime, 
+and whether a user has friends or photos}, are in fact likely to be co-dependent, leading to an overweight of 
+these features during classification. The wealth of expert features available to a SNS, many of which are 
+not independent, makes this the least compelling classifier choice that we examined.
 

### 2012-05-09 docs/paper/paper.tex
update

@@
 
-Unlike spam messages, we did not have access to a comparable corpus of human-labeled ham messages.
-To simplify the labeling of ham messages, we made the assumption that messages remaining in the inboxes of active users after a period of several months would most likely not be spam.
-Consequently, we extracted a working set of 2 million ham messages that were sent in the beginning of January, 2012 and still existed in the recipient's inbox.
-To reduce the possibility of collecting messages sent by uncaught spammers to inactive or dormant users, we selected messages only from users and to recipients who had logged in within the last two weeks.
-As in the case of spam, we collected two million messages with associated account data and merged them into text files for feature extraction.
+Unlike spam messages, we did not have access to a comparable corpus of human-labeled ham 
+messages. To simplify the labeling of ham messages, we made the assumption that messages remaining in 
+the inboxes of active users after a period of several months would most likely not be spam. 
+Consequently, we extracted a working set of 2 million ham messages that were sent in late December 2011 and early 
+January 2012 and still existed in the recipient's inbox in March 2012. To reduce the 
+possibility of collecting messages sent by uncaught spammers to inactive or dormant users, 
+we selected messages only from users and to recipients who had logged in within the last two 
+weeks. As in the case of spam, we collected two million messages with associated account 
+data and merged them into text files for feature extraction.
 
@@
 
-We attempted to train the multiple logistic regression classifier implemented by the ScalaNLP package, but encountered a bug that prevented the model from being generated correctly.
-With the expectation that SVMs would outperform logistic regression, we decided to focus our efforts on tuning the SVM instead of pursuing logistic regression.
+We attempted to train the multiple logistic regression classifier implemented by the ScalaNLP package, 
+but encountered a bug that prevented the model from being generated correctly. With the expectation 
+that SVMs would outperform logistic regression, we decided to focus our efforts on tuning the SVM 
+instead of pursuing logistic regression.
 
@@
 ($\gamma$, $\rho$, $d$), as well as a penalty parameter ($C$). We used Perl to convert the sparse feature 
-matrices generated for use with ScalaNLP to a format recognized by LIBSVM. We then used one of the package'
-s tools to perform simple scaling on the data.
+matrices generated for use with ScalaNLP to a format recognized by LIBSVM. We then used one of the package's 
+tools to perform simple scaling on the data.
 
@@
 
+Our initial results using the Pegasos SVM solver included in ScalaNLP were disappointing. The 
+classifier trained on the bag-of-words feature set (from 22k messages) yielded a 77\% mean accuracy 
+over 10-fold cross-validation. The addition of expert features actually decreased this accuracy to 
+51\%. Table \{tab:svmfeats} provides a list of top features for the bag-of-words and expert 
+feature models. We see that for the bag-of-words classifier, the words "year," "new," "happy,"  
+"2012," "wish," and "you" all have high weights, which makes sense given that the ham data was 
+collected around the beginning of January.
+
 \begin{table}
@@
 \section{Future Work}
+
+\subsection{Data Collection}
+
+The ham message collection methodology used in this project may have been flawed, leading questionable 
+performance for several classifiers. The 2 million ham messages were sent over a span of 10 days, from 
+December 23, 2011 to January 1, 2012. While we chose the corpus' start date as exactly four months before 
+extracting the data, the coincidental timing with a global holiday seems to have introduced unusual 
+skew in the bag-of-words features. In contrast, the majority of messages in the spam corpus span a period of 
+7 months, with only a slight linear increase in volume over the interval.  Consequently, as seen in 
+the Linear Regression and SVM feature weight tables, words typically associated with the New Year gained 
+undue wieght during classification.  We plan to address this issue by creating a new ham message corpus 
+through sampling that matches the temporal distribution of spam messages more closely.
 

### 2012-05-09 docs/paper/paper.tex
update

@@
 over 10-fold cross-validation. The addition of expert features actually decreased this accuracy to 
-51\%. Table \{tab:svmfeats} provides a list of top features for the bag-of-words and expert 
+51\%. Table \ref{tab:svmfeats} provides a list of top features for the bag-of-words and expert 
 feature models. We see that for the bag-of-words classifier, the words "year," "new," "happy,"  
@@
 
-The ham message collection methodology used in this project may have been flawed, leading questionable 
+The ham message collection methodology used in this project may have been flawed, leading to questionable 
 performance for several classifiers. The 2 million ham messages were sent over a span of 10 days, from 

### 2012-05-09 docs/paper/paper.tex
update

@@
 attracts a wide variety of financial scams, ranging from Nigerian
-"419" scams to romance scams. Another prevalent problem is spam with
+``419" scams to romance scams. Another prevalent problem is spam with
 links to third-party websites, directing users to various porn/webcam
@@
 100 million private messages and another 2 million messages that have been 
-labeled as spam by users. Other data includes 40 million or so "wall" 
+labeled as spam by users. Other data includes 40 million or so ``wall" 
 comments, 5 million photos, and 8 million photo comments.
@@
 \item Month of birth 
-\item Whether sender has friends (boolean)
-\item Whether sender has photo(s) (boolean)
-\item Profile country and IP-detected country match (boolean)
-\item IP-detected country is Ghana, Nigeria, Senegal, Malaysia, Turkey, Gambia, Ivory Coast, Togo (``high-risk" countries) (boolean)
+\item Whether sender has friends , boolean
+\item Whether sender has photo(s) , boolean
+\item Profile country and IP-detected country match , boolean
+\item IP-detected country is Ghana, Nigeria, Senegal, Malaysia, Turkey, Gambia, Ivory Coast, Togo (``high-risk" countries) , boolean
 \end{itemize}
@@
 % Modified from Reynold's HW1  write-up
-%We use a na\"ive Bayes model, in which the documents are generated according to the following process.  
+%We use a na\``ive Bayes model, in which the documents are generated according to the following process.  
 %First, a class \(c\) is chosen from a probability distribution \(\Pr(c)\).
@@
 
-The decline in accuracy after adding expert features is most likely explained by the "naive" assumption in this 
+The decline in accuracy after adding expert features is most likely explained by the ``naive" assumption in this 
 algorithm that the features are independent. Several of the features that we used, \eg{account lifetime, 
@@
 We tried a number of values for $\lambda$: 1, 10, 50, 100, 500, 1000, 1500, 3000. Figure \ref{fig:lambda}
+\ref{fig:roclin}
 
@@
     \label{fig:lambda}
+\end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/linear-roc.pdf}
+    \caption{Linear regression ROC curves for classifier using bag-of-words and bag-of-words with ``expert" features.}
+    \label{fig:roclin}
 \end{figure}
@@
 \begin{table}
-\begin{tabular}{l|l||l|l}
+\begin{tabular}{l|r||l|r}
 \multicolumn{2}{c}{\textbf{Ham}} & 
@@
 \end{tabular}
-\caption{Linear Regression top features (bag-of-words + expert features) with weights. Expert features
-are denoted by the 'F-' prefix.}
+\caption{Linear Regression top features (bag-of-words \& expert features) with weights. Expert features
+are denoted by an 'F-' prefix.}
 \label{tab:lrfeats}
@@
 
-Our initial results using the Pegasos SVM solver included in ScalaNLP were disappointing. The 
+Our initial results using the Pegasos SVM solver included in ScalaNLP were lower than expected. The 
 classifier trained on the bag-of-words feature set (from 22k messages) yielded a 77\% mean accuracy 
@@
 51\%. Table \ref{tab:svmfeats} provides a list of top features for the bag-of-words and expert 
-feature models. We see that for the bag-of-words classifier, the words "year," "new," "happy,"  
-"2012," "wish," and "you" all have high weights, which makes sense given that the ham data was 
-collected around the beginning of January.
+feature models. We see that for the bag-of-words classifier, the words ``year," ``new," ``happy,"  
+``2012," ``wish," and ``you" all have high weights, which makes sense given that the ham data was 
+collected around the beginning of 2012. The table shows that three expert features: account lifetime, 
+friends and recipient age, are weighted very heavily in comparison to the remaining features. These 
+skewed weights (and the inability to tune the SVM solver via ScalaNLP) prompted us to test LIBSVM 
+on the same data.
 
 \begin{table}
-\begin{tabular}{l|l||l|l}
+\begin{tabular}{l|r||l|r}
 \multicolumn{2}{c}{\textbf{Bag-of-Words}} & 
@@
 \hline
-\textbf{Feature} & \textbf{Ham Weight} & \textbf{Feature} & \textbf{Ham Weight} \\
+\textbf{Feature} & \textbf{Ham Wt.} & \textbf{Feature} & \textbf{Ham Wt.} \\
 \hline
 
-year & 0.2253 & F-account\_lifetime & -0.98\\
-new & 0.225 & F-friends & -0.3671\\
+year & 0.2253 & F-account\_lifetime & -0.9800\\
+new & 0.2250 & F-friends & -0.3671\\
 am & -0.1722 & F-recip\_age & -0.2129\\
-happy & 0.16 & F-age & -0.183\\
+happy & 0.1600 & F-age & -0.1830\\
 d & 0.1513 & i & -0.05231\\
 ok & -0.1497 & you & -0.0483\\
-yahoo & -0.136 & to & -0.04026\\
-it & 0.12 & and & -0.03636\\
-me & -0.118 & F-month\_of\_birth & -0.03452\\
+yahoo & -0.1360 & to & -0.04026\\
+it & 0.1200 & and & -0.03636\\
+me & -0.1180 & F-month\_of\_birth & -0.03452\\
 too & 0.1173 & F-country\_match & -0.0269\\
 i'm & 0.1116 & me & -0.0224\\
-haha & 0.0998 & am & -0.022\\
+haha & 0.0998 & am & -0.0220\\
 can & -0.0971 & my & -0.0209\\
@@
 \end{table}
+
+Our results with LIBSVM's RBF kernel were consistently better than any of the other classifiers we tested. We 
+evaluated the classifier's performance on five pairs of bag-of-words and expert feature data sets ranging in 
+size from 2000 to 100k messages. Table \ref{tab:svmacc} shows results of 10-fold cross-validation. Our mean 
+accuracy over the bag-of-words data sets was 85.73\% and 98.41\% over the bag-of-words with expert features. We 
+observed a slight decrease in performance on the bag-of-words data sets as the data size grew. Figure \ref{fig:roc-25}
+shows the AUCs and ROC curves for the 50k data sets, making clear the the significantly greater AUC resulting 
+from the addition of expert features.
+
+\begin{table}
+\begin{tabular}{r|r|r}
+\hline
+\textbf{Dataset Size} & \textbf{Bag-of-Words} & \textbf{Expert Features} \\
+\hline
+2000 & 86.70\% & 98.35\% \\
+10000 & 86.23\% & 98.49\% \\
+22000 & 86.81\% & 98.33\% \\
+50000 & 84.59\% & 98.43\% \\
+100000 & 84.33\% & 98.45\% \\
+\end{tabular}
+\caption{SVM 10-fold cross-validation accuracy over data sets of varying sizes (half spam, half ham)} 
+\label{tab:svmacc}
+\end{table}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/roc-25.pdf}
+    \caption{LIBSVM ROC curves for classifier using bag-of-words and bag-of-words with ``expert" features on 50k messages 
+     (25k spam, 25k ham)}
+    \label{fig:roc-25}
+\end{figure}
 
@@
 
-figure \ref{fig:cprof}
-figure \ref{fig:cip}
-figure \ref{fig:drop}
-figure \ref{fig:friends}
-figure \ref{fig:roclin}
-figure \ref{fig:photos}
-figure \ref{fig:recipage}
-figure \ref{fig:sendage}
-figure \ref{fig:sendsex}
-figure \ref{fig:roc-25}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figures/linear-roc.pdf}
-    \caption{Linear regression ROC curves for classifier using bag-of-words and bag-of-words with ``expert" features.}
-    \label{fig:roclin}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figures/roc-25.pdf}
-    \caption{LIBSVM ROC curves for classifier using bag-of-words and bag-of-words with ``expert" features on 50k messages.}
-    \label{fig:roc-25}
-\end{figure}
-
-\begin{figure}[h]
-    \centering
-    \includegraphics[width=\linewidth]{figures/sex.pdf}
-    \caption{Sender sex as stated on profile.}
-    \label{fig:sendsex}
-\end{figure}
+%\begin{figure}[h]
+%    \centering
+%    \includegraphics[width=\linewidth]{figures/sex.pdf}
+%    \caption{Sender sex as stated on profile.}
+%    \label{fig:sendsex}
+%\end{figure}
+%
 

### 2012-05-09 docs/paper/paper.tex
update

@@
 the extracted features, comparing the ham and spam corpora. These statistics were generated 
-using SQL on a table created from the merged and cleaned data generated in the process described above.
+using SQL on a table created from the merged and cleaned data generated in the process described 
+above. We then chose a subset of these features to use for training and evaluating our classifiers.
 
@@
 and omit spaces after these characters). Second, we calculated the ratio of uppercase to 
-lowercase letters in the text, after observing relatively high amounts of uppercase text in spam messages.
+lowercase letters in the text, after observing relatively high amounts of uppercase text in 
+spam messages.
 %% add numbers
@@
 
-We chose message and user account features to analyze primarily by the amount of data that we had for each them, while making sure to include the most vital account information.
-Table \ref{tab:feats} provides averages on a number of the account statistics that we examined.
+We chose message and user account features to analyze primarily by the amount of data that 
+we had for each them, while making sure to include the most vital account information. 
+Table \ref{tab:feats} offers statistics on a number of the account features that we examined.
+
 \begin{table}
@@
 with the 73\% that US-based IP addresses comprised for ham users whose profiles stated that 
-they were in the United States (Figure \ref{fig:usaham}).
+they were in the United States (Figure \ref{fig:usaham}). During feature extraction, we also 
+generated a boolean feature (IP mismatch) indicating whether the IP-detected country and 
+the profile country matched.
 
@@
 
-\textbf{Sender IPs:} Spammers in general had a smaller number of unique IP addresses associated with them than ham users, as per figure \ref{fig:uniqip}.
-The mean for spammers was 11.82 with a standard deviation of 52.52 and a median of 2, whereas ham users saw a mean of 175.51, a standard deviation of 322.19 and median of 64.
-This is likely due to their comparatively low length of time on the site and the fact that many legitimate users are behind NATs that yield a high number of dynamic IP addresses.
-One possibility for future research is to investigate the ratio of unique IPs for users as a function of the time they have been registered.
-Likewise, examining unique /16 network blocks instead of unique IP addresses could mitigate the influence of NATs.
+\textbf{Sender IPs:} Spammers in general had a smaller number of unique IP addresses associated 
+with them than ham users, as per figure \ref{fig:uniqip}. The median number of IP addresses for 
+spam users was 2, whereas ham users saw a median of 64. This is likely due to the comparatively 
+shorter length of time that spammers remain on the site before being deleted by moderators (Figure 
+\label{fig:lifetime} show a histogram of the time elapsed between registration and message sending 
+time).  Similarly, we believe that many legitimate users are behind NATs, yielding a high number 
+of dynamic IP addresses. One possibility for future research is to investigate the ratio of 
+unique IPs for users as a function of the time they have been registered. Likewise, examining 
+unique /16 network blocks instead of unique IP addresses could mitigate the influence of NATs.
 
@@
 
-\textbf{Extra Profile Features:} We found that the number of photos that user has uploaded is also indicative of their spam or ham reputation. Spammers
+\textbf{Photos and Friends:} We found that both the number of ``friends" that a user has (connections 
+to other users) as well as the number of photos that user has uploaded were also indicative of their 
+spam or ham reputation. Spammers typically had one or no photos, while ham users typically had larger 
+quantities of both, as visible in Figures \ref{fig:photos} and \ref{fig:friends}. It is important to note, 
+however, that we would expect a legitimate user's average number of photos and friends to grow over 
+as a function of account lifetime. It would be quite reasonable for new ham users to have very photos 
+or friends and we plan to further examine the relationship of these attributes over account lifetimes 
+for both ham and spam users.
 
@@
 
-\textbf{Sender Age:}
-
-\textbf{Sender Name:}
-
-\textbf{Username, Name:} While we collected these fields, we did not end up processing or analyzing them further.
-That said, we saw considerable repetition in the usernames of scammers and it seems that analyzing the substrings in these fields could yield useful features.
-
-\textbf{Sender Sex:}
-
-\subsection{Mixed Features}
-
-\textbf{Account Lifetime.}
-
-\textbf{IP Mismatch.} This
+\textbf{Username, Name:} While we collected these fields, we did not analyze them further, due to 
+time constraints. Nevertheless, we saw considerable repetition in the usernames of scammers 
+and it seems that analyzing the substrings in these fields could yield useful features.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Classification}
+\section{Classification \& Results}
 
@@
 
+\subsection{Na\"ive Bayes Results}
+
+One of the advantages of the Na\"ive Bayes classifier was that it took very little time to train the classifier, 
+and the steps involved were all trivially parallelizable, which makes it convenient for large data sets.
+
+The decline in accuracy after adding expert features is most likely explained by the ``naive" assumption in this 
+algorithm that the features are independent. Several of the features that we used, \eg{account lifetime, 
+and whether a user has friends or photos}, are in fact likely to be co-dependent, leading to an overweight of 
+these features during classification. The wealth of expert features available to a SNS, many of which are 
+not independent, makes this the least compelling classifier choice that we examined.
+
 \subsection{Linear Regression}
@@
 
-\subsection{Logistic Regression}
-
-We attempted to train the multiple logistic regression classifier implemented by the ScalaNLP package, 
-but encountered a bug that prevented the model from being generated correctly. With the expectation 
-that SVMs would outperform logistic regression, we decided to focus our efforts on tuning the SVM 
-instead of pursuing logistic regression.
-
-\subsection{Support Vector Machines (SVMs)}
-
-%intro to svms
-
-Support Vector Machines are a classification method \cite{boser}, \cite{cortesv95}
-
-We initially used ScalaNLP's built-in SVM solver, which implements the Pegasos algorithm \cite{pegasos} as extended 
-by Wang, Crammer and Vucetic\cite{wang2}. The optimizer runs stochastic subgradient descent on the primal objective 
-using batches provided. However, the ScalaNLP SVM interface does not allow changing kernel functions or parameters, 
-which seems to have led to the classifier's sub-par performance.
-
-We then turned to LIBSVM, a library for SVM training and classification in C++ and Java, with wrappers for 
-Python and other languages. The package allows one to specify the type of kernel to use, kernel parameters 
-($\gamma$, $\rho$, $d$), as well as a penalty parameter ($C$). We used Perl to convert the sparse feature 
-matrices generated for use with ScalaNLP to a format recognized by LIBSVM. We then used one of the package's 
-tools to perform simple scaling on the data.
-
-Given the complexity and time-intensive nature of choosing and testing parameters, the 
-package includes a Python script that performs cross-validation to suggest appropriate 
-$\gamma$ and $C$ values for the recommended radial basis function (RBF) kernel. The RBF kernel is 
-a real-valued function such that: 
-$$K(\mathbf{x}_i, \mathbf{x}_j) = (\gamma\mathbf{x}_i^\intercal\mathbf{x}_j+r)^d, \gamma>0$$
-We ran this script on a subset of 2500 ham and 2500 spam messages to obtain $C = 8$ and $\gamma = 0.0078125$.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results \& Analysis}
-
-\subsection{Na\"ive Bayes}
-
-One of the advantages of the Na\"ive Bayes classifier was that it took very little time to train the classifier, 
-and the steps involved were all trivially parallelizable, which makes it convenient for large data sets.
-
-The decline in accuracy after adding expert features is most likely explained by the ``naive" assumption in this 
-algorithm that the features are independent. Several of the features that we used, \eg{account lifetime, 
-and whether a user has friends or photos}, are in fact likely to be co-dependent, leading to an overweight of 
-these features during classification. The wealth of expert features available to a SNS, many of which are 
-not independent, makes this the least compelling classifier choice that we examined.
-
-\subsection{Linear Regression}
+\subsection{Linear Regression Results}
 
@@
 
-\subsection{SVMs}
+\subsection{Logistic Regression}
+
+We attempted to train the multiple logistic regression classifier implemented by the ScalaNLP package, 
+but encountered a bug that prevented the model from being generated correctly. With the expectation 
+that SVMs would outperform logistic regression, we decided to focus our efforts on tuning the SVM 
+instead of pursuing logistic regression.
+
+\subsection{Support Vector Machines (SVMs)}
+
+%intro to svms
+
+Support Vector Machines are a classification method \cite{boser}, \cite{cortesv95}
+
+We initially used ScalaNLP's built-in SVM solver, which implements the Pegasos algorithm \cite{pegasos} as extended 
+by Wang, Crammer and Vucetic\cite{wang2}. The optimizer runs stochastic subgradient descent on the primal objective 
+using batches provided. However, the ScalaNLP SVM interface does not allow changing kernel functions or parameters, 
+which seems to have led to the classifier's sub-par performance.
+
+We then turned to LIBSVM, a library for SVM training and classification in C++ and Java, with wrappers for 
+Python and other languages. The package allows one to specify the type of kernel to use, kernel parameters 
+($\gamma$, $\rho$, $d$), as well as a penalty parameter ($C$). We used Perl to convert the sparse feature 
+matrices generated for use with ScalaNLP to a format recognized by LIBSVM. We then used one of the package's 
+tools to perform simple scaling on the data.
+
+Given the complexity and time-intensive nature of choosing and testing parameters, the 
+package includes a Python script that performs cross-validation to suggest appropriate 
+$\gamma$ and $C$ values for the recommended radial basis function (RBF) kernel. The RBF kernel is 
+a real-valued function such that: 
+$$K(\mathbf{x}_i, \mathbf{x}_j) = (\gamma\mathbf{x}_i^\intercal\mathbf{x}_j+r)^d, \gamma>0$$
+We ran this script on a subset of 2500 ham and 2500 spam messages to obtain $C = 8$ and $\gamma = 0.0078125$.
+
+\subsection{SVM Results}
 
@@
 
+Another direction for future research would be to examine data other than messages. We have access to millions 
+of wall posts and photo comments, making it possible to investigate the characteristics of spam in these genres. 
+Wall posts and comments tend to be comparatively short, so we might
+
 \subsection{Spam Grouping}
 
-As a part of this project, we performed some qualitative analysis on spam and differentiated between the major categories of undesirable messages found on this SNS. However, we
+As a part of this project, we performed some qualitative analysis on spam and differentiated between 
+some major categories of undesirable messages that we observed on the InterPals web site. However, we would 
+like to gain a better understanding of the relative volume of each type of scam
 
@@
 
-It is also anecdotally clear from moderator log messages that spammers re-use their profile photos and these photos are often of celebrities, porn stars or simply stock photos. Hashing these images
+It is also anecdotally clear from moderator log messages that spammers re-use their profile photos and 
+these photos are often of celebrities or simply stock photos. Hashing these images
 

### 2012-05-09 docs/paper/paper.tex
update

@@
 
-\title{Detecting Spam on Social Networking Sites}
+\title{Feature Selection and Classification of \\
+    Spam on Social Networking Sites}
 
@@
 of data on the sender, which can be used to build more accurate detection mechanisms. 
-We analyze public and private data from a popular social network in order to 
-gain insight into the various features spam messages and the accompanying user 
-accounts and profiles available to site operators. By examining the ....
+We analyze 4 million private messages as well as other public and private data from 
+a popular social network in order to gain insight into the various features of 
+spam messages and the accompanying user accounts data available to site operators. 
+We use these insights to choose features that best differentiate spammers from 
+legitimate, ``ham," users.  Finally, we extract these features from the site's data 
+and use them to train and evaluate classifiers.
 \end{abstract}
@@
 
-Social networks of any significant size see constant spam, scams
-and phishing attacks. The nature of these attacks can be quite diverse
-and difficult detect. Marketers can spam members with unwanted
-advertisements, fraudsters lure users with advance fee frauds and
-other confidence tricks, while others may attempt to steal user
-information by directing users to external phishing pages. 
-Sites with global reach see communication among its members in a 
-number foreign languages with varying levels
-of ability. This means that much benign content shares characteristics
-like misspellings, awkward phrases, etc. that might have made certain
-types of common frauds and spam more easy to distinguish on US-based
-(or English-language) sites.
+Social networks of any significant size see a constant stream of spam, scams and phishing attacks. 
+The nature of this unwanted activity can be quite diverse, specific to a site's target audience and 
+often not trivial to detect. Marketers can spam members with unwanted advertisements, fraudsters 
+lure users with advance fee frauds and other confidence tricks, while others may attempt to steal user
+information by directing users to external phishing pages. In contrast to email spam, social spam 
+has a more personal component, since both spammers and legitimate users (``ham" users) have accompanying 
+profiles with descriptive information.
+
+Sites with global reach see communication among its members in a number foreign languages with 
+varying levels of ability. This means that much benign content shares characteristics like misspellings, 
+awkward phrases, etc. that might have made certain types of common frauds and spam more easy to 
+distinguish on US-based (or English-language) sites.
+
+Recent research has begun to evaluate Email spam and techniques with which to combat it has been studied 
+extensively. Like
+
 % focus on extra impact of non-public features vs. public (bag-of-words)
@@
 
-Email spam and techniques with which to combat it has been studied 
-extensively. Like
 This paper examines in detail the types of malicious and benign
@@
 
-In the course of collecting the data, we observed a number of different types of undesirable messages.
-Our ongoing observation of spam on the InterPals website provided direct intuition into the major classes of unwanted behavior that we classified as spam for this project.
-We noted the following broad categories:
+In the course of collecting the data, we observed a number of different types of 
+undesirable messages. Our ongoing observation of spam on the InterPals website provided 
+direct intuition into the major classes of unwanted behavior that we classified as spam for 
+this project. We noted the following broad categories:
 
-\textbf{Advance Fee Fraud} We observed this scam to come in various flavors.  (inheritance, lottery,  etc)
-\textbf{Romance Scams:} These 
-\textbf{Sexual Solicitation}
-\textbf{Porn Ads}
-\textbf{Advertising External Sites}
-\textbf{Webcam Ads}
-\textbf{Money Muling}
-\textbf{Begging \& Gift Requests}
-\textbf{Business Proposals}
+\begin{itemize}
+\item \textbf{Advance Fee Fraud} We observed this scam to come in various flavors.  (inheritance, lottery,  etc)
+\item \textbf{Romance Scams:} These 
+\item \textbf{Sexual Solicitation}
+\item \textbf{Porn Ads}
+\item \textbf{Advertising External Sites}
+\item \textbf{Webcam Ads}
+\item \textbf{Money Muling}
+\item \textbf{Begging \& Gift Requests}
+\item \textbf{Business Proposals}
+\end{itemize}
 
-This categorizations are based only on observation. Analyzing the relevant terms in log comments entered by moderators when deleting spammer accounts provides the following breakdown:
+These categorizations are based only on cursory manual observation of a sample of several tens of thousands 
+spam-labeled messages. We plan to quantify the volume of messages in each of these categories and 
+attempt to provide a finer granularity of categorization in future research.
+
+%Analyzing the relevant terms in log comments entered by moderators when 
+%deleting spammer accounts provides the following breakdown:
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analysis}
-% Vern wants site owner perspective on how to apply classifiers
-From a site's perspective, ...
-
-- plan to implement SVM, despite speed tradeoff
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Future Work}
@@
 of wall posts and photo comments, making it possible to investigate the characteristics of spam in these genres. 
-Wall posts and comments tend to be comparatively short, so we might
+Wall posts and comments tend to be comparatively short, which should present a slightly different classification 
+problem.
+
+We have also recently observed spammers attacking the report system by reporting large numbers of legitimate users 
+with complaints consisting of nonsensical text or form messages. In addition to offering another application for 
+classifiers, the data we are collecting should offer a window into the retaliatory techniques of miscreants on 
+SNSs.
 
@@
 some major categories of undesirable messages that we observed on the InterPals web site. However, we would 
-like to gain a better understanding of the relative volume of each type of scam
+like to gain a better understanding of the relative volume of each type of scam. To this end, we plan to 
+experiment with clustering algorithms to attempt to identify and quantify subclasses of spam.
 
-Characteristics of spam messages vary widely depending on the type of spam
-we currently lump all spam into a single category; we plan to build on the classes we mentioned in section()
-We are currently in the process of extending the moderation interface to include more specific tagging capabilities
-%- certain classifiers might work better on certain types of spam
-%- clustering to see how well these classes fit the spam types on the site
+Furthermore, characteristics of spam messages vary substantially depending on the type of spam or scam in question. 
+The classifiers that we trained and evaluated in this project all spam messages as a single category. We are currently 
+investigating multiclass classifiers as a potential tool to identify the type of 
+spam in question with greater precision. Specifically, we are considering multiclass SVMs, which reduce 
+multiclass problems into multiple binary classification problems.\cite{duan}
+
+From a site's perspective, this finer granularity is desirable not only 
+because specialty classifiers could be trained to recognize specific classes of spam, but also in order to 
+trigger different courses of action (\eg automatic account deletion, warning messages, reports to human moderators) 
+based on the class and probability of spam detected.
+
+We are currently extending the moderation interface to include more specific tagging capabilities, which should aid 
+in generating labels for multiclass classifiers and for seeding semi-supervised clustering algorithms (see Basu et al. 
+\cite{basu}).
 
@@
 
-In this project, we chose only a small subset of ``expert" features to use.
-However, we believe that it would be beneficial to train and evaluate the classifiers on a wider range of features.
-The features would include both the ones we extracted in this project but omitted from the classifier features matrices, as well as new features.
-Adding N-grams and message similarity tests seem particularly promising.
+In this project, we chose only a small subset of ``expert" features to use. However, we believe that 
+it would be beneficial to train and evaluate the classifiers on a wider range of features. The features 
+would include both the ones we extracted in this project but omitted from the classifier features matrices, 
+as well as new features. Adding N-grams and message similarity tests seem particularly promising. 
 N-grams would allow the classifier to take into account certain phrases \eg{``Western Union" or ``money transfer"}.
 
-We have also recently begun to collect browser fingerprints from users based on user agent and JavaScript-acquired browser plugin details, time zone, screen size, color depth, system fonts, keyboard layout and locale information.
-Subsequent classifiers could leverage this information either as a hash that effectively tags individual known spammers or in part, adding features based on time zone or keyboard layout.
+We have also recently begun to collect browser fingerprints from users based on user agent 
+and JavaScript-acquired browser plugin details, time zone, screen size, color depth, system fonts, 
+keyboard layout and locale information. Subsequent classifiers could leverage this information either 
+as a hash that effectively tags individual known spammers or in part, adding features based on 
+time zone or keyboard layout.
 
 It is also anecdotally clear from moderator log messages that spammers re-use their profile photos and 
-these photos are often of celebrities or simply stock photos. Hashing these images
+these photos are often of celebrities or simply stock photos. Fingerprinting and perceptual hash techniques 
+such as Marr wavelet, discrete cosine transform (DCT), color histograms, etc., which provided by the 
+pHash library, \footnote{http://www.phash.org} seem to promising in detecting image re-use.
 
-Server access logs, search logs and user viewing histories from InterPals, all of which are available to us, offer another avenue for future research.
-It is conceivable that certain spammers may have similar site usage patterns, in terms of HTTP request intervals,
+Server access logs, search logs and user viewing histories from InterPals, all of which 
+are available to us, offer another avenue for future research. It is conceivable that certain 
+spammers may have similar site usage patterns, in terms of HTTP request intervals, search queries, 
+and so on, which might prove useful in their detection.
 
-\subsection{Testing \& Evaluation}
+\subsection{Site Implementation}
 
-- test SVM and other classifiers on newer data
--
+Given the promising results of SVM classifiers in this project, we plan to test and implement a classifier that will 
+be integrated into the site application code itself. Initially, we envision this as a system that automatically flags 
+potential spam messages and brings them to the attention of human moderators. Moderators will then be able to manually 
+make a decision as to whether the messages are spam or not. In order to diminish the effect of subjectivity or moderator 
+error, each message will be vetted by several moderators, and the majority decision will be accepted as ground ``truth." 
+Recording these messages and moderator decision will allow us to further evaluate the performance of these classifiers in 
+the ``wild."
 

### 2012-05-09 docs/paper/paper.tex
update

@@
 Korea, UK, France and Germany combining for 24\%. This distribution of profile countries can 
-be seen in Figure \ref{fig:cprof}
+be seen in Figure \ref{fig:cprof}.
 
@@
 respectively). For spam user accounts, Hotmail, Yahoo and Gmail accounted for 7\%, 72\% and 6\% respectively. This striking 
-predominance of Yahoo email accounts among spammers can be seen in figure \ref{fig:email}. While we have no certain 
+predominance of Yahoo email accounts among spammers can be seen in Figure \ref{fig:email}. While we have no certain 
 explanation for this, Figure \ref{fig:yspam} shows that Ghana and Nigeria alone account for 38\% of the IP 
@@
 
-figure \ref{fig:day}
-figure \ref{fig:month}
-figure \ref{fig:drop}
 \begin{figure}[h]

### 2012-05-09 docs/paper/paper.tex
update

@@
 \usepackage{multirow}
-
+\usepackage{amsmath}
+
+\DeclareMathOperator{\Argmax}{argmax}
+\newcommand{\PosC}{\mathrm{Spam}}
+\newcommand{\NegC}{\mathrm{Ham}}
 \newcommand{\ie}{{\em i.e.,}~}
@@
 
-Social networks of any significant size see a constant stream of spam, scams and phishing attacks. 
+Social networks of any significant size witness a constant flow of spam, scams and phishing attacks. 
 The nature of this unwanted activity can be quite diverse, specific to a site's target audience and 
@@
 
-We investigated several machine learning techniques to automatically 
-detect spam and scams in the private message data, and chose to use 
-train and evaluate Na\"ive Bayes, Linear Regression, and 
-Support Vector Machine (SVM) classifiers. Our implementations used 
-a variety of tools, including Matlab, ScalaNLP, LIBSVM, Lucene and Spark, 
-an in-memory distributed computing framework designed for machine 
-learning and iterative computation.
+We investigate several machine learning techniques to automatically 
+detect spam and scams in the private message data, and train and 
+evaluate Na\"ive Bayes, Linear Regression, and Support Vector Machine (SVM) 
+classifiers. Our implementations used a variety of tools, including Matlab, ScalaNLP, LIBSVM, Lucene and Spark, 
+an in-memory distributed computing framework designed for machine learning and iterative computation.
 
@@
 
-The rapid growth of social media has made Social Networking Services (SNSs) 
-increasingly attractive targets for spam and fraud, leading to a proliferation 
-of sophisticated attacks. This trend is reflected in recent research, as papers 
+The rapid growth of social media has made SNSs increasingly attractive targets 
+for spam and fraud, leading to a proliferation of sophisticated attacks. This 
+trend is reflected in recent research, as papers 
 have focused on identifying and classifying the various types of social media spam. 
@@
 
-For this project, we had unlimited access to data from InterPals, a SNS for users who wish to communicate with others from other countries, whether for language practice, cultural exchange or friendship.
-Users sign up by completing a registration form with information about themselves, including age, sex and location.
-After registering, users can expand their profile page to include descriptions, interests, languages they speak, etc.
-and upload photographs.
-After clicking on a link sent to their email address, a user can begin to interact with others on the site via private message, public ``wall" posts, comments on photos and on a bulletin board system.
+For this project, we had unlimited access to data from InterPals, a SNS for users who 
+wish to communicate with others from other countries, whether for language practice, cultural 
+exchange or friendship. Users sign up by completing a registration form with 
+information about themselves, including age, sex and location. After registering, users 
+can expand their profile page to include descriptions, interests, languages they speak, etc. 
+and upload photographs. After clicking on a link sent to their email address, a user can 
+begin to interact with others on the site via private message, public ``wall" posts, comments 
+on photos and on a bulletin board system.
 
@@
 \item Month of birth 
-\item Whether sender has friends , boolean
-\item Whether sender has photo(s) , boolean
-\item Profile country and IP-detected country match , boolean
-\item IP-detected country is Ghana, Nigeria, Senegal, Malaysia, Turkey, Gambia, Ivory Coast, Togo (``high-risk" countries) , boolean
+\item Whether sender has friends, boolean
+\item Whether sender has photo(s), boolean
+\item Profile country and IP-detected country match, boolean
+\item IP-detected country is a ``high-risk"\footnote{As determined by us for the purposes of this project, based on 
+the IP-detected country histogram of spam users. This list includes Ghana, Nigeria, Senegal, Malaysia, Turkey, 
+Gambia, Ivory Coast and Togo.} spam country, boolean
 \end{itemize}
@@
 
-% Modified from Reynold's HW1  write-up
+The main idea of the Naive Bayes classification algorithm is that one can calculate the probability 
+of an item belonging to a certain class by applying Bayes' theorem on the assumption that the probability 
+of one word/feature that appears in a document is independent of the probability of any other feature. 
+While is very rarely the case (hence ``naive"), Naive Bayes still performs surprisingly well in many cases.
+
+Using the naive Bayes classifier, for a class \(c \in \{\PosC, \NegC\}\) and document \(d\) consisting of words \(w_{d,1},\dotsc,w_{d,m_d}\) (where each word occurs at most once in the Bernoulli version), we have
+\[\Pr(c|d)=\frac{\Pr(d|c)\Pr(c)}{\Pr(d)}\]
+where
+\[\Pr(d|c)=\sum_{i=1}^{m_d}\Pr(w_i|c)\ldotp\]
+The training procedure sets \(\Pr(c)\) to be the fraction of documents with class \(c\),%
+
+\[\Pr(w|c)=\frac{n_{w,c}+\alpha}{\sum_{w'}(n_{w',c}+\alpha)}\ldotp\]
+When testing, our model classifies a document \(d\) as \(\Argmax_c\Pr(c|d)\).
+
 %We use a na\``ive Bayes model, in which the documents are generated according to the following process.  
@@
 
+The Na\"ive Bayes classifier had an accuracy of 82\% when trained on the bag-of-words. Its accuracy dropped 
+to 77\% when expert features were added. The decline in accuracy after adding expert features is most likely 
+explained by the ``naive" assumption in this algorithm that the features are independent. Several of 
+the features that we used, \eg{account lifetime, and whether a user has friends or photos}, are in 
+fact likely to be co-dependent, leading to an overweight of these features during classification.
+
 One of the advantages of the Na\"ive Bayes classifier was that it took very little time to train the classifier, 
 and the steps involved were all trivially parallelizable, which makes it convenient for large data sets. 
-
-The decline in accuracy after adding expert features is most likely explained by the ``naive" assumption in this 
-algorithm that the features are independent. Several of the features that we used, \eg{account lifetime, 
-and whether a user has friends or photos}, are in fact likely to be co-dependent, leading to an overweight of 
-these features during classification. The wealth of expert features available to a SNS, many of which are 
-not independent, makes this the least compelling classifier choice that we examined.
+However, in view of the wealth of expert features available to a SNS, many of which are not independent, this was 
+the least compelling classifier choice that we examined.
 
@@
 over 10-fold cross-validation. The addition of expert features actually decreased this accuracy to 
-51\%. Table \ref{tab:svmfeats} provides a list of top features for the bag-of-words and expert 
+51\%.
+
+Table \ref{tab:svmfeats} provides a list of top features for the bag-of-words and expert 
 feature models. We see that for the bag-of-words classifier, the words ``year," ``new," ``happy,"  
 ``2012," ``wish," and ``you" all have high weights, which makes sense given that the ham data was 
-collected around the beginning of 2012. The table shows that three expert features: account lifetime, 
-friends and recipient age, are weighted very heavily in comparison to the remaining features. These 
+collected around the beginning of 2012. The lists include a number of stopwords, such as ``ok," 
+``me" and ``it," which should ideally not be highly weighted. While we did not remove stopwords for this 
+project, this is something that we could consider doing in the future.
+
+The table of top terms reveals that three expert features: account lifetime, friends and 
+recipient age, are weighted very heavily in comparison to the remaining features. These 
 skewed weights (and the inability to tune the SVM solver via ScalaNLP) prompted us to test LIBSVM 
@@
 wish & 0.0836 & the & -0.01412\\
-doing & -0.0779 & f-sex & -0.01258\\
+doing & -0.0779 & F-sex & -0.01258\\
 
@@
 shows the AUCs and ROC curves for the 50k data sets, making clear the the significantly greater AUC resulting 
-from the addition of expert features.
+from the addition of expert features. As expected, however, both the ScalaNLP and LIBSVM implementations 
+took substantially longer to train than NB and LR.
 
@@
 grouped together without regard to their specific angle of attack, share a number of characteristics that differentiate 
-them from legitimate users. To simulate an approach 
-These features allowed us to improve the accuracy of classifiers
-
-%\begin{figure}[h]
-%    \centering
-%    \includegraphics[width=\linewidth]{figures/sex.pdf}
-%    \caption{Sender sex as stated on profile.}
-%    \label{fig:sendsex}
-%\end{figure}
-%
+them from legitimate users. We trained and evaluated Na\"ive Bayes, Linear Regression and SVM classifiers on a subset 
+of the 4 million messages on which we collected feature statistics. Our evaluation compared the performance of 
+each classifier on bag-of-word representations of message text vis-\`{a}-vis the combination of bag-of-word features 
+with a number of ``expert features." We found that SVMs using an RBF kernel on LIBSVM outperformed the other 
+classifiers. In both the linear regression and SVM cases, the addition of expert features substantially increased 
+the classifier's accuracy, underscoring the impact of supplementary account-related features in detecting spam on 
+SNSs.
 

### 2012-05-09 docs/paper/paper.tex
update

@@
 \begin{itemize}
-\item \textbf{Advance Fee Fraud} We observed this scam to come in various flavors.  (inheritance, lottery,  etc)
-\item \textbf{Romance Scams:} These 
+\item \textbf{Advance Fee Fraud} 
+\item \textbf{Romance Scams} 
 \item \textbf{Sexual Solicitation}
@@
 
-Our goal in classification was not only to build an effective classifier for spam detection, but also to see how the presence of ``expert features", or features based on account and message metadata, could affect classifier accuracy.
-To this end, with of each of the classifiers we chose to implement, we evaluated performance on a bag-of-words representation of message data as well as a combination of bag-of-words with expert features.
+Our goal in classification was not only to build an effective classifier for spam detection, 
+but also to see how the presence of ``expert features", or features based on account and 
+message metadata, could affect classifier accuracy. To this end, with of each of the classifiers 
+we chose to implement, we evaluated performance on a bag-of-words representation of message data 
+as well as a combination of bag-of-words with expert features.
 
@@
 %- used matlab for linear regression with ridge regularization.
-%- also attempted logistics regression with scalanlp, but training result was buggy (always output 0 probably. perhaps a bug in scalanlp)
-%
-%- naive bayes is super fast. accuracy around 80%. when expert features are added, accuracy dropped.
-%
-%- SVM is slow, and doesn’t produce very good result. Suspect we don’t have enough training data (22k training examples, 5k features)
-%
-%- linear regression performed the best, with an AUC of 0.xx. When expert features are added, AUC went up to 0.xx.
-%
 
@@
 All of our linear regression code was implemented in MATLAB because of its ease-of-use 
-and optimizations on matrix multiplications. We read in the text-based feature dictionaries and feature matrix
-representations created above,
-
-%We parse the messages from the text format produced by our cleaning script, using the dictionary generated earlier map words to integers, saving the feature data into two sparse matrices.
-The first matrix only includes the bag-of-words representation of the message body, while the second includes the expert features as well.
-Matlab uses Compressed Sparse Column format, so we actually stored the matrix $X'$ because the dictionary is very sparse.
-
-%For each feature set, the actual matrix multiplications were done using the sparse $X$ matrix along with our $y$ vector and the ridge parameter.
-The multiplication $X'X$ takes up too much memory if we use the entire dictionary, so we chose to limit the dictionary size.
-We found that a dictionary of the 10,000 most frequent words gave sufficient features while also being computable using MATLAB's built-in sparse matrix multiplications.
-We also tried using a dictionary of 15,000 words, but only had a marginal improvement along with much slower execution.
-Using our 10,000 word dictionary, we calculated the performance of our matrix multiply to be 0.123 Gflops/s by dividing the number of non-zero elements in our matrices by the time to compute the multiplication.
-%
-%
-%Since we used ridge regression, we had the ability to tune the $\lambda$ parameter in order to get the best performance.
-As shown in figure \ref{fig:lambda}, we observed an AUC improvement of about .5\% between the default value of 1 and a tuned parameter of 1500.
-This is a fairly major difference since the AUC is at 93\% using the default value.
-Additionally, we also noticed a major difference in the top words when using better $\lambda$ values.
-Using the default value of 1, our top words included obscure words including ``\_c'', ``\_j'', ``donovan'', and ``matlock.'' When we used the tuned $\lambda$ value of 1500, the words were much more representative, as shown in table \ref{tab:words}.
-%
+and optimizations on matrix multiplications. We read in the text-based feature dictionaries and the feature matrix
+representations created above, which is stored as a sparse matrix (efficiently represented in Matlab as 
+in Cmmpressed Sparse Column format).
+
+We used ridge regression, which gave us the opportunity to tune a $\lambda$ parameter in order to 
+regularize the vector components for better performance.
 
@@
 
-We tried a number of values for $\lambda$: 1, 10, 50, 100, 500, 1000, 1500, 3000. Figure \ref{fig:lambda}
-\ref{fig:roclin}
+We performed 10-fold cross-validation with $\lambda=50$, which resulted in a mean AUC of 0.886 and mean lift of 
+29.216 for the bag-of-words LR model. After adding the expert features, the mean AUC climbed to 0.951 with a mean 
+lift of 41.584. Figure \ref{fig:roclin} show ROC curves for 1-fold cross-validation on both of these models.
+
+As shown in figure \ref{fig:lambda}, we tested a number of values for $\lambda$, the ridge regression parameter. 
+Since the AUCs corresponding to $\lambda$ for both models appear to peak between 30 and 50, we may have been able to 
+get a slightly higher AUC by using a lower $\lambda$.
 
@@
 \end{figure}
+
+Tables \ref{tab:lrfeats} and \ref{tab:lrfeatse} list the top-weighted terms for the bag-of-words and expert feature 
+models. The spam-related words appear in line with our expectations, with sex-related terms prevalent, as well as
+words that might indicate advertisements (``cafe4tune," for example, appears to be a competing SNS). Table \ref{lrfeatse} 
+shows that IP-profile country mismatch and membership of the IP-detected country in our ``high-risk" list were 
+strong predictors that a message might be spam. Likewise, whether or not a user had photos was a righ-ranking indicator 
+of ham.
 
@@
 \hline
-
 requests & -0.1349 & F-country\_match & 0.3365\\
@@
 daichi & -0.0775 & k?  & 0.0806\\
-
 \end{tabular}
 \caption{Linear Regression top features (bag-of-words \& expert features) with weights. Expert features
-are denoted by an 'F-' prefix.}
-\label{tab:lrfeats}
+are denoted by an `F-' prefix.}
+\label{tab:lrfeatse}
 \end{table}

### 2012-05-09 docs/paper/paper.tex
update

@@
 
-Support Vector Machines are a classification method \cite{boser}, \cite{cortesv95}
-
-We initially used ScalaNLP's built-in SVM solver, which implements the Pegasos algorithm \cite{pegasos} as extended 
-by Wang, Crammer and Vucetic\cite{wang2}. The optimizer runs stochastic subgradient descent on the primal objective 
+Support Vector Machines are a classification method that maps class examples (\eg messages) to points 
+in space and aims to maximize the margin around a hyperplane separating the classes.\cite{boser}, \cite{cortesv95} 
+Because the sets of points may not be linearly separable in the original dimension, a kernel trick can 
+be used to fit the maximum-margin hyperplane in a higher-dimensional feature space (which may appear 
+non-linear in the original input space).
+
+We initially used ScalaNLP's built-in SVM solver, which implements the Pegasos maximization algorithm\cite{pegasos} 
+as extended by Wang, Crammer and Vucetic\cite{wang2}. 
+The optimizer runs stochastic subgradient descent on the primal objective 
 using batches provided. However, the ScalaNLP SVM interface does not allow changing kernel functions or parameters, 
@@
 Python and other languages. The package allows one to specify the type of kernel to use, kernel parameters 
-($\gamma$, $\rho$, $d$), as well as a penalty parameter ($C$). We used Perl to convert the sparse feature 
+($\gamma$, $\rho$, $d$), as well as a soft margin parameter ($C$). We used Perl to convert the sparse feature 
 matrices generated for use with ScalaNLP to a format recognized by LIBSVM. We then used one of the package's 
@@
 package includes a Python script that performs cross-validation to suggest appropriate 
-$\gamma$ and $C$ values for the recommended radial basis function (RBF) kernel. The RBF kernel is 
+$\gamma$ and $C$ values for the recommended Gaussian radial basis function (RBF) kernel. The RBF kernel is 
 a real-valued function such that: 
 $$K(\mathbf{x}_i, \mathbf{x}_j) = (\gamma\mathbf{x}_i^\intercal\mathbf{x}_j+r)^d, \gamma>0$$
-We ran this script on a subset of 2500 ham and 2500 spam messages to obtain $C = 8$ and $\gamma = 0.0078125$.
+We ran this script on a subset of 5000 messages with expert features to obtain $C = 8$ and $\gamma = 0.0078125$, 
+which we used to tune the SVM we used to classify both the bag-of-words and expert feature models.
 

### 2012-05-09 docs/paper/paper.tex
update

@@
 Social networks of any significant size witness a constant flow of spam, scams and phishing attacks. 
-The nature of this unwanted activity can be quite diverse, specific to a site's target audience and 
-often not trivial to detect. Marketers can spam members with unwanted advertisements, fraudsters 
-lure users with advance fee frauds and other confidence tricks, while others may attempt to steal user
-information by directing users to external phishing pages. In contrast to email spam, social spam 
-has a more personal component, since both spammers and legitimate users (``ham" users) have accompanying 
-profiles with descriptive information.
+The nature of this unwanted activity, which we henceforth refer to collectively as ``spam" can be 
+quite diverse, specific to a site's target audience and often not trivial to detect. Marketers 
+spam members with unwanted advertisements, fraudsters lure users with advance fee frauds and 
+other confidence tricks, while others may attempt to steal user information by directing users 
+to external phishing pages.
 
 Sites with global reach see communication among its members in a number foreign languages with 
-varying levels of ability. This means that much benign content shares characteristics like misspellings, 
-awkward phrases, etc. that might have made certain types of common frauds and spam more easy to 
-distinguish on US-based (or English-language) sites.
-
-Recent research has begun to evaluate Email spam and techniques with which to combat it has been studied 
-extensively. Like
-
-% focus on extra impact of non-public features vs. public (bag-of-words)
+varying levels of ability. This means that much benign content shares characteristics including 
+misspellings, awkward phrases, etc. that might have made certain types of common frauds and spam 
+more easy to distinguish on US-based (or English-language) sites. Likewise, simple regional IP-based 
+filtering to target high-spam countries like Nigeria and Ghana would prevent legitimate users 
+located on these networks from accessing the site.
+
+In contrast to email spam, social spam has a more personal component, since both spammers 
+and legitimate users (``ham" users) have accompanying profiles with descriptive information. This can 
+lead to more drawn-out, conversational attempts by spammers to approach users, since communication 
+is more immediate (in the sense that users may see that they are online at the same time, may 
+visit each others' profiles, etc.). However, the additional user data inherent in SNSs also 
+offers a bountiful supply of data that site operators can mine to detect spam more effectively.
+
+Most research on social networking spam has been done at a distance, using data collected 
+either through scraping or artificially attracting spammers through ``honeypot" accounts. This 
+paper leverages access to private messages, metadata and account details from a popular SNS 
+to study the characteristics of social spam as well as the features and classifiers that 
+sites can use to detect it.
 

### 2012-05-09 docs/paper/paper.tex
update

@@
 
-This paper examines in detail the types of malicious and benign
-content that are encountered on social networks by analyzing
-experimental data available from InterPals, an international social network
+We examine in detail the types of malicious and benign
+content that users encounter on social networks. We do this by analyzing
+data available from InterPals, an international social network
 for cultural exchange and language practice. For example, the site
@@
 on the site, including those measures that have already been taken
-(e.g. various heuristics including IP/location anomaly detection,
+(\eg various heuristics including IP-based location anomaly detection,
 frequency capping, duplicate account detection, etc.). We then analyze 
-message and user account data points to try to differentiate legitimate users
-from malicious ones. By mining this data, we extract features 
-to build and evaluate classifiers that can detect unwanted behavior 
+message and user account data to try to identify characteristics that 
+best differentiate legitimate users from malicious ones. By mining this data, 
+ we extract features to build and evaluate classifiers that can detect unwanted behavior 
 programmatically. The large volume of data available to us, although we do 
@@
 
-We investigate several machine learning techniques to automatically 
-detect spam and scams in the private message data, and train and 
-evaluate Na\"ive Bayes, Linear Regression, and Support Vector Machine (SVM) 
-classifiers. Our implementations used a variety of tools, including Matlab, ScalaNLP, LIBSVM, Lucene and Spark, 
-an in-memory distributed computing framework designed for machine learning and iterative computation.
+This paper investigate several machine learning techniques to 
+detect spam in private messages. We train and evaluate Na\"ive Bayes, 
+linear regression, and support vector machine (SVM) 
+classifiers. Our implementations used a variety of tools, 
+including Matlab, ScalaNLP, LIBSVM, Lucene and Spark, 
+an in-memory distributed computing framework designed for machine 
+learning and iterative computation.
 
@@
 
-We initially trained and evaluated three classifiers on a small subset of the data: Na\"ive Bayes, Linear Regression and SVM. 
+We initially trained and evaluated three classifiers on a small subset of the data: Na\"ive Bayes, linear regression and SVM. 
 We chose an 11k spam and 11k ham sample for training, using this particular sample size due to the size of the spam corpus 
@@
 \end{tabular}
-\caption{Linear Regression top features (bag-of-words) with weights}
+\caption{linear regression top features (bag-of-words) with weights}
 \label{tab:lrfeats}
@@
 \end{tabular}
-\caption{Linear Regression top features (bag-of-words \& expert features) with weights. Expert features
+\caption{linear regression top features (bag-of-words \& expert features) with weights. Expert features
 are denoted by an `F-' prefix.}
@@
 7 months, with only a slight linear increase in volume over the interval.  Consequently, as seen in 
-the Linear Regression and SVM feature weight tables, words typically associated with the New Year gained 
+the linear regression and SVM feature weight tables, words typically associated with the New Year gained 
 undue wieght during classification.  We plan to address this issue by creating a new ham message corpus 
@@
 grouped together without regard to their specific angle of attack, share a number of characteristics that differentiate 
-them from legitimate users. We trained and evaluated Na\"ive Bayes, Linear Regression and SVM classifiers on a subset 
+them from legitimate users. We trained and evaluated Na\"ive Bayes, linear regression and SVM classifiers on a subset 
 of the 4 million messages on which we collected feature statistics. Our evaluation compared the performance of 

### 2012-05-09 docs/paper/paper.tex
update

@@
 \section{Introduction}
-% Modified from project proposal
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Related Work}
@@
 and MySpace. Other work looks at SNSs as platforms to collect data about users in 
-order to aid direct attacks on the user's computers or to compromise a large number of accounts. \cite{patsakis, huber}
+order to aid direct attacks on the user's computers or to compromise a large 
+number of accounts. \cite{patsakis, huber}
 
@@
 Unfortunately, the differing characteristics of various social networks \eg the length 
-of messages in Facebook vs. Twitter, can reduce the benefits of sharing spam corpora across diverse sites.
+of messages in Facebook vs. Twitter, can reduce the benefits of sharing spam corpora 
+across diverse sites.
 
@@
 include significant detail on the various types and characteristics of undesirable 
-activity on the site, including fake profiles, harassment, compromised accounts, malware and spam.
+activity on the site, including fake profiles, harassment, compromised accounts, malware 
+and spam.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Data set}
@@
 
-Currently, the site combats spam and other Terms of Service violations through volunteer moderators.
-Users can report content, including private messages, profiles and photographs to moderators, using a form that includes a drop-down menu of pre-selected reasons with the option to add a more detailed message in a text field.
-Likewise, the message interface allows a user to report a private message spam with a single click.
-Moderators have access to a queue of these reports.
-In addition to the material being reported, moderators are able to make decisions based on the data from the reported user's account, including outgoing private messages, IP addresses, as well as a list of other users who have logged in from the same computer (determined via IP address as well as by browser fingerprints).
-Moderators can then decide to delete the user, send a warning, or clear reports on a user.
-All moderator actions require them to annotate their decision with a brief log message.
-When deleting a user, moderators have the option of flagging the reported message (if there is one) as spam, or of flagging all of the user's outgoing private messages as spam.
-
-Other anti-spam measures include widespread use of CAPTCHAs across the site and frequency caps.
-Short-window frequency caps are in place for all users, limiting the number of messages that can be sent per short time interval (1, 5 and 10 minutes).
-New users are also subject to a per-day cap on the number of unique users with whom they are able to initiate contact.
-
-Users are able to filter messages by age, sex, continent and country. At this time,
+Currently, the site combats spam and other Terms of Service violations 
+through volunteer moderators. Users can report content, including private messages, 
+profiles and photographs to moderators, using a form that includes a drop-down menu of 
+pre-selected reasons with the option to add a more detailed message in a text field. 
+Likewise, the message interface allows a user to report a private message spam with a single 
+click. Moderators have access to a queue of these reports. In addition to the material 
+being reported, moderators are able to make decisions based on the data from the reported 
+user's account, including outgoing private messages, IP addresses, as well as a list of 
+other users who have logged in from the same computer (determined via IP address as well 
+as by browser fingerprints). Moderators can then decide to delete the user, send a warning, 
+or clear reports on a user. All moderator actions require them to annotate their decision with a 
+brief log message. When deleting a user, moderators have the option of flagging the 
+reported message (if there is one) as spam, or of flagging all of the user's outgoing private messages 
+as spam.
+
+Other anti-spam measures include widespread use of CAPTCHAs across the site and 
+frequency caps. Short-window frequency caps are in place for all users, limiting the 
+number of messages that can be sent per short time interval (1, 5 and 10 minutes). New users 
+are also subject to a per-day cap on the number of unique users with whom they are able to initiate 
+contact.
+
+%- Users are able to filter messages by age, sex, continent and country. At this time, 
+%- Duplicate account detection
 
@@
 
-At the beginning of this research project, slightly two million messages had been flagged by moderators as spam.
-We extracted the contents of exactly two million spam-labeled messages from accounts deleted by moderators between October 2011 and March 2012.
-As moderators can flag a user's entire list of outgoing messages as spam on deletion, the earliest sent dates of some messages in our data set begin in May 2010.
+At the beginning of this research project, slightly two million messages had 
+been flagged by moderators as spam. We extracted the contents of exactly 2 million 
+spam-labeled messages from accounts deleted by moderators between October 2011 
+and March 2012. As moderators can flag a user's entire list of outgoing messages 
+as spam on deletion, the earliest sent dates of some messages in our data set begin in 
+May 2010.
 
 InterPals stores private messages and user account data in a number of separate MySQL tables. 
-Account information for deleted users is stored for six months, so account details for all spam and ham accounts was available.
-Extracting the data of interest to this project required dumping the contents from a query with multiple joins and aggregations across 8 tables with a combined size of 154 GB.
-To minimize the impact on the load of the site's production database instances, we took an LVM snapshot of slave instance and cloned the database to a spare server.
-In addition to using SQL to extract the data, we wrote Perl scripts to clean the data, tokenize messages and prepare data for further processing.
-The methodology section (X) offers more details on precisely which items of data were collected and why.
+Account information for deleted users is stored for six months, so account details for all spam 
+and ham accounts was available. Extracting the data of interest to this project required 
+dumping the contents from a query with multiple joins and aggregations across 8 tables with a 
+combined size of 154 GB. To minimize the impact on the load of the site's production database 
+instances, we took an LVM snapshot of slave instance and cloned the database to a spare server. 
+In addition to using SQL to extract the data, we wrote Perl scripts to clean the data, tokenize 
+messages and prepare data for further processing. The methodology section (X) offers more details 
+on precisely which items of data were collected and why.
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Methodology}
@@
 
-In order to train our classifiers, we created a feature matrix from the message and account data. We did this using 
-Perl, Spark, and Lucene analyzers. The first step was to generate a dictionary sorted by document frequencies in descending 
-order, mapping each feature (word or expert feature value) to an integer key. The second step was to create a sparse 
-matrix representation of each document (in this case, spam or ham message) and the associated features. Spark provides a 
-parallelized in-memory computation framework that dramatically reduced the time necessary to extract these features, 
-build the dictionary and generate the feature matrix.
-
-We did not use all of the expert features available to us, due to time constraints. Likewise, for many expert features could 
-have a wide range of values, which would expand the dictionary size considerably. In addition to the bag-of-words from the 
-message text, we used the following features:
+In order to train our classifiers, we created a feature matrix from the message and account data. 
+We did this using Perl, Spark, and Lucene analyzers. The first step was to generate a dictionary sorted by document 
+frequencies in descending order, mapping each feature (word or expert feature value) to an integer 
+key. The second step was to create a sparse matrix representation of each document (in this case, 
+spam or ham message) and the associated features. Spark provides a parallelized in-memory computation 
+framework that dramatically reduced the time necessary to extract these features, build the dictionary 
+and generate the feature matrix.
+
+We did not use all of the expert features available to us, due to time constraints. Likewise, for 
+many expert features could have a wide range of values, which would expand the dictionary size 
+considerably. In addition to the bag-of-words from the message text, we used the following features:
 \begin{itemize}
@@
 
-We initially trained and evaluated three classifiers on a small subset of the data: Na\"ive Bayes, linear regression and SVM. 
-We chose an 11k spam and 11k ham sample for training, using this particular sample size due to the size of the spam corpus 
+We initially trained and evaluated three classifiers on a small subset of the data: 
+Na\"ive Bayes, linear regression and SVM.  We chose an 11k spam and 11k ham sample for 
+training, using this particular sample size due to the size of the spam corpus 
 section containing recipient age feature data.
-
-%classification:
-%- implemented 3 models: naive bayes, linear regression, and SVM.
-%- used scalanlp for naive bayes
-%- used matlab for linear regression with ridge regularization.
 
@@
 When testing, our model classifies a document \(d\) as \(\Argmax_c\Pr(c|d)\).
-
-%We use a na\``ive Bayes model, in which the documents are generated according to the following process.  
-%First, a class \(c\) is chosen from a probability distribution \(\Pr(c)\).
-%Then, a document from class \(c\) is generated.
-%We tried a multinomial model and a Bernoulli model to model this process.
-%\begin{itemize}
-%\item \emph{Multinomial model.}  A document length \(m_d\) is chosen and \(m_d\) words \(w_{d,1}, \dotsc, w_{d,m_d}\) are chosen independently from a distribution \(\Pr(w|c)\).
-%\item \emph{Bernoulli model.}  For each word \(w\) in the vocabulary, \(w\) is added to the document with probability \(\Pr(w|c)\).
-%\end{itemize}
-%The distributions \(\Pr(c)\) and \(\Pr(w|c)\) are parameters of the model.
-%
-%
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Future Work}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}

### 2012-05-09 docs/paper/paper.tex
typos

@@
 
-Facebook \cite{stein} provides an overview of their ``immune system" defences 
+Facebook \cite{stein} provides an overview of their ``immune system" defenses 
 against phishing, fraud and spam. The system is composed of classifier services, an 
@@
 features for classification and a policy engine to take action on suspected misuse. While 
-the discussion remains high-level and includes few implemention particulars, it does 
+the discussion remains high-level and includes few implementation particulars, it does 
 include significant detail on the various types and characteristics of undesirable 
@@
 have applied machine learning \cite{kolari, nag} and NLP \cite{mishne} techniques to the 
-problem. Likewise, Markines et al. \cite{markines} apply similar techniques to spam on social bookmarking sites.
+problem. Likewise, Markines et al. \cite{markines} apply similar techniques to spam on 
+social bookmarking sites.
 
@@
 Figure \ref{cite:drop}, it seems likely that this is due to an unwillingness on the part of spam users to 
-scroll down to lower options (as well as, perhaps, a hesitance to disclose real birthdates). A similar trend is 
+scroll down to lower options (as well as, perhaps, a hesitance to disclose real birth dates). A similar trend is 
 observed in the countries with the most discrepancies between IP-detected countries and countries stated on spam 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Classification \& Results}
@@
 $y$, and explanatory variables, $X$. The output variable is a linear sum of the explanatory 
-variables multiplied by their corresponding coeffecients, represented by $\beta$. 
+variables multiplied by their corresponding coefficients, represented by $\beta$. 
 In our case, $y$ represents a message's label as spam or ham (1 or 0), 
@@
 In this project, we performed ordinary least squares regression. By minimizing the sum of squared 
-residuals, we can solve for the unkown parameter vector $\beta$ using a closed form solution. 
+residuals, we can solve for the unknown parameter vector $\beta$ using a closed form solution. 
 $$\beta=(X'X)^{-1}X'y$$ In order to ensure that the matrix $X'X$ is invertible, we perform ridge regression. 
@@
 representations created above, which is stored as a sparse matrix (efficiently represented in Matlab as 
-in Cmmpressed Sparse Column format).
+in Compressed Sparse Column format).
 
@@
 ``2012," ``wish," and ``you" all have high weights, which makes sense given that the ham data was 
-collected around the beginning of 2012. The lists include a number of stopwords, such as ``ok," 
-``me" and ``it," which should ideally not be highly weighted. While we did not remove stopwords for this 
+collected around the beginning of 2012. The lists include a number of stop words, such as ``ok," 
+``me" and ``it," which should ideally not be highly weighted. While we did not remove stop words for this 
 project, this is something that we could consider doing in the future.
@@
 the linear regression and SVM feature weight tables, words typically associated with the New Year gained 
-undue wieght during classification.  We plan to address this issue by creating a new ham message corpus 
+undue weight during classification.  We plan to address this issue by creating a new ham message corpus 
 through sampling that matches the temporal distribution of spam messages more closely.

### 2012-05-09 docs/paper/paper.tex
update

@@
 Social networking sites (SNSs) see a variety of spam and scams targeted at 
-their users. In contrast to the more limited amounts of secondary information 
-available when analyzing email spam, spam on SNSs is accompanied by a wealth 
-of data on the sender, which can be used to build more accurate detection mechanisms. 
+their users. In contrast to the limited amounts of information available beyond 
+message text and headers when analyzing email spam, spam on SNSs is often accompanied 
+by a wealth of data on the sender, which can be used to build more accurate detection mechanisms. 
 We analyze 4 million private messages as well as other public and private data from 
@@
 The nature of this unwanted activity, which we henceforth refer to collectively as ``spam" can be 
-quite diverse, specific to a site's target audience and often not trivial to detect. Marketers 
+quite diverse, specific to a site's target audience and often not easily detectable. Marketers 
 spam members with unwanted advertisements, fraudsters lure users with advance fee frauds and 
@@
 
-\subsection{Spam Set}
+\subsection{Spam Data Set}
 
@@
 
-\subsection{Ham Set}
+\subsection{Ham Data Set}
 
@@
 
-\section{Classification \& Results}
+\section{Classifiers \& Results}
 

### 2012-05-09 docs/paper/paper.tex
update

@@
 
-Social networks of any significant size witness a constant flow of spam, scams and phishing attacks. 
-The nature of this unwanted activity, which we henceforth refer to collectively as ``spam" can be 
+Social networking sites (SNSs) of any significant size witness a constant flow of spam, 
+scams and phishing attacks.  The nature of this unwanted activity, which we henceforth refer to 
+collectively as ``spam" can be 
 quite diverse, specific to a site's target audience and often not easily detectable. Marketers 
@@
 
-Sites with global reach see communication among its members in a number foreign languages with 
+Sites with global reach see communication among members in a variety foreign languages with 
 varying levels of ability. This means that much benign content shares characteristics including 
-misspellings, awkward phrases, etc. that might have made certain types of common frauds and spam 
+misspellings, awkward phrases, and so on, that might have made certain types of common frauds and spam 
 more easy to distinguish on US-based (or English-language) sites. Likewise, simple regional IP-based 
 filtering to target high-spam countries like Nigeria and Ghana would prevent legitimate users 
-located on these networks from accessing the site.
+located on blocked networks from accessing the site.
 
@@
 
-\subsection{Spam Email \& Web}
+\subsection{Email \& Web Spam}
 

### 2012-05-09 docs/paper/paper.tex
update

@@
 data available from InterPals, an international social network
-for cultural exchange and language practice. For example, the site
+for cultural exchange and language practice. The site
 attracts a wide variety of financial scams, ranging from Nigerian
@@
 question & -0.1095 & displaying & 0.1112\\
-xxx & -0.1081 & ghana & 0.111\\
+xxx & -0.1081 & ghana & 0.1110\\
 joyce & -0.1078 & correctly & 0.1102\\
@@
 \end{tabular}
-\caption{linear regression top features (bag-of-words) with weights}
+\caption{Linear regression top features (bag-of-words) with weights}
 \label{tab:lrfeats}

### 2012-05-09 docs/paper/paper.tex
update

@@
     \includegraphics[width=\linewidth]{figures/country-prof.pdf}
-    \caption{Country as stated on user profile.}
+    \caption{Country as stated on user profile}
     \label{fig:cprof}
@@
     \includegraphics[width=\linewidth]{figures/country-ip.pdf}
-    \caption{Country detected by IP address using MaxMind GeoIP database.}
+    \caption{Country detected by IP address using MaxMind GeoIP database}
     \label{fig:cip}
@@
     \includegraphics[width=\linewidth]{figures/ips-usa-spam.pdf}
-    \caption{Distribution of IPs by country for spammers with profiles stating a USA location.}
+    \caption{Distribution of IPs by country for spammers with profiles stating a USA location}
     \label{fig:usaspam}
@@
     \includegraphics[width=\linewidth]{figures/ips-usa-ham.pdf}
-    \caption{Distribution of IPs by country for ham users with profiles stating a USA location.}
+    \caption{Distribution of IPs by country for ham users with profiles stating a USA location}
     \label{fig:usaham}
@@
     \includegraphics[width=\linewidth]{figures/unique-ips.pdf}
-    \caption{Number of unique IP addresses associated with user account.}
+    \caption{Number of unique IP addresses associated with user account}
     \label{fig:uniqip}
@@
     \includegraphics[width=\linewidth]{figures/lifetime.pdf}
-    \caption{Sender account lifetime at the time that the message was sent.}
+    \caption{Sender account lifetime at the time that the message was sent}
     \label{fig:lifetime}
@@
     \includegraphics[width=\linewidth]{figures/email.pdf}
-    \caption{Domain name of email provided at signup.}
+    \caption{Domain name of email provided at signup}
     \label{fig:email}
@@
     \includegraphics[width=\linewidth]{figures/yahoo-spam.pdf}
-    \caption{IP distribution for scam accounts with Yahoo email addresses.}
+    \caption{IP distribution for scam accounts with Yahoo email addresses}
     \label{fig:yspam}
@@
     \includegraphics[width=\linewidth]{figures/recip-age.pdf}
-    \caption{Age of message recipient as stated on user profile. Note that 2 million message spam sample 
-        had only 11,000 messages with this information.}
+    \caption{Age of message recipient as stated on user profile (Note that 2 million message spam sample 
+        had only 11,000 messages with this information)}
     \label{fig:recipage}
@@
     \includegraphics[width=\linewidth]{figures/dob-day.pdf}
-    \caption{Date of month from birthday submitted by user at signup.}
+    \caption{Date of month from birthday submitted by user at signup}
     \label{fig:day}
@@
     \includegraphics[width=\linewidth]{figures/dob-month.pdf}
-    \caption{Month of birthday as submitted by user at signup.}
+    \caption{Month of birthday as submitted by user at signup}
     \label{fig:month}
@@
     \includegraphics[width=\linewidth]{figures/dropdown.png}
-    \caption{Drop-down select menu on signup form.}
+    \caption{Drop-down select menu on signup form}
     \label{fig:drop}
@@
     \includegraphics[width=\linewidth]{figures/photos.pdf}
-    \caption{Number of photos associated with sender account.}
+    \caption{Number of photos associated with sender account}
     \label{fig:photos}
@@
     \includegraphics[width=\linewidth]{figures/friends.pdf}
-    \caption{Number of friends associated with sender account.}
+    \caption{Number of friends associated with sender account}
     \label{fig:friends}
@@
     \includegraphics[width=\linewidth]{figures/linear-ridge.pdf}
-    \caption{AUC based on varying lambda values for Ridge regression.}
+    \caption{AUC based on varying lambda values for Ridge regression}
     \label{fig:lambda}
@@
     \includegraphics[width=\linewidth]{figures/linear-roc.pdf}
-    \caption{Linear regression ROC curves for classifier using bag-of-words and bag-of-words with ``expert" features.}
+    \caption{Linear regression ROC curves for classifier using bag-of-words and bag-of-words with ``expert" features}
     \label{fig:roclin}
@@
 \end{tabular}
-\caption{linear regression top features (bag-of-words \& expert features) with weights. Expert features
-are denoted by an `F-' prefix.}
+\caption{Linear regression top features (bag-of-words \& expert features) with weights (Expert features
+are denoted by an `F-' prefix)}
 \label{tab:lrfeatse}

### 2012-05-09 docs/paper/paper.tex
update

@@
 This paper investigate several machine learning techniques to 
-detect spam in private messages. We train and evaluate Na\"ive Bayes, 
+detect spam in private messages. We train and evaluate Naive Bayes, 
 linear regression, and support vector machine (SVM) 
@@
 on MySpace are enough to train an effective social spam classifier. They note that
-C4.5 decision tree algorithms provide better performance than Na\"ive Bayes in this case. 
+C4.5 decision tree algorithms provide better performance than Naive Bayes in this case. 
 As in other works, this only examines publicly available profile information collected 
@@
 researchers tackle this adversarial problem. Dalvi et al. \cite{dalvi} offer a 
-modified Na\"ive Bayes classifier to detect and reclassify data taking into account 
+modified Naive Bayes classifier to detect and reclassify data taking into account 
 the optimal modification strategy that an adversary could choose. Lowd and Meek \cite{lowd} 
@@
 
-\section{Data set}
+\section{Data Sets}
 
@@
 we chose to implement, we evaluated performance on a bag-of-words representation of message data 
-as well as a combination of bag-of-words with expert features.
+as well as a combination of bag-of-words with expert features. Based on our investigation of 
+previous research in this area, we focused our attention on Naive Bayes, linear regression, 
+logistic regression and support vector machines, all of which have been used with varying degrees of 
+success for email, web and social spam classification, as described in our overview of related work.
 
@@
 We initially trained and evaluated three classifiers on a small subset of the data: 
-Na\"ive Bayes, linear regression and SVM.  We chose an 11k spam and 11k ham sample for 
+Naive Bayes, linear regression and SVM.  We chose an 11k spam and 11k ham sample for 
 training, using this particular sample size due to the size of the spam corpus 
@@
 
-\subsection{Na\"ive Bayes}
+\subsection{Naive Bayes}
 
@@
 
-\subsection{Na\"ive Bayes Results}
-
-The Na\"ive Bayes classifier had an accuracy of 82\% when trained on the bag-of-words. Its accuracy dropped 
+\subsection{Naive Bayes Results}
+
+The Naive Bayes classifier had an accuracy of 82\% when trained on the bag-of-words. Its accuracy dropped 
 to 77\% when expert features were added. The decline in accuracy after adding expert features is most likely 
@@
 
-One of the advantages of the Na\"ive Bayes classifier was that it took very little time to train the classifier, 
+One of the advantages of the Naive Bayes classifier was that it took very little time to train the classifier, 
 and the steps involved were all trivially parallelizable, which makes it convenient for large data sets. 
@@
 grouped together without regard to their specific angle of attack, share a number of characteristics that differentiate 
-them from legitimate users. We trained and evaluated Na\"ive Bayes, linear regression and SVM classifiers on a subset 
+them from legitimate users. We trained and evaluated Naive Bayes, linear regression and SVM classifiers on a subset 
 of the 4 million messages on which we collected feature statistics. Our evaluation compared the performance of 

### 2012-05-09 docs/paper/paper.tex
display changes

@@
 \usepackage{amsmath}
+\usepackage[hang,small,bf]{caption}
+\DeclareCaptionType{copyrightbox}
 
@@
 \newcommand{\eg}{{\em e.g.,}~}
+
+\renewcommand{\normalsize}{\fontsize{10.5pt}{12.5pt}\selectfont}
 
@@
 
-In contrast to email spam, social spam has a more personal component, since both spammers 
+In contrast to email spam, social spam often has a more personal component, since both spammers 
 and legitimate users (``ham" users) have accompanying profiles with descriptive information. This can 
@@
 \begin{table}
+\small
 \begin{tabular}{|l|l|r|r|r|}
@@
 \begin{table}
+\small
 \begin{tabular}{l|l||l|l}
@@
 \begin{table}
+\small
 \begin{tabular}{l|r||l|r}
@@
 \begin{table}
+\small
 \begin{tabular}{r|r|r}

### 2012-05-09 docs/paper/paper.tex
update

@@
 
-We examine in detail the types of malicious and benign
+We examine in detail the classes of malicious and benign
 content that users encounter on social networks. We do this by analyzing
@@
 Figures \ref{fig:day} and \ref{fig:month}. Given the drop-down select menus on the web site's registration form, reproduced 
-Figure \ref{cite:drop}, it seems likely that this is due to an unwillingness on the part of spam users to 
+Figure \ref{fig:drop}, it seems likely that this is due to an unwillingness on the part of spam users to 
 scroll down to lower options (as well as, perhaps, a hesitance to disclose real birth dates). A similar trend is 
@@
 
-\textbf{Photos and Friends:} We found that both the number of ``friends" that a user has (connections 
+\textbf{Photos \& Friends:} We found that both the number of ``friends" that a user has (connections 
 to other users) as well as the number of photos that user has uploaded were also indicative of their 
@@
 models. The spam-related words appear in line with our expectations, with sex-related terms prevalent, as well as
-words that might indicate advertisements (``cafe4tune," for example, appears to be a competing SNS). Table \ref{lrfeatse} 
+words that might indicate advertisements (``cafe4tune," for example, appears to be a competing SNS). Table \ref{tab:lrfeatse} 
 shows that IP-profile country mismatch and membership of the IP-detected country in our ``high-risk" list were 

### 2012-05-09 docs/paper/paper.tex
update

@@
 We enjoyed unrestricted access to the data of InterPals.net, a SNS with 
-over 1.3 million active members. This data includes a corpus of over a 
+over 1.3 million active members. The data from this site includes a corpus of over a 
 100 million private messages and another 2 million messages that have been 
-labeled as spam by users. Other data includes 40 million or so ``wall" 
-comments, 5 million photos, and 8 million photo comments.
+labeled as spam  and deleted by users. Other data includes 40 million or so ``wall" 
+comments, 5 million photos and 8 million photo comments.
 
@@
 addresses associated with these Yahoo accounts. This popularity of Yahoo among West African users has 
-been noted before\cite{burrell}, and appears to be due to the early penetration of Yahoo services in the region.
+been noted before\cite{burrell} and appears to be due to the early penetration of Yahoo services in the region.
 
@@
 In order to train our classifiers, we created a feature matrix from the message and account data. 
-We did this using Perl, Spark, and Lucene analyzers. The first step was to generate a dictionary sorted by document 
+We did this using Perl, Spark and Lucene analyzers. The first step was to generate a dictionary sorted by document 
 frequencies in descending order, mapping each feature (word or expert feature value) to an integer 
@@
 explained by the ``naive" assumption in this algorithm that the features are independent. Several of 
-the features that we used, \eg{account lifetime, and whether a user has friends or photos}, are in 
+the features that we used, \eg{account lifetime and whether a user has friends or photos}, are in 
 fact likely to be co-dependent, leading to an overweight of these features during classification.

### 2012-05-09 docs/paper/paper.tex
update

@@
 Nevertheless, most research has been carried out on publicly-available data 
-from SNSs, making it difficult up until now to measure the effect of private user 
+from SNSs, making it unfeasible up until now to measure the effect of private user 
 data on algorithms for detecting site misuse.
@@
 sent from compromised accounts. Benevenuto et al. \cite{benevenuto} identify 
-attributes of spam on video SNSs and use a Support Vector Machine (SVM) for classification.
+social attributes of spam and ham on video SNSs (in this case, YouTube), including video 
+view counts, comment counts and user public profile attributes. They then use a 
+support vector machine (SVM) for classification, yielding 96\% accuracy in detecting 
+advertisers (``promoters"), but only accurately identifying 57\% of examples of more general spam.
 

### 2012-05-09 docs/paper/paper.tex
update

@@
 features from MySpace profiles, but the relatively poor performance of their classifier highlights 
-the difficulties in manual classification social network spam. Several 
-studies take the approach of baiting spammers with social ``honeypots", profiles 
-created with the sole intent of attracting spam.\cite{stringhini, lee} They 
+the difficulties in manual classification of social network spam. Several 
+studies \cite{stringhini, lee} take the approach of baiting spammers with social ``honeypots", profiles 
+created with the sole intent of attracting spam. They 
 then use the data collected to train classifiers with features including 
@@
 the honeypot approach as well and provide examples of various types of spammers, 
-the typical demographics of their profiles as well as the web pages that they tend to advertise.
+the typical demographics of their profiles, as well as the web pages that they tend to advertise.
 
@@
 sent from compromised accounts. Benevenuto et al. \cite{benevenuto} identify 
-social attributes of spam and ham on video SNSs (in this case, YouTube), including video 
-view counts, comment counts and user public profile attributes. They then use a 
-support vector machine (SVM) for classification, yielding 96\% accuracy in detecting 
+social attributes of spam and ham on video SNSs (in this case, they scraped data from 
+YouTube), including video view counts, comment counts and user public profile attributes. 
+They then use a support vector machine (SVM) for classification, yielding 96\% accuracy in detecting 
 advertisers (``promoters"), but only accurately identifying 57\% of examples of more general spam.
@@
 associative classification, in which, for example, a message would be classified as 
-spam if it contained a link to a web page that had a high probability of being spam. 
+spam if it contains a link to a web page that has a high probability of being spam. 
 Unfortunately, the differing characteristics of various social networks \eg the length 
@@
 against phishing, fraud and spam. The system is composed of classifier services, an 
-ML-derived Feature Extraction Language (FXL), feature loops to aggregate and prepare 
-features for classification and a policy engine to take action on suspected misuse. While 
+ML-derived Feature Extraction Language (FXL), ``feature loops" or services that aggregate 
+and prepare features for classification and a policy engine to take action on suspected misuse. While 
 the discussion remains high-level and includes few implementation particulars, it does 
@@
 engine based on lightweight features such as geographic distance between sender and 
-receiver, geolocation anomalies and diurnal patterns. While the target was conventional 
-spam, these and similar features are applicable to spam on SNSs as well.
+receiver and diurnal patterns. While the target was conventional 
+spam, monitoring sender reputation and using similar features (\eg time-of-day when messages 
+were sent), seesm applicable to spam on SNSs as well.
 
@@
 Blog comment spam have also attracted considerable attention from researchers who 
-have applied machine learning \cite{kolari, nag} and NLP \cite{mishne} techniques to the 
-problem. Likewise, Markines et al. \cite{markines} apply similar techniques to spam on 
-social bookmarking sites.
+have applied machine learning \cite{kolari, nag} techniques including SVMs and Bayesian classifiers.
+Mishe et al.\cite{mishne} employ language modeling to find semantic discrepancies between 
+the blogs on which link spam comments are posted and the target sites (which might, for example, 
+contain adult content). Likewise, Markines et al. \cite{markines} apply similar techniques 
+including SVMs and boosting (AdaBoost) to spam on social bookmarking sites.
 
@@
 researchers tackle this adversarial problem. Dalvi et al. \cite{dalvi} offer a 
-modified Naive Bayes classifier to detect and reclassify data taking into account 
-the optimal modification strategy that an adversary could choose. Lowd and Meek \cite{lowd} 
-provide a framework for reverse engineering a classifier to determine whether 
-an adversary can efficiently learn enough about a classifier to effectively defeat it.
+modified naive Bayes classifier to detect and reclassify data taking into account 
+the optimal modification strategy that an adversary could choose, given full knowledge of 
+the classifier model and parameters, but without the ability to tamper with the 
+classifier's training data. The authors show that the optimized classifier's counter-strategy 
+is substantially more effective than a standard classifers in the cases examined: 
+adversaries adding words, adding length and substituting words with synonyms. 
+Lowd and Meek \cite{lowd} provide a framework for reverse engineering a 
+classifier to determine whether an adversary can efficiently learn enough 
+about a classifier to effectively defeat it.
 
@@
 
-For this project, we had unlimited access to data from InterPals, a SNS for users who 
-wish to communicate with others from other countries, whether for language practice, cultural 
+For this project, we had unlimited access to data from InterPals, a site for users who 
+wish to communicate with people from other countries, whether for language practice, cultural 
 exchange or friendship. Users sign up by completing a registration form with 
 information about themselves, including age, sex and location. After registering, users 
-can expand their profile page to include descriptions, interests, languages they speak, etc. 
-and upload photographs. After clicking on a link sent to their email address, a user can 
+can expand their profile page to include self-descriptions, interests, languages they speak,  
+photographs, etc. After clicking on an activation link sent to their email address, a user can 
 begin to interact with others on the site via private message, public ``wall" posts, comments 
@@
 through volunteer moderators. Users can report content, including private messages, 
-profiles and photographs to moderators, using a form that includes a drop-down menu of 
+profiles and photographs, to moderators by using a form that includes a drop-down menu of 
 pre-selected reasons with the option to add a more detailed message in a text field. 
-Likewise, the message interface allows a user to report a private message spam with a single 
+Likewise, the message interface allows a user to report a private message as spam with a single 
 click. Moderators have access to a queue of these reports. In addition to the material 
@@
 user's account, including outgoing private messages, IP addresses, as well as a list of 
-other users who have logged in from the same computer (determined via IP address as well 
-as by browser fingerprints). Moderators can then decide to delete the user, send a warning, 
+other users who have logged in from the same computer (determined via IP address, cookies, as well 
+as by rudimentary browser fingerprints). Moderators can then decide to delete the user, send a warning, 
 or clear reports on a user. All moderator actions require them to annotate their decision with a 
 brief log message. When deleting a user, moderators have the option of flagging the 
-reported message (if there is one) as spam, or of flagging all of the user's outgoing private messages 
+reported message (if there is one) as spam or of flagging all of the user's outgoing private messages 
 as spam.
@@
 Other anti-spam measures include widespread use of CAPTCHAs across the site and 
-frequency caps. Short-window frequency caps are in place for all users, limiting the 
+frequency caps on activities that involve contact with other users. Short-window 
+frequency caps are in place for all users, limiting the 
 number of messages that can be sent per short time interval (1, 5 and 10 minutes). New users 
@@
 
-At the beginning of this research project, slightly two million messages had 
+At the beginning of this research project, slightly over two million messages had 
 been flagged by moderators as spam. We extracted the contents of exactly 2 million 
@@
 Account information for deleted users is stored for six months, so account details for all spam 
-and ham accounts was available. Extracting the data of interest to this project required 
+and ham accounts were available. Extracting the data of interest to this project required 
 dumping the contents from a query with multiple joins and aggregations across 8 tables with a 
@@
 user logs in to the site. By examining the distribution of countries (using the MaxMind GeoIP
-database) associated with the unique IP addresses (Figure \ref{fig:cip}), we see that the top 
+database\cite{maxmind}) associated with the unique IP addresses (Figure \ref{fig:cip}), we see that the top 
 countries for spammers are significantly different, while the top ham countries are 
@@
 such as Marr wavelet, discrete cosine transform (DCT), color histograms, etc., which provided by the 
-pHash library, \footnote{http://www.phash.org} seem to promising in detecting image re-use.
+pHash library\cite{phash}, seem to promising in detecting image re-use.
 

### 2012-05-09 docs/paper/paper.tex
update

@@
 combined size of 154 GB. To minimize the impact on the load of the site's production database 
-instances, we took an LVM snapshot of slave instance and cloned the database to a spare server. 
+instances, we took an LVM snapshot of a replication instance (the production database has one master and one 
+slave) and cloned the database to a spare server. 
 In addition to using SQL to extract the data, we wrote Perl scripts to clean the data, tokenize 
-messages and prepare data for further processing. The methodology section (X) offers more details 
+messages and prepare data for further processing. The methodology section offers more details 
 on precisely which items of data were collected and why.
@@
 Consequently, we extracted a working set of 2 million ham messages that were sent in late December 2011 and early 
-January 2012 and still existed in the recipient's inbox in March 2012. To reduce the 
+January 2012 and still existed in the recipient's inbox as of March 2012. To reduce the 
 possibility of collecting messages sent by uncaught spammers to inactive or dormant users, 
-we selected messages only from users and to recipients who had logged in within the last two 
-weeks. As in the case of spam, we collected two million messages with associated account 
+we selected messages only from senders and to recipients who had logged in within the last two 
+weeks. As in the case of the spam data set, we collected two million messages with associated account 
 data and merged them into text files for feature extraction.
@@
 
-In the course of collecting the data, we observed a number of different types of 
+In the course of collecting the data, we observed a number of distinct classes of 
 undesirable messages. Our ongoing observation of spam on the InterPals website provided 
@@
 \begin{itemize}
-\item \textbf{Advance Fee Fraud} 
-\item \textbf{Romance Scams} 
-\item \textbf{Sexual Solicitation}
-\item \textbf{Porn Ads}
-\item \textbf{Advertising External Sites}
-\item \textbf{Webcam Ads}
-\item \textbf{Money Muling}
-\item \textbf{Begging \& Gift Requests}
-\item \textbf{Business Proposals}
+\item \textbf{Advance fee fraud}, including inheritance, lotter, visa and customs-clearance scams
+\item \textbf{Romance scams}, including ``mail-order bride" and military scams
+\item \textbf{Sexual solicitation}
+\item \textbf{Ads for porn sites}, primarily adult webcam and live chat sites.
+\item \textbf{Ads for miscellaneous external sites}, often other SNSs
+\item \textbf{Money muling}, often in the guise of high-paying ``stay-at-home" jobs or ``mystery" shopping
+\item \textbf{Begging \& gift requests}
+\item \textbf{Business proposals}
 \end{itemize}
@@
 spam-labeled messages. We plan to quantify the volume of messages in each of these categories and 
-attempt to provide a finer granularity of categorization in future research.
+attempt to provide a finer granularity of categorization (and more detailed level of description) in 
+future research.
 
@@
 We chose to focus on spam in private messages for this paper, given that messages account for 
-a majority of user spam reports. In addition to using the bag-of-words representation of 
-the message content, we aimed to identify a subset of additional, relevant ``expert" features 
+a majority of user spam reports on the site. In addition to using the bag-of-words representation of 
+the message content, we aimed to identify a subset of relevant ``expert" features 
 based on public and non-public message and account information that would augment the classifier's 

### 2012-05-09 docs/paper/paper.tex
update

@@
 spam messages would, for example, put spaces before commas, periods and quotation marks, 
-and omit spaces after these characters). Second, we calculated the ratio of uppercase to 
+while omitting spaces after these characters). Second, we calculated the ratio of uppercase to 
 lowercase letters in the text, after observing relatively high amounts of uppercase text in 

### 2012-05-09 docs/paper/paper.tex
complete draft

@@
 recipient, as listed on their profile. Unfortunately, due to data collection issues, this data was only available 
-for the most recent 11,000 spam messages. While the male-to-female ratio of the recipients was very close to that 
-of non-spam recipients (both were 47\% male to 53\% female), we found that the age of recipients was typically higher.
+for the most recent 11,000 spam messages. While the male-to-female ratio of the spam recipients was very close to that 
+of non-spam recipients (both were 47\% male to 53\% female), we found that the age of recipients was typically higher 
+the average ham recipient age.
 
@@
 revealed by their IP address differed remarkably between the spam and ham user groups. 
-48\% of spam users claimed to be from the United States, with the next most popular 
-countries being the UK, Ghana, Senegal, Germany and Canada taking combined 25\%. 
-In contrast, only 15\% ham users claimed to be from the United States, with Russia, 
-Korea, UK, France and Germany combining for 24\%. This distribution of profile countries can 
-be seen in Figure \ref{fig:cprof}.
+48\% of spam users claimed to be from the United States, with the UK, Ghana, Senegal, 
+Germany and Canada comprising the next 25\%.  In contrast, only 15\% ham users claimed to be 
+from the United States, with Russia, Korea, UK, France and Germany making up the next 24\%. 
+This distribution of profile countries can be seen in Figure \ref{fig:cprof}.
 
@@
 Users are free to choose any country on their profile. We record each IP address from which a 
-user logs in to the site. By examining the distribution of countries (using the MaxMind GeoIP
-database\cite{maxmind}) associated with the unique IP addresses (Figure \ref{fig:cip}), we see that the top 
+user logs in to the site. By examining the distribution of countries  associated with the unique IP addresses
+(using the MaxMind GeoIP database\cite{maxmind}), we see that the top 
 countries for spammers are significantly different, while the top ham countries are 
-virtually unchanged. Furthermore, only 30\% of the IP addresses of spam users who 
+virtually unchanged (Figure \ref{fig:cip}). Furthermore, only 30\% of the IP addresses of spam users who 
 claimed to be from the United States actually mapped back to a US-based ISP, with a combined 
@@
 
-\textbf{Sender IPs:} Spammers in general had a smaller number of unique IP addresses associated 
-with them than ham users, as per figure \ref{fig:uniqip}. The median number of IP addresses for 
+\textbf{Sender IPs:} Spammers in general logged in from a smaller number of unique IP addresses 
+than did ham users (Figure \ref{fig:uniqip}). The median number of IP addresses for 
 spam users was 2, whereas ham users saw a median of 64. This is likely due to the comparatively 
 shorter length of time that spammers remain on the site before being deleted by moderators (Figure 
-\label{fig:lifetime} show a histogram of the time elapsed between registration and message sending 
+\ref{fig:lifetime} shows a histogram of the time elapsed between registration and message sending 
 time).  Similarly, we believe that many legitimate users are behind NATs, yielding a high number 
@@
 
-\textbf{Sender Email:} Users must provide email addresses upon signup. The site does not allow users to contact other 
-members before verifying that the email address provided is valid, via an account activation link sent to the signup 
-address. Subsequent changes to a user's email address on file require a similar confirmation process. Thus, we can be certain 
+\textbf{Sender Email:} Users must provide valid email addresses upon signup and these addresses are 
+verified by confirmation links before users can communicate with other site members. Subsequent changes 
+to a user's email address on file require a similar confirmation process. Thus, we can be certain 
 that the email accounts that we extracted were in use by the senders, at least at the time of registration.
@@
 In the data processing step, we extracted the domain name from each email address. We found that ham user email domains 
-were distributed across top email providers in similar proportions (Hotmail, Yahoo and Gmail accounted for 22\%, 17\% and 17\% 
-respectively). For spam user accounts, Hotmail, Yahoo and Gmail accounted for 7\%, 72\% and 6\% respectively. This striking 
+were distributed across top email providers in similar proportions: Hotmail, Yahoo and Gmail accounted for 22\%, 17\% and 17\% 
+respectively. For spam user accounts, Hotmail, Yahoo and Gmail accounted for 7\%, 72\% and 6\% respectively. The striking 
 predominance of Yahoo email accounts among spammers can be seen in Figure \ref{fig:email}. While we have no certain 
@@
 age of message recipients was higher for spam recipients that for recipients of ham messages, with 
-median ages of 41 and 22 (Figure \ref{fig:recipage}). The sample size for spam recipient age 
+median ages of 41 and 22 (Figure \ref{fig:recipage}). The sample size for spam recipient ages 
 was only 11,000 messages due to late data collection. We posit that the higher ages indicated on spam profiles, 
@@
     \caption{Age of message recipient as stated on user profile (Note that 2 million message spam sample 
-        had only 11,000 messages with this information)}
+        had only 11,000 messages containing these fields.)}
     \label{fig:recipage}
@@
 observed in the countries with the most discrepancies between IP-detected countries and countries stated on spam 
-user profiles, with Afghanistan and Albania comprising 19\%, despite their relatively low representation on the 
-site.
+user profiles, with Afghanistan and Albania listed on 19\% of such profiles, despite their relatively low 
+representation on the site.
 
@@
 however, that we would expect a legitimate user's average number of photos and friends to grow over 
-as a function of account lifetime. It would be quite reasonable for new ham users to have very photos 
+as a function of account lifetime. It would be reasonable for new ham users to have very few photos 
 or friends and we plan to further examine the relationship of these attributes over account lifetimes 
@@
 
-\textbf{Username, Name:} While we collected these fields, we did not analyze them further, due to 
-time constraints. Nevertheless, we saw considerable repetition in the usernames of scammers 
-and it seems that analyzing the substrings in these fields could yield useful features.
+\textbf{Username \& Name:} While we collected these fields, we did not analyze them further, due to 
+time constraints. Nevertheless, we observed considerable repetition in the usernames of scammers 
+and we expect that analyzing the substrings in these fields could yield useful features.
 
@@
 frequencies in descending order, mapping each feature (word or expert feature value) to an integer 
-key. The second step was to create a sparse matrix representation of each document (in this case, 
-spam or ham message) and the associated features. Spark provides a parallelized in-memory computation 
+key. The second step was to create a sparse matrix representation of each document, in this case, 
+spam or ham message, and the associated features. Spark provides a parallelized in-memory computation 
 framework that dramatically reduced the time necessary to extract these features, build the dictionary 
@@
 
-We did not use all of the expert features available to us, due to time constraints. Likewise, for 
-many expert features could have a wide range of values, which would expand the dictionary size 
-considerably. In addition to the bag-of-words from the message text, we used the following features:
+We did not use all of the expert features described in the previous section, due to time constraints. Likewise, 
+many expert features have a wide range of possible values, which would expand the dictionary size 
+considerably. This can usually be addressed by bucketing values or converting them to binary features, 
+which we did for several features at the expense of some loss of specificity. In addition to 
+the bag-of-words from the message text, we used the following features:
 \begin{itemize}
@@
 \item Whether sender has friends, boolean
-\item Whether sender has photo(s), boolean
+\item Whether sender has photos, boolean
 \item Profile country and IP-detected country match, boolean
@@
 The main idea of the Naive Bayes classification algorithm is that one can calculate the probability 
-of an item belonging to a certain class by applying Bayes' theorem on the assumption that the probability 
-of one word/feature that appears in a document is independent of the probability of any other feature. 
+that an item belongs to a certain class by applying Bayes' theorem. This theorem assumes that the probability 
+that one feature appears in a document is independent of the probability of any other feature also appearing.  
 While is very rarely the case (hence ``naive"), Naive Bayes still performs surprisingly well in many cases.
@@
 
-One of the advantages of the Naive Bayes classifier was that it took very little time to train the classifier, 
+One of the advantages of the Naive Bayes classifier was that it took very little time to train the classifier 
 and the steps involved were all trivially parallelizable, which makes it convenient for large data sets. 
@@
 
-In this project, we performed ordinary least squares regression. By minimizing the sum of squared 
+In this project, we perform ordinary least squares regression. By minimizing the sum of squared 
 residuals, we can solve for the unknown parameter vector $\beta$ using a closed form solution. 
@@
 
-We used ridge regression, which gave us the opportunity to tune a $\lambda$ parameter in order to 
+Using ridge regression also gave us the opportunity to tune a $\lambda$ parameter in order to 
 regularize the vector components for better performance.
@@
 29.216 for the bag-of-words LR model. After adding the expert features, the mean AUC climbed to 0.951 with a mean 
-lift of 41.584. Figure \ref{fig:roclin} show ROC curves for 1-fold cross-validation on both of these models.
-
-As shown in figure \ref{fig:lambda}, we tested a number of values for $\lambda$, the ridge regression parameter. 
-Since the AUCs corresponding to $\lambda$ for both models appear to peak between 30 and 50, we may have been able to 
-get a slightly higher AUC by using a lower $\lambda$.
+lift of 41.584. Figure \ref{fig:roclin} shows ROC curves for 1-fold cross-validation on both of these models.
+
+As shown in Figure \ref{fig:lambda}, we tested a number of values for $\lambda$, the ridge regression parameter. 
+Since the AUCs corresponding to $\lambda$ for both models appear to peak between 30 and 50, we might have been able to 
+get a slightly higher AUC by using a lower value for $\lambda$.
 
@@
 
-Support Vector Machines are a classification method that maps class examples (\eg messages) to points 
-in space and aims to maximize the margin around a hyperplane separating the classes.\cite{boser}, \cite{cortesv95} 
+SVMs are a classification method that maps class examples (\eg messages) to points 
+in space and aims to maximize the margin around a hyperplane separating the classes.\cite{boser, cortesv95} 
 Because the sets of points may not be linearly separable in the original dimension, a kernel trick can 
@@
 The optimizer runs stochastic subgradient descent on the primal objective 
-using batches provided. However, the ScalaNLP SVM interface does not allow changing kernel functions or parameters, 
-which seems to have led to the classifier's sub-par performance.
+using the batches provided. However, the ScalaNLP SVM interface does not allow users to change kernel 
+functions or parameters, which seems to have led to the classifier's sub-par performance.
 
@@
 Python and other languages. The package allows one to specify the type of kernel to use, kernel parameters 
-($\gamma$, $\rho$, $d$), as well as a soft margin parameter ($C$). We used Perl to convert the sparse feature 
-matrices generated for use with ScalaNLP to a format recognized by LIBSVM. We then used one of the package's 
+($\gamma$, $\rho$, $d$), as well as a soft margin parameter ($C$). We used Perl to convert our existing sparse feature 
+matrices generated for use with ScalaNLP to a format recognized by LIBSVM. We then ran one of the package's 
 tools to perform simple scaling on the data.
@@
 $$K(\mathbf{x}_i, \mathbf{x}_j) = (\gamma\mathbf{x}_i^\intercal\mathbf{x}_j+r)^d, \gamma>0$$
-We ran this script on a subset of 5000 messages with expert features to obtain $C = 8$ and $\gamma = 0.0078125$, 
-which we used to tune the SVM we used to classify both the bag-of-words and expert feature models.
+We ran this script with 10-fold cross-validation on a subset of 5000 messages with 
+expert features to obtain $C = 8$ and $\gamma = 0.0078125$, which we used to tune 
+the SVM we used to classify both the bag-of-words and expert feature models.
 
@@
 Our initial results using the Pegasos SVM solver included in ScalaNLP were lower than expected. The 
-classifier trained on the bag-of-words feature set (from 22k messages) yielded a 77\% mean accuracy 
+classifier trained on the bag-of-words feature set (22k messages) yielded a 77\% mean accuracy 
 over 10-fold cross-validation. The addition of expert features actually decreased this accuracy to 
@@
 size from 2000 to 100k messages. Table \ref{tab:svmacc} shows results of 10-fold cross-validation. Our mean 
-accuracy over the bag-of-words data sets was 85.73\% and 98.41\% over the bag-of-words with expert features. We 
-observed a slight decrease in performance on the bag-of-words data sets as the data size grew. Figure \ref{fig:roc-25}
+accuracy over the bag-of-words data sets was 85.73\% and 98.41\% for the bag-of-words with expert features. We 
+observed a slight decrease in performance on the bag-of-words data sets as the data size grew, but do not 
+have a hypothesis as to why this occurred. Figure \ref{fig:roc-25}
 shows the AUCs and ROC curves for the 50k data sets, making clear the the significantly greater AUC resulting 
@@
 of wall posts and photo comments, making it possible to investigate the characteristics of spam in these genres. 
-Wall posts and comments tend to be comparatively short, which should present a slightly different classification 
+Wall posts and comments tend to be comparatively short, which should present a somewhat different classification 
 problem.
@@
 As a part of this project, we performed some qualitative analysis on spam and differentiated between 
-some major categories of undesirable messages that we observed on the InterPals web site. However, we would 
+some major categories of the undesirable messages that we observed on the InterPals web site. However, we would 
 like to gain a better understanding of the relative volume of each type of scam. To this end, we plan to 
@@
 Furthermore, characteristics of spam messages vary substantially depending on the type of spam or scam in question. 
-The classifiers that we trained and evaluated in this project all spam messages as a single category. We are currently 
+The classifiers that we trained and evaluated in this project treat all spam messages as a single category. We are currently 
 investigating multiclass classifiers as a potential tool to identify the type of 
@@
 From a site's perspective, this finer granularity is desirable not only 
-because specialty classifiers could be trained to recognize specific classes of spam, but also in order to 
+because specialty classifiers could be trained to recognize specific classes of spam more effectively, but also in order to 
 trigger different courses of action (\eg automatic account deletion, warning messages, reports to human moderators) 
@@
 would include both the ones we extracted in this project but omitted from the classifier features matrices, 
-as well as new features. Adding N-grams and message similarity tests seem particularly promising. 
+as well as new features. Adding N-grams and message similarity tests seems particularly promising. 
 N-grams would allow the classifier to take into account certain phrases \eg{``Western Union" or ``money transfer"}. 
-
-We have also recently begun to collect browser fingerprints from users based on user agent 
+Using cosine similarity tests, data-punctutaed token trees or computing Jaccard indexes could allow us to 
+detect whether a user has sent similar messages to multiple recipients in the past.
+
+We have also recently begun to collect browser fingerprints from users based on user-agent strings 
 and JavaScript-acquired browser plugin details, time zone, screen size, color depth, system fonts, 
 keyboard layout and locale information. Subsequent classifiers could leverage this information either 
-as a hash that effectively tags individual known spammers or in part, adding features based on 
+as a hash that effectively tags individual known spammers or in discrte parts, adding features based on 
 time zone or keyboard layout.
@@
 such as Marr wavelet, discrete cosine transform (DCT), color histograms, etc., which provided by the 
-pHash library\cite{phash}, seem to promising in detecting image re-use.
+pHash library\cite{phash}, seem promising in detecting image re-use.
 

### 2012-05-09 docs/paper/paper.tex
reorg
typo

@@
 \begin{itemize}
-\item \textbf{Advance fee fraud}, including inheritance, lotter, visa and customs-clearance scams
+\item \textbf{Advance fee fraud}, including inheritance, lottery, visa and customs-clearance scams
 \item \textbf{Romance scams}, including ``mail-order bride" and military scams
@@
 
-\section{Classifiers \& Results}
+\section{Classifiers}
 
@@
 
-\subsection{Naive Bayes Results}
-
-The Naive Bayes classifier had an accuracy of 82\% when trained on the bag-of-words. Its accuracy dropped 
-to 77\% when expert features were added. The decline in accuracy after adding expert features is most likely 
-explained by the ``naive" assumption in this algorithm that the features are independent. Several of 
-the features that we used, \eg{account lifetime and whether a user has friends or photos}, are in 
-fact likely to be co-dependent, leading to an overweight of these features during classification.
-
-One of the advantages of the Naive Bayes classifier was that it took very little time to train the classifier 
-and the steps involved were all trivially parallelizable, which makes it convenient for large data sets. 
-However, in view of the wealth of expert features available to a SNS, many of which are not independent, this was 
-the least compelling classifier choice that we examined.
-
 \subsection{Linear Regression}
@@
 regularize the vector components for better performance.
+
+\subsection{Logistic Regression}
+
+We attempted to train the multiple logistic regression classifier implemented by the ScalaNLP package, 
+but encountered a bug that prevented the model from being generated correctly. With the expectation 
+that SVMs would outperform logistic regression, we decided to focus our efforts on tuning the SVM 
+instead of pursuing logistic regression.
+
+\subsection{Support Vector Machines (SVMs)}
+
+SVMs are a classification method that maps class examples (\eg messages) to points 
+in space and aims to maximize the margin around a hyperplane separating the classes.\cite{boser, cortesv95} 
+Because the sets of points may not be linearly separable in the original dimension, a kernel trick can 
+be used to fit the maximum-margin hyperplane in a higher-dimensional feature space (which may appear 
+non-linear in the original input space).
+
+We initially used ScalaNLP's built-in SVM solver, which implements the Pegasos maximization algorithm\cite{pegasos} 
+as extended by Wang, Crammer and Vucetic\cite{wang2}. 
+The optimizer runs stochastic subgradient descent on the primal objective 
+using the batches provided. However, the ScalaNLP SVM interface does not allow users to change kernel 
+functions or parameters, which seems to have led to the classifier's sub-par performance.
+
+We then turned to LIBSVM, a library for SVM training and classification in C++ and Java, with wrappers for 
+Python and other languages. The package allows one to specify the type of kernel to use, kernel parameters 
+($\gamma$, $\rho$, $d$), as well as a soft margin parameter ($C$). We used Perl to convert our existing sparse feature 
+matrices generated for use with ScalaNLP to a format recognized by LIBSVM. We then ran one of the package's 
+tools to perform simple scaling on the data.
+
+Given the complexity and time-intensive nature of choosing and testing parameters, the 
+package includes a Python script that performs cross-validation to suggest appropriate 
+$\gamma$ and $C$ values for the recommended Gaussian radial basis function (RBF) kernel. The RBF kernel is 
+a real-valued function such that: 
+$$K(\mathbf{x}_i, \mathbf{x}_j) = (\gamma\mathbf{x}_i^\intercal\mathbf{x}_j+r)^d, \gamma>0$$
+We ran this script with 10-fold cross-validation on a subset of 5000 messages with 
+expert features to obtain $C = 8$ and $\gamma = 0.0078125$, which we used to tune 
+the SVM we used to classify both the bag-of-words and expert feature models.
+
+\section{Performance \& Results}
+
+\subsection{Naive Bayes Results}
+
+The Naive Bayes classifier had an accuracy of 82\% when trained on the bag-of-words. Its accuracy dropped 
+to 77\% when expert features were added. The decline in accuracy after adding expert features is most likely 
+explained by the ``naive" assumption in this algorithm that the features are independent. Several of 
+the features that we used, \eg{account lifetime and whether a user has friends or photos}, are in 
+fact likely to be co-dependent, leading to an overweight of these features during classification.
+
+One of the advantages of the Naive Bayes classifier was that it took very little time to train the classifier 
+and the steps involved were all trivially parallelizable, which makes it convenient for large data sets. 
+However, in view of the wealth of expert features available to a SNS, many of which are not independent, this was 
+the least compelling classifier choice that we examined.
 
@@
 
-\subsection{Logistic Regression}
-
-We attempted to train the multiple logistic regression classifier implemented by the ScalaNLP package, 
-but encountered a bug that prevented the model from being generated correctly. With the expectation 
-that SVMs would outperform logistic regression, we decided to focus our efforts on tuning the SVM 
-instead of pursuing logistic regression.
-
-\subsection{Support Vector Machines (SVMs)}
-
-%intro to svms
-
-SVMs are a classification method that maps class examples (\eg messages) to points 
-in space and aims to maximize the margin around a hyperplane separating the classes.\cite{boser, cortesv95} 
-Because the sets of points may not be linearly separable in the original dimension, a kernel trick can 
-be used to fit the maximum-margin hyperplane in a higher-dimensional feature space (which may appear 
-non-linear in the original input space).
-
-We initially used ScalaNLP's built-in SVM solver, which implements the Pegasos maximization algorithm\cite{pegasos} 
-as extended by Wang, Crammer and Vucetic\cite{wang2}. 
-The optimizer runs stochastic subgradient descent on the primal objective 
-using the batches provided. However, the ScalaNLP SVM interface does not allow users to change kernel 
-functions or parameters, which seems to have led to the classifier's sub-par performance.
-
-We then turned to LIBSVM, a library for SVM training and classification in C++ and Java, with wrappers for 
-Python and other languages. The package allows one to specify the type of kernel to use, kernel parameters 
-($\gamma$, $\rho$, $d$), as well as a soft margin parameter ($C$). We used Perl to convert our existing sparse feature 
-matrices generated for use with ScalaNLP to a format recognized by LIBSVM. We then ran one of the package's 
-tools to perform simple scaling on the data.
-
-Given the complexity and time-intensive nature of choosing and testing parameters, the 
-package includes a Python script that performs cross-validation to suggest appropriate 
-$\gamma$ and $C$ values for the recommended Gaussian radial basis function (RBF) kernel. The RBF kernel is 
-a real-valued function such that: 
-$$K(\mathbf{x}_i, \mathbf{x}_j) = (\gamma\mathbf{x}_i^\intercal\mathbf{x}_j+r)^d, \gamma>0$$
-We ran this script with 10-fold cross-validation on a subset of 5000 messages with 
-expert features to obtain $C = 8$ and $\gamma = 0.0078125$, which we used to tune 
-the SVM we used to classify both the bag-of-words and expert feature models.
-
 \subsection{SVM Results}
-
 Our initial results using the Pegasos SVM solver included in ScalaNLP were lower than expected. The 
